### 2011-08-04 test/DESeq/suppl_ii.Rnw
DESeq is used for OMZ175.

\documentclass{article}

\usepackage{Sweave}
\usepackage[a4paper]{geometry}
\usepackage{graphicx}

% Labels, hard-coded to make sure they stay fixed
\input{labels.tex}

\SweaveOpts{keep.source=TRUE,eps=FALSE,width=4,height=4}

% meant to be run with R 2.11, DESeq 1.1.6, edgeR 1.6.0

\begin{document}

\begin{minipage}{\textwidth}
\center
{\large{\textbf{\textsf{Differential expression analysis for sequence count data}}}}
\medskip

Simon Anders, Wolfgang Huber
\bigskip

\LARGE{\textsc{\textsf{\textbf{Supplement II}}}}
\vspace{4ex}
\end{minipage}

\noindent In order to make the analyses presented in our paper transparent and reproducible, 
we document in the present supplement the complete R code used to perform
them and to create all the plots in the paper and in Supplement I. The purpose of this 
document is \emph{not} to explain the use of the \emph{DESeq} package; for this, 
please read the vignette (manual) supplied with the package. 
\medskip

\section{Required packages}

We start with loading all requried packages:
<<results=hide>>=
library( DESeq )
library( edgeR )
library( hexbin )
library( latticeExtra )
library( gplots )
library( geneplotter )
@

This Sweave file has been run with these package versions:

<<>>=
sessionInfo()
@

\section{Reading in the data}

All data files necessary to run the code shown here are supplied in the supplementary data 
tarball available with the paper.

\subsection{Fly RNA-Seq data}

The RNA-Seq data from gene overexpression in fly embryos is unpublished, an excerpt
of which was kindly provided to us by Wilczynski et al.\ (Furlong Lab, EMBL Heidelberg)
in the form of the following table (file \texttt{fly\_RNA\_counts.tsv}, supplied 
with supplementary data):

<<>>=
countsTableFly <- read.delim( "fly_RNA_counts.tsv" )
condsFly <- c( "A", "A", "B", "B" )
# Add dummy names to avoid confusion later
rownames( countsTableFly ) <- paste( "Gene", 1:nrow(countsTableFly), sep="_" )
@

\subsection{Neural stem cell RNA-Seq data}

The RNA-Seq data comparing neural stem cell lines from gliablastomas and non-cancerous
tissue was kindly provided to us by Engstr\"om et al.\ (Bertone Lab, EMBL-EBI Hinxton)
in the form of the following table (file \texttt{NeuralStemCellData.tab}, supplied 
with supplementary data):

<<>>=
countsTableNS <- read.delim( "NeuralStemCellData.tab", row.names=1 )

head(countsTableNS)
@

The conditions are

<<>>=
condsNS <- c( "GNS_dupe", "GNS", "GNS", "GNSL", "NS", "NS" )
@

where the first entry is for the sample that will be eliminated later as it is derived
from the same patient as the second one.

\subsection{Yeast RNA-Seq data}

We have downloaded the deposited reads for the yeast RNA-Seq experiment by Nagalakshmi et al.\
from the NCBI Short Read Archive, remapped them to the yeast genome with Bowtie\footnote{Langmead et al.,
Ultrafast and memory-efficient alignment of short DNA sequences to the human genome. Genome Biol.\ 10, R25} 
and the counted the coverage of each gene with 
htseq-count\footnote{http://www-huber.embl.de/users/anders/HTSeq/doc/count.html}.

We read in this data with

<<>>=
countsYeast <- read.delim( "counts_Nagalakshmi_et_al.tab", row.names=1 ) 
head( countsYeast )
@

\section{ChIP-Seq data}

The supplementary file \texttt{PolII\_counts.tsv} contains one column for each sample
in the Kasowski et al.\ paper, and one row for each polymerase-II binding region
that the authors identified, giving the number of reads counted to map within each binding
region. This file has been derived from the original data file in Kasowski et al.'s 
supplementary data with a simple reformatting.

We read in the data with

<<>>=
countsPolII <- read.delim( "PolII_counts.tsv" )

# make rownames
row.names(countsPolII) <- sprintf( "%s:5d-%d", countsPolII$chrom, 
   countsPolII$start, countsPolII$stop )

# The IDs should start with "GM", not "X":
colnames(countsPolII) <- sub( "X", "GM", colnames(countsPolII) )

# A few individuals in this data file have four or more replicates. We have picked 
# (at random) two of them:
countsPolII <- countsPolII[ , grepl( "GM12878", colnames(countsPolII) ) | 
   grepl( "GM12891", colnames(countsPolII) ) ]

head(countsPolII)
@

\section{Analysis of fly data}

\subsection{Analysis with DESeq}

For the standard DESeq analysis, we instantiate a \emph{CountDataSet}, then estimate
the size factors and the variance functions, and finally perform the binomial test.

<<>>=
cdsFly <- newCountDataSet( countsTableFly, condsFly )
cdsFly <- estimateSizeFactors( cdsFly )
cdsFly <- estimateVarianceFunctions( cdsFly )
resFly <- nbinomTest( cdsFly, "A", "B" )
@

\subsection{Analysis with edgeR}

For the edgeR analysis, we test for different modes: One choice is whether edgeR 
should run in the common (``C'') or the tag-wise (``T'') dispersion mode, the other
is whether edgeR should be given the actual number of sequenced reads (``N''), as
suggested in the edgeR manual, or the same size factors\footnote{As edgeR expects read counts, we 
multiply these by $10^7$ so that they get a plausible magnitude. edgeR only 
cares about the relative ratios of the \texttt{lib.size} values, and hence, this factor 
is unimportant. However, it expect them to be large numbers (it seems to round them 
to integers), and hence, the ``\texttt{1e7*}'' below ensures that no information is lost in 
rounding. Using another large number does not change the results.} as DESeq used (``D'').

We instantiate four \emph{DGEList} objects and perform the analysis four times:

<<>>=
# using DESeq's size factors:

dglFlyD <- DGEList( counts=countsTableFly, group=condsFly, lib.size=1e7*sizeFactors(cdsFly) )
dglFlyD <- estimateCommonDisp( dglFlyD )
dglFlyD <- estimateTagwiseDisp( dglFlyD )

# test with tag-wise dispersion:

edgerResFlyDT <- exactTest( dglFlyD, common.disp=FALSE )
edgerResFlyDTPadj <- p.adjust( edgerResFlyDT$table$p.value, method="BH" )

# common dispersion, DESeq's size factors:

edgerResFlyDC <- exactTest( dglFlyD, common.disp=TRUE )
edgerResFlyDCPadj <- p.adjust( edgerResFlyDC$table$p.value, method="BH" )

# with the original library sizes:

# the total number of sequenced counts were as follows:
libsizesFly <- c( A1=13613446, A2=12106517, B1=12920058, B2=11599459 )

dglFlyN <- DGEList( counts=countsTableFly, group=condsFly, lib.size=libsizesFly )
dglFlyN <- estimateCommonDisp( dglFlyN )
dglFlyN <- estimateTagwiseDisp( dglFlyN )

# test with tag-wise dispersion:

edgerResFlyNT <- exactTest( dglFlyN, common.disp=FALSE )
edgerResFlyNTPadj <- p.adjust( edgerResFlyNT$table$p.value, method="BH" )

# common dispersion, DESeq's size factors:

edgerResFlyNC <- exactTest( dglFlyN, common.disp=TRUE )
edgerResFlyNCPadj <- p.adjust( edgerResFlyNC$table$p.value, method="BH" )
@

\subsection{Plots and comparison}

\subsubsection{Figures on variance estimation}

We start with the plots of Fig.\ \ref{varestFig} in the paper.

The common dispersion values estimated by edgeR are:

<<>>=
dglFlyD$common.dispersion
dglFlyN$common.dispersion
@

The values for the two library size settings hardly differ, we shall use the first one
for the dashed orange line.

<<>>=
alpha <- dglFlyD$common.dispersion
@

The \emph{DESeq} package offers the function \texttt{rawVarFunc}, that
returns the fitted raw variance function for a condition $\rho$, i.e., 
the function $\hat v_\rho(q)$ defined in Eq.\ (9) of the paper. For the solid orange line
in the plot, we define a function \texttt{basevarFunc} that returns the fitted variance
function $w_\rho(q)$ by undoing the subtraction of Eq.\ (9) that \texttt{rawVarFunc}
does internally:

<<>>=
baseVarFunc <- function( cds, cond ) {
   rvf <- rawVarFunc( cds, cond )
   sf <- sizeFactors(cds)[ conditions(cds) == cond ]
   xim <- sum(1/sf) / length(sf)
   function( q ) rvf( q ) + xim * q
}   
@

We also need the gene-wise variance estimates (for the scatter plot part of Fig.\ \ref{varestFig})

<<>>=
diagForA <- varianceFitDiagnostics( cdsFly, "A" )
@

The scatter points in the plots are actual hexagonal bins, produced with the \emph{hexbin}
package. While this looks nice, it unfortunately requires a rather involved call to routines
from the \emph{lattice} to get a proper log-log plot:

<<var_mean_fly,fig=TRUE,  width=4,height=4>>=
# pdf( "var_mean_fly.pdf", width=4, height=4 )

# two functions to make nice exponents for the axis labels

xscale.components.log <- function( ... ) {
   res = xscale.components.default( ... )
   res$bottom$labels$labels = do.call( expression, 
      lapply( res$bottom$labels$at, function(a)
         substitute( 10^b, list(b=a) ) ) )
   res
}

yscale.components.log = function( ... ) {
   res = yscale.components.default( ... )    
   res$left$labels$labels = do.call( expression, 
      lapply( res$left$labels$at, function(a)
         substitute( 10^b, list(b=a) ) ) )
   res
}

xg <- seq( log10(1/3), 
   log10(max( diagForA$baseMean )), length.out=1000 )

print( xyplot(    
   baseVar ~ baseMean, 
   diagForA[ diagForA$baseMean>0,], 
   panel = function( ... ) { 
      panel.hexbinplot( ... )

      # The solid orange line with the fitted base variance
      # (Note that we have to manually switch between natural
      # and logarithmic axis scaling, because hexbin cannot deal
      # with automatic logarithmic scaling):

      llines( xg, log10( baseVarFunc(cdsFly,"A")( 10^xg ) ),
         col="orange", lwd=2 )

      # The dashed orange line with edgeR's variance fit,
      # i.e., variance = mean + alpha * mean^2:

      llines( xg, log10( 10^xg + alpha * (10^xg)^2 ),
         col="orange", lwd=2, lty="dashed" )

      # For each sample with condition "A", plot a purple line
      # with the shot noise, variance = mean, where the mean
      # needs to be scaled by the size factor:

      for( sf in sizeFactors( cdsFly )[ conditions(cdsFly) == "A" ] )
         llines( xg, log10( 10^xg*sf ), col="purple", lwd=1.5 ) },

   trans=function(x) x^(1/6), 
   inv=function(x) x^6,
   xbins=80,
   scales = list( 
      x = list( log=TRUE, axs="i" ),
      y = list( log=TRUE, axs="i", limits=c( 3e-5,3e8 ), tick.number=8 ) ),
   xlab="mean", ylab="variance",
   xscale.components = xscale.components.log,
   yscale.components = yscale.components.log
) )

# dev.off()
@

For Fig.\ \ref{varestFig}b, the code is quite similar:

<<cv2_mean_fly, fig=TRUE,  width=4, height=4>>=
# pdf( "cv2_mean_fly.pdf", width=4, height=4 )

print(xyplot(    
   I( baseVar/baseMean^2 ) ~ baseMean, 
   diagForA[ diagForA$baseMean>0,], 
   panel = function( ... ) { 
      panel.hexbinplot( ... )

      # As we convert from a variance to an SCV, we should adjust for
      # bias, as described in Supplementary Note C. The function
      # adjustScvForBias takes care of this. It requires, as second
      # argument, the information how many samples were used for the
      # variance estimate. The 'attr(...)' expression supplies this,
      # returning in this case simply the value 2:

      llines( xg, 
         adjustScvForBias( 
	    baseVarFunc(cdsFly,"A")( 10^xg ) / (10^xg)^2, 
	    attr( rawVarFunc(cdsFly,"A")(NA), "size" ) ),
         col="orange", lwd=2 )

      # The other lines are as before, only without the log of
      # the y value but with a division by mean^2.

      llines( xg, ( 10^xg + alpha * (10^xg)^2 ) / (10^xg)^2,
         col="orange", lwd=2, lty="dashed" )

      for( sf in sizeFactors( cdsFly )[ conditions(cdsFly) == "A" ] )
         llines( xg, 1 / (10^xg * sf), col="purple", lwd=1.5 ) },

   trans = function(x) x^(1/4), 
   inv = function(x) x^4,
   xbins = 80,
   scales = list( 
      x = list( log=TRUE, axs="i" ),
      y = list( log=FALSE, axs="i", tick.number=8, limits=c(0,.2) ) ),
   xlab = "mean", ylab="squared coefficient of variation",
   xscale.components = xscale.components.log,
))

# dev.off()
@

Just for completeness, the code for Suppl.\ Fig.\ \ref{scvZoomOut}, which is nearly identical to 
the previous code block:

<<cv2_mean_full_fly, fig=TRUE,  width=4, height=4>>=
# pdf( "cv2_mean_full_fly.pdf", width=4, height=4 )

print(xyplot(    
   I( baseVar/baseMean^2 ) ~ baseMean, 
   diagForA[ diagForA$baseMean>0,], 
   panel = function( ... ) { 
      panel.hexbinplot( ... ) 
      llines( xg, 
         adjustScvForBias( 
	    baseVarFunc(cdsFly,"A")( 10^xg ) / (10^xg)^2, 
	    attr( rawVarFunc(cdsFly,"A")(NA), "size" ) ),
         col="orange", lwd=2 )
      llines( xg, ( 10^xg + alpha * (10^xg)^2 ) / (10^xg)^2,
         col="orange", lwd=2, lty="dashed" )
      for( sf in sizeFactors( cdsFly )[ conditions(cdsFly) == "A" ] )
         llines( xg, 1 / (10^xg * sf), col="purple", lwd=1.5 ) },
   trans = function(x) x^(1/4), 
   inv = function(x) x^4,
   xbins = 80,
   scales = list( x = list( log=TRUE, axs="i" ),
      y = list( log=FALSE, axs="i", tick.number=8 ) ),
   xlab = "mean", ylab="squared coefficient of variation",
   xscale.components = xscale.components.log
))

# dev.off()   
@

Finally, for the residuals ECDF plot (Fig.\ \ref{resEcdfFly}), the apropriate function
of DESeq was used:

<<resEcdf_A_fly, fig=TRUE, width=8, height=8>>=
# pdf( "resEcdf_A_fly.pdf", width=8, height=8 )

residualsEcdfPlot( cdsFly, "A" )

# dev.off()
@

\subsection{Figures for DE analysis}

For Fig.\ \ref{DEFig}, we plot one dot for each gene, using semi-transparent colour values
(i.e., specified with alpha channel) in order to make overplotting of dots apparent.
(The \emph{hexbin} package cannot be used as we have two colours, black and red,
to indicate significance.) Extra code, in the form of a panel function, is needed in 
order to draw the small arrows indicating genes with fold changes outside the displayed
value range. We also customise the axis function to get nicer labels.

<<DE_fly, fig=TRUE,  width=4, height=4>>=
# pdf( "DE_fly.pdf", width=4, height=4 )

print(xyplot( 
   log2FoldChange ~ I( baseMean ),
   resFly,  
   pch=16, cex=.3, 
   col=ifelse( resFly$padj < .1, "#FF000040", "#00000040" ),
   panel = function( x, y, col, ... ) {
      above <- (y > 5.8)
      below <- (y < -5.8)
      inside <- !( above | below )
      panel.xyplot( x=x[inside], y=y[inside], col=col[inside], ... ) 
      panel.arrows( x[above], 5.8, x[above], 5.95, col=col[above],
         length=".1", unit="native" )
      panel.arrows( x[below], -5.8, x[below], -5.95, col=col[below],
         length=".1", unit="native" ) },
   axis = function( side, ... ) {
      if( side=="left") {
         panel.axis( side, outside=TRUE, at=seq(-14,14,by=1), labels=FALSE )
         panel.axis( side, outside=TRUE, at=seq(-10,10,by=5), labels=TRUE )
      } 
      if( side=="bottom") {
         panel.axis( side, outside=TRUE, at=seq(-2,10,by=1), rot=0,
            labels = do.call( expression, 
	       lapply( seq(-2,10,by=1), function(a) 
	          substitute( 10^b, list(b=a) ) ) ) )
      } },
   xlab = "mean", ylab = "log2 fold change",
   scales = list( 
      x = list( log=TRUE ), 
      y = list( log=FALSE, limits=c( -6, 6 ) ) ) ))

# dev.off()
@

The volcano plot (Fig.\ \ref{vulcano}) has been produced with:

<<vulcano_fly, fig=TRUE,  width=4, height=4>>=
# pdf( "vulcano_fly.pdf", width=4, height=4 )

print(xyplot( -log10( pval ) ~ log2FoldChange, 
   resFly, 
   pch=20, cex=.2,
   col=ifelse( resFly$padj<.1, "#FF000050", "#00000050" ),
   axis = function( side, ... ) {
      if( side=="bottom") {
         panel.axis( side, outside=TRUE, at=seq(-14,14,by=1), labels=FALSE )
         panel.axis( side, outside=TRUE, at=seq(-10,10,by=5), labels=TRUE )
      }
      if( side=="left") {
         panel.axis( side, outside=TRUE, at=seq(0,25,by=1), labels=FALSE )
         panel.axis( side, outside=TRUE, at=seq(0,25,by=5), 
            labels = do.call( expression, 
	       lapply( seq(0,25,by=5), function(a) 
	          substitute( 10^-b, list(b=a) ) ) ) )
      } },
   xlab = "log2 fold change", ylab = "p value",
   scales = list( 
      x = list( limits=c( -6, 6 ) ),
      y = list( limits=c( 0, 25 ) ) ) ))

# dev.off()
@

\subsection{Hit density}

As edgeR removes rows with zero counts, we should only look at those which are present
in edgeR's result table:

<<>>=
goodRows <- resFly$id %in% rownames(edgerResFlyDT$table)
@

We calculate the densities of hits as follows:

<<>>=
densAll   <- density( log10(resFly$baseMean)[ goodRows ] )
densHits  <- density( log10(resFly$baseMean)[ goodRows & resFly$padj < .1 ] )
densedgDT <- density( log10(resFly$baseMean)[ goodRows ][ edgerResFlyDTPadj < .1 ] )
densedgDC <- density( log10(resFly$baseMean)[ goodRows ][ edgerResFlyDCPadj < .1 ] )
densedgNT <- density( log10(resFly$baseMean)[ goodRows ][ edgerResFlyNTPadj < .1 ] )
densedgNC <- density( log10(resFly$baseMean)[ goodRows ][ edgerResFlyNCPadj < .1 ] )
@

We plot these for densities as follows:

<<hits_dens_fly_all, fig=TRUE,  width=6, height=4>>=
# pdf( "hits_dens_fly_all.pdf", width=6, height=4 )

plot( densAll$x, densAll$y / 7, type="l", ylim=c(0,.05),
   xlab="log10 mean", ylab="density", xaxs="i", yaxs="i", col="gray" )
lines( densHits$x, densHits$y * densHits$n / densAll$n, 
   col="red", lwd=2 )
lines( densedgDT$x, densedgDT$y * densedgDT$n / densAll$n, 
   col="blue", lwd=2, lty="dashed" )
lines( densedgDC$x, densedgDC$y * densedgDC$n / densAll$n, 
   col="lightblue", lwd=2, lty="dashed" )
lines( densedgNT$x, densedgNT$y * densedgNT$n / densAll$n, 
   col="blue", lwd=2, lty="solid" )
lines( densedgNC$x, densedgNC$y * densedgNC$n / densAll$n, 
   col="lightblue", lwd=2, lty="solid" )
text( densAll$x[310]+.2, densAll$y[310]/7, "x7", col="gray" )

# dev.off()
@

For this data set, there is hardly a noticeably difference between the two ways of 
specifying library size. Hence, we omitted the dashed lines from the figure in the paper:

<<hits_dens_fly, fig=TRUE,  width=6, height=4>>=
# pdf( "hits_dens_fly.pdf", width=6, height=4 )

plot( densAll$x, densAll$y / 7, type="l", ylim=c(0,.05),
   xlab="log10 mean", ylab="density", xaxs="i", yaxs="i", col="gray" )
lines( densHits$x, densHits$y * densHits$n / densAll$n, 
   col="red", lwd=2 )
lines( densedgNT$x, densedgNT$y * densedgNT$n / densAll$n, 
   col="blue", lwd=2, lty="solid" )
lines( densedgNC$x, densedgNC$y * densedgNC$n / densAll$n, 
   col="lightblue", lwd=2, lty="solid" )
text( densAll$x[310]+.2, densAll$y[310]/7, "x7", col="gray" )

# dev.off()
@

The effect of the tag-wise dispersion estimate (i.e., of moderated testing) is also hardly
appreciable. This is fully expected because of the low number of replicates.
\medskip

The paper gives numbers of hits, obtained as follows:

Percentage of genes with DESeq $p$ value below 5\%:

<<>>=
length(which( resFly$pval < .05 )) / nrow( resFly )
@

Number of hits at 10\% FDR, found by DESeq:

<<>>=
length( which( resFly$padj < .1 ) )
@

The number of genes judged significant by edgeR is a bit higher and not very sensitive
to the choice of library size specification and dispersion mode:

<<>>=
c( DT = length( which( edgerResFlyDTPadj < .1 ) ),
   DC = length( which( edgerResFlyDCPadj < .1 ) ),
   NT = length( which( edgerResFlyNTPadj < .1 ) ),
   NC = length( which( edgerResFlyNCPadj < .1 ) ) )
@

The numbers of hits shared between DESeq and edgeR are as follows:

<<>>=
list( 
  table( DESeq = resFly$padj[goodRows] < .1, edgeR_DT = edgerResFlyDTPadj < .1 ),
  table( DESeq = resFly$padj[goodRows] < .1, edgeR_DC = edgerResFlyDCPadj < .1 ),
  table( DESeq = resFly$padj[goodRows] < .1, edgeR_NT = edgerResFlyNTPadj < .1 ),
  table( DESeq = resFly$padj[goodRows] < .1, edgeR_NC = edgerResFlyDCPadj < .1 ) )
@

\subsection{Same-vs-same comparison}

For the same-vs-same comparison of the two ``A'' samples, we first remove those
genes which have zero counts in both samples and make a new CountDataSet, then run
the standard analysis:\footnote{With a DESeq version prior to 1.1.6, use 
``\texttt{pool=TRUE}'' rather than ``\texttt{method="blind"}''}

<<>>=
a <- counts(cdsFly)[,conditions(cdsFly)=="A"]
cdsA <- newCountDataSet( a[rowSums(a)>0,], c( "A1", "A2" ) )
cdsA <- estimateSizeFactors( cdsA )
cdsA <- estimateVarianceFunctions( cdsA, method="blind" )
resA <- nbinomTest( cdsA, "A1", "A2" )
@

We do the same with edgeR. (As edgeR does not have an option 
for ``blind'' variance estimation as DESeq does, we tell it first 
that both samples are replicates, then estimate the dispersion,
and then change the group labels before the actual test:

<<>>=
dglA <- DGEList( counts(cdsA), sizeFactors( cdsA ) * 1e7, 
   c( "A", "A" ) )
dglA <- estimateCommonDisp( dglA )
dglA <- estimateTagwiseDisp( dglA )
dglA$samples$group = c( "A1", "A2" )
eResA <- exactTest( dglA, common.disp=FALSE  ) 
@

Finally, we assume a Poisson model and fit a GLM:

<<>>=
pvalPois <- 
   apply( counts(cdsA), 1, function(cl) 
      anova( 
         glm( 
	    counts ~ sample, 
	    data.frame( counts=cl, sample=colnames(counts(cdsA)) ), 
            family=poisson ), 
	 test="Chisq" )$`P(>|Chi|)`[2] )
@

This code generates the ECDF curves (Fig.\ \ref{repl_pval}a):

<<repl_pval_fly_A_ecdf, fig=TRUE,  width=6, height=6>>=
# pdf( file="repl_pval_fly_A_ecdf.pdf", width=6, height=6)

print(ecdfplot( ~ data | which, 
   make.groups( 
      `DESeq, below 100` = resA$pval[ resA$baseMean < 100 ],
      `DESeq, above 100` = resA$pval[ resA$baseMean > 100 ],   
      `DESeq, all` = resA$pval,
      `edgeR, below 100` = eResA$table$p.value[ 
          resA$baseMean < 100 ],
      `edgeR, above 100` = eResA$table$p.value[ 
          resA$baseMean > 100 ],
      `edgeR, all` = eResA$table$p.value,
      `Poisson, below 100` = pvalPois[ resA$baseMean < 100 ],
      `Poisson, above 100` = pvalPois[ resA$baseMean > 100 ],	
      `Poisson, all` = pvalPois ),
   layout=c(3,3), as.table=TRUE, xlab = "p value",
   panel = function( ...) {
      lsegments( 0, 0, 1, 1, col="darkgray" )
      panel.ecdfplot( ... ) },
   scales = list( tick.number = 4, cex=1 ) ))

# dev.off()
@

The zoom-in (Fig.\ \ref{repl_pval}b):

<<repl_pval_fly_A_ecdf_zoom, fig=TRUE,  width=6, height=6>>=
# pdf( file="repl_pval_fly_A_ecdf_zoom.pdf", width=6, height=6)

print(ecdfplot( ~ data | which, 
   make.groups( 
      `DESeq, below 100` = resA$pval[ resA$baseMean < 100 ],
      `DESeq, above 100` = resA$pval[ resA$baseMean > 100 ],   
      `DESeq, all` = resA$pval,
      `edgeR, below 100` = eResA$table$p.value[ 
          resA$baseMean < 100 ],
      `edgeR, above 100` = eResA$table$p.value[ 
          resA$baseMean > 100 ],
      `edgeR, all` = eResA$table$p.value,
      `Poisson, below 100` = pvalPois[ resA$baseMean < 100 ],
      `Poisson, above 100` = pvalPois[ resA$baseMean > 100 ],	
      `Poisson, all` = pvalPois ),
   layout=c(3,3), as.table=TRUE, xlab = "p value",
   panel = function( ...) {
      lsegments( 0, 0, 1, 1, col="darkgray" )
      panel.ecdfplot( ... ) },
   xlim=c( -.005, .1 ), ylim=c( -.005, .1 ),
   scales = list( tick.number = 4, cex=1 ) ))

# dev.off()
@

\section{Analysis of neural stem cell data}

We performed the same analysis as described in the previous section for
the neural stem cell set. For completeness, we give here the full code to
produce all the figures relating to this data set, but with only few comments,
as the code is very similar to the already described code.

\subsection{Analysis with DESeq and edgeR}

As stated above and in the paper, we omit the first sample. Apart from that, 
the analysis is performed as before:

<<>>=
cdsNS <- newCountDataSet( countsTableNS[,-1], condsNS[-1] )
cdsNS <- estimateSizeFactors( cdsNS )
cdsNS <- estimateVarianceFunctions( cdsNS )
resNS <- nbinomTest( cdsNS, "GNS", "NS" )
@

For edgeR, we also have to explicitly remove the ``GNSL'' column. Apart from that, as before:

<<>>=
colGNSL <- which( conditions(cdsNS) == "GNSL" )

dglNS.D <- DGEList( 
   counts = counts(cdsNS)[,-colGNSL], 
   group = factor( conditions(cdsNS)[-colGNSL] ), 
   lib.size = 1e7*sizeFactors(cdsNS)[-colGNSL] )
dglNS.D <- estimateCommonDisp( dglNS.D )
dglNS.D <- estimateTagwiseDisp( dglNS.D )

edgerResNS.DT <- exactTest( dglNS.D, common.disp=FALSE )
edgerResNS.DTPadj <- p.adjust( edgerResNS.DT$table$p.value, method="BH" )

edgerResNS.DC <- exactTest( dglNS.D, common.disp=TRUE )
edgerResNS.DCPadj <- p.adjust( edgerResNS.DC$table$p.value, method="BH" )

libSizesNS <- c( G144=7604834, G166=13625570, G179=12291910,
   CB541=12872125, CB660=10502656 )

dglNS.N <- DGEList( 
   counts = counts(cdsNS)[,-colGNSL], 
   group = factor( conditions(cdsNS)[-colGNSL] ), 
   lib.size = libSizesNS[-colGNSL] )
dglNS.N <- estimateCommonDisp( dglNS.N )
dglNS.N <- estimateTagwiseDisp( dglNS.N )

edgerResNS.NT <- exactTest( dglNS.N, common.disp=FALSE )
edgerResNS.NTPadj <- p.adjust( edgerResNS.NT$table$p.value, method="BH" )

edgerResNS.NC <- exactTest( dglNS.N, common.disp=TRUE )
edgerResNS.NCPadj <- p.adjust( edgerResNS.NC$table$p.value, method="BH" )
@

DESeq finds this number of significant hits at 10\% FDR:

<<>>=
length( which( resNS$padj < .1 ) )
@

For edgeR, the numbers are:

<<>>=
c( DT = length( which( edgerResNS.DTPadj < .1 ) ),
   DC = length( which( edgerResNS.DCPadj < .1 ) ),
   NT = length( which( edgerResNS.NTPadj < .1 ) ),
   NC = length( which( edgerResNS.NCPadj < .1 ) ) )
@

\subsection{Figures}

Figure \ref{cv2_Paul} has been made with this code:

<<cv2_mean, fig=TRUE,  width=4, height=4>>=
# pdf( "cv2_mean.pdf", width=4, height=4 )

diagForGNS <- varianceFitDiagnostics( cdsNS, "GNS" )
alpha <- dglNS.D$common.dispersion

print(xyplot(    
   I( baseVar/baseMean^2 ) ~ baseMean, 
   diagForGNS[ diagForGNS$baseMean>0,], 
   panel = function( ... ) { 
      panel.hexbinplot( ... ) 
      llines( 
         xg, 
         adjustScvForBias( 
           baseVarFunc(cdsNS,"GNS")( 10^xg ) / (10^xg)^2, 
           attr( rawVarFunc(cdsNS,"GNS")(NA), "size" ) ),
         col="orange", lwd=2 )
      llines( xg, ( 10^xg + alpha * (10^xg)^2 ) / (10^xg)^2,
         col="orange", lwd=2, lty="dashed" )
      for( sf in sizeFactors( cdsNS )[ conditions(cdsNS) == "GNS" ] )
         llines( xg, 1 / (10^xg * sf), col="purple", lwd=1.5 ) },
   trans = function(x) x^(1/4), 
   inv = function(x) x^4,
   xbins = 80,
   scales = list( x = list( log=TRUE, axs="i" ),
      y = list( log=FALSE, axs="i", tick.number=8 ) ),
   xlab = "mean", ylab="squared coefficient of variation"
))

# dev.off()
@

For Fig.\ \ref{DE_Paul}, this code was used:

<<DE,fig=TRUE,  width=4, height=4>>=
# pdf( "DE.pdf", width=4, height=4 )

print(xyplot( 
   log2FoldChange ~ I( baseMean ),
   resNS, 
   pch=16, cex=.3, 
   col=ifelse( resNS$padj < .1, "#FF000040", "#00000040" ),
   axis = function( side, ... ) {
      if( side=="left") {
         panel.axis( side, outside=TRUE, at=seq(-14,14,by=1), labels=FALSE )
         panel.axis( side, outside=TRUE, at=seq(-10,10,by=5), labels=TRUE )
      } 
      if( side=="bottom") {
         panel.axis( side, outside=TRUE, at=seq(-2,10,by=1), 
            labels=paste( "10^", seq(-2,10,by=1), sep="" ) )
      } },
   xlab = "mean", ylab = "log2 fold change",
   scales = list( x=list(log=TRUE), y=list(log=FALSE) ) ))

# dev.off()
@

Figure \ref{hits_dens_Paul}, with the density estimates, was done with:

<<hits_dens,fig=TRUE,  width=6, height=4 >>=
# pdf( "hits_dens.pdf", width=6, height=4 )

goodRowsNS <- resNS$id %in% rownames(edgerResNS.DT$table)

densAll.NS   <- density( log10(resNS$baseMean)[ goodRowsNS ] )
densHits.NS  <- density( log10(resNS$baseMean)[ goodRowsNS & resNS$padj < .1 ] )
densedgDT.NS <- density( log10(resNS$baseMean)[ goodRowsNS ][ edgerResNS.DTPadj < .1 ] )
densedgDC.NS <- density( log10(resNS$baseMean)[ goodRowsNS ][ edgerResNS.DCPadj < .1 ] )
densedgNT.NS <- density( log10(resNS$baseMean)[ goodRowsNS ][ edgerResNS.NTPadj < .1 ] )
densedgNC.NS <- density( log10(resNS$baseMean)[ goodRowsNS ][ edgerResNS.NCPadj < .1 ] )

plot( densAll.NS$x, densAll.NS$y / 7, type="l", ylim=c(0,.07),
   xlab="log10 mean", ylab="density", xaxs="i", yaxs="i", 
   col="gray" )
lines( densHits.NS$x, densHits.NS$y * densHits.NS$n / densAll.NS$n, 
   col="red", lwd=2 )
lines( densedgDT.NS$x, densedgDT.NS$y * densedgDT.NS$n / densAll.NS$n, 
   col="blue", lwd=2, lty="dashed" )
lines( densedgDC.NS$x, densedgDC.NS$y * densedgDC.NS$n / densAll.NS$n, 
   col="lightblue", lwd=2, lty="dashed" )
lines( densedgNT.NS$x, densedgNT.NS$y * densedgNT.NS$n / densAll.NS$n, 
   col="blue", lwd=2, lty="solid" )
lines( densedgNC.NS$x, densedgNC.NS$y * densedgNC.NS$n / densAll.NS$n, 
   col="lightblue", lwd=2, lty="solid" )
text( densAll.NS$x[310]+.2, densAll.NS$y[310]/7, "x7", col="gray" )

# dev.off()
@

Finally, Figure \ref{repl_pval_Paul} with the type-I error control histograms

<<repl_pval_A_ecdf, fig=TRUE,  width=8, height=6 >>=
# pdf( file="repl_pval_A_ecdf.pdf", width=6, height=4.5)

cdsGNS <- newCountDataSet( counts(cdsNS)[,1:2], c( "GNS_A", "GNS_B" ) )
cdsGNS <- estimateSizeFactors( cdsGNS )
cdsGNS <- estimateVarianceFunctions( cdsGNS, method="blind" )
resGNS <- nbinomTest( cdsGNS, "GNS_A", "GNS_B" )
medGNS <- median( getBaseMeansAndVariances( counts(cdsGNS), 
             sizeFactors(cdsGNS) )$baseMean )

dglGNS <- DGEList( counts(cdsGNS), sizeFactors( cdsGNS ) * 1e7, 
   c( "GNS", "GNS" ) )
dglGNS <- estimateCommonDisp( dglGNS )
dglGNS <- estimateTagwiseDisp( dglGNS )
dglGNS$samples$group = c( "GNS_1", "GNS_2" )
eResGNS <- exactTest( dglGNS, common.disp=FALSE ) 
eMedGNS <- median( rowMeans(dglGNS$pseudo.alt) )

print(ecdfplot( ~ data | which, 
   make.groups( 
      `DESeq lower half` = resGNS$pval[ resGNS$baseMean < medGNS ],
      `DESeq upper half` = resGNS$pval[ resGNS$baseMean > medGNS ],   
      `DESeq all` = resGNS$pval,
      `edgeR lower half` = eResGNS$table$p.value[ 
          rowMeans(dglGNS$pseudo.alt) < eMedGNS ],
      `edgeR upper half` = eResGNS$table$p.value[ 
          rowMeans(dglGNS$pseudo.alt) > eMedGNS ],
      `edgeR all` = eResGNS$table$p.value ),
   layout=c(3,2), xlab = "p value", as.table=TRUE, 
   panel = function( ...) {
      lsegments( 0, 0, 1, 1, col="darkgray" )
      panel.ecdfplot( ... ) },
   scales = list( tick.number=4, cex=1 ) ))

# dev.off()   
@

\section{Working without replicates}

To see how much hits we lose without replicates, we reduce the fly data set
to only one sample per condition

<<>>=
cdsFly2 <- cdsFly[ ,c( "A1", "B1" ) ]
cdsFly2 <- estimateVarianceFunctions( cdsFly2, method="blind" )
resFly2 <- nbinomTest( cdsFly2, "A", "B" )
@

Here we see a comparison of the hits with the full data set vs the reduced data set:

<<>>=
addmargins( table( full = resFly$padj < .1, reduced = resFly2$padj < .1 ) )
@

The same for the neural stem cell data set:

<<>>=
cdsNS2 <- cdsNS[ ,c( "G144", "CB541" ) ]
cdsNS2 <- estimateVarianceFunctions( cdsNS2, method="blind" )
resNS2 <- nbinomTest( cdsNS2, "GNS", "NS" )
addmargins( table( full = resNS$padj < .1, reduced = resNS2$padj < .1 ) )
@

\section{VST}

To demonstrate the VST, we first make a new CountDataSet from the neural cell data,
this time with all samples. To make clear that the dendrogram we get in the end was made
without using class labels, we do not pass condition labels.

<<heatmap,fig=TRUE,width=6,height=6>>=
# pdf( "heatmap.pdf" )

cdsNS.all <- newCountDataSet( countsTableNS, rep( "dummy", ncol(countsTableNS) ) )
cdsNS.all <- estimateSizeFactors( cdsNS.all )

vsdNS <- getVarianceStabilizedData( cdsNS.all )
distMatrix <- as.matrix(dist(t( vsdNS )))
rownames( distMatrix ) <- c( "GNS (*)", "GNS (*)", "GNS", "GNS (L)", "NS", "NS" )

heatmap.2( distMatrix, density.info="none", trace="none", revC=TRUE,
   col=colorRampPalette(c("darkblue","yellow"))(100) )

# dev.off()
@

\section{Yeast RNA-Seq data}

For the yeast data, we have replicates, but no conditions to compare. DESeq's CountDataSet
class is not designed to deal with this situation, and we fall back to the set of DESeq
functions provided to directly work with matrix data:

<<>>=
sfYeast <- estimateSizeFactorsForMatrix( countsYeast )

# A function to get a base variance function:
get_bvf <- function( cl ) {
   rvf <- estimateVarianceFunctionForMatrix( countsYeast[,cl], sfYeast[cl] )
   xim <- sum(1/sfYeast[cl])/length(cl)
   function( q ) rvf( q ) + xim * q
}

# The variance estimated between the replicates:
bvf_dT_tech <- get_bvf( c("dT_ori","dT_tech") )
bvf_dT_bio  <- get_bvf( c("dT_ori","dT_bio" ) )
bvf_RH_tech <- get_bvf( c("RH_ori","RH_tech") )
bvf_RH_bio  <- get_bvf( c("RH_ori","RH_bio" ) )

# The common-scale means
meansYeast <- getBaseMeansAndVariances( countsYeast, sfYeast )$baseMean
@

The plot of Fig. \ref{noiseNaga} is produced with

<<naga, fig=TRUE,  width=4, height=4>>=
#pdf("naga.pdf", width=4, height=4)

layout( cbind(1:2), heights = c( 1, 3 ) )
par( cex=.75 )

par(mai=c(0.2, 1, 0.5, 0.5))
plot( NULL, main="", xaxs="i", yaxs="i", ylab="density", 
   xlim=c( -.3, 4 ), ylim=c( 0, 1200 ), xaxt="n" )
hist( log10( meansYeast ), breaks=30, add=TRUE )

par(mai=c(1, 1, 0.01, 0.5))
plot( NULL, log="x", ylim=c(0, .65), xaxs="i", yaxs="i",
   xlab="mean", ylab="squared coefficient of variation",
   xlim=c( 10^-.3, 10^4 ) )
xg <- 10^seq( 0, 4, length.out=1000 ) 
rect( xg[-1000], 1/(xg[-1000]*max(sfYeast)), 
   xg[-1], 1/(xg[-1000]*min(sfYeast)), col="#FFD0FF", lty="blank" )
lines( xg, bvf_dT_bio(xg)/xg^2, col="brown" )
lines( xg, bvf_dT_tech(xg)/xg^2, col="darkblue" )
lines( xg, bvf_RH_bio(xg)/xg^2, col="brown", lty="dashed" )
lines( xg, bvf_RH_tech(xg)/xg^2, col="darkblue", lty="dashed" )
lines( xg, 1/xg, col="#FFD0FF" )

# dev.off()
@

\section{ChIP-Seq}

We first look at a heat map for the Pol-II count data:

<<PolII_heatmap, fig=TRUE>>=
cdsPolII.1 <- newCountDataSet( countsPolII, rep( "dummy", 12 ) )
cdsPolII.1 <- estimateSizeFactors( cdsPolII.1 )
heatmap( as.matrix( dist( t( getVarianceStabilizedData( cdsPolII.1 ) ) ) ),
   cexCol=.7, cexRow=.7, symm=TRUE,
   col=colorRampPalette(c("darkblue","yellow"))(100) )
@

The heatmap shows clear evidence of batch effects for GM12878: Replicates 1 to 4
form one batch and 6 and 7 another one. For GM12891, replicates 1 to 3 form
one batch and 4 to 6 another one.

For Fig.\ \ref{figJan}, we have to chose two replicates for each of the groups A1, A2,
B1, and B2. In order to be balanced with respect to batches, we chose the
replicates such that each batch is present in each group, as follows 
(Unused samples get denoted with an ``x''):

<<>>=
condsPolII <- c( 
   GM12878_1 = "A1", GM12878_2 = "A2", GM12878_3 = "x", 
   GM12878_4 = "x",  GM12878_6 = "A1", GM12878_7 = "A2",
   GM12891_1 = "B1", GM12891_2 = "B2", GM12891_3 = "x",
   GM12891_4 = "B1", GM12891_5 = "B2", GM12891_6 = "x" )

cdsPolII.2 <- newCountDataSet( countsPolII, condsPolII )
cdsPolII.2 <- estimateSizeFactors( cdsPolII.2 )
cdsPolII.2 <- estimateVarianceFunctions( cdsPolII.2 )
resPolII.A1vsA2 <- nbinomTest( cdsPolII.2, "A1", "A2" )
resPolII.B1vsB2 <- nbinomTest( cdsPolII.2, "B1", "B2" )
resPolII.A1vsB1 <- nbinomTest( cdsPolII.2, "A1", "B1" )
@

For the comparison of 4 replicates against 4, we use only the
first letter:

<<>>=
cdsPolII.3 <- newCountDataSet( countsPolII, substr( condsPolII, 1, 1 ) )
conditions( cdsPolII.3 )
@

<<>>=
cdsPolII.3 <- estimateSizeFactors( cdsPolII.3 )
cdsPolII.3 <- estimateVarianceFunctions( cdsPolII.3 )
resPolII.AvsB <- nbinomTest( cdsPolII.3, "A", "B" )
@

Number of hits at 10\% FDR:

<<>>=
table( resPolII.A1vsA2$padj < .1 )
table( resPolII.B1vsB2$padj < .1 )
table( resPolII.A1vsB1$padj < .1 )
table( resPolII.AvsB$padj < .1 )

@

The same with a Poisson GLM:

<<>>=
do_glm <- function( cds, cols )
   apply( counts(cds)[, cols], 1, function(r)
      anova( 
         glm( r ~ conditions(cds)[cols], family=poisson ), 
         test="Chisq" )$`P(>|Chi|)`[2] )

pvalPois.A1vsA2 <- do_glm( cdsPolII.2, which( conditions(cdsPolII.2) %in% c("A1","A2") ) )
pvalPois.B1vsB2 <- do_glm( cdsPolII.2, which( conditions(cdsPolII.2) %in% c("B1","B2") ) )
pvalPois.A1vsB1 <- do_glm( cdsPolII.2, which( conditions(cdsPolII.2) %in% c("A1","B1") ) )
pvalPois.AvsB   <- do_glm( cdsPolII.2, which( conditions(cdsPolII.3) %in% c("A","B") ) )
@

The ECDF plot of Fig.\ \ref{figJan} is now made with:

<<ecdf_Jan_pois, fig=TRUE,  width = 6, height = 4.5>>=

# pdf( "ecdf_Jan_pois.pdf", width = 6, height = 4.5 )
print(ecdfplot( ~ data | which, 
   make.groups( 
      `D: A1 vs A2` = resPolII.A1vsA2$pval,
      `D: B1 vs B2` = resPolII.B1vsB2$pval,
      `D: A1 vs B1` = resPolII.A1vsB1$pval,
      `D: A vs B`   = resPolII.AvsB$pval,
      `P: A1 vs A2` = pvalPois.A1vsA2,
      `P: B1 vs B2` = pvalPois.B1vsB2,
      `P: A1 vs B1` = pvalPois.A1vsB1,
      `P: A vs B`   = pvalPois.AvsB ),
   layout=c(4,2), xlab = "p value", as.table=TRUE,  
   panel = function( ...) {
      lsegments( 0, 0, 1, 1, col="darkgray" )
      panel.ecdfplot( ... ) },
   scales = list( tick.number=4, cex=1 ) ))

# dev.off()      
@

\section{Simulation}

The details of simulations in Supplementary Note G are given here.

\subsection{Constant dispersion}

The count data was drawn as follows

<<>>=

# Number of genes to simulate
ngenes <- 20000

# True mean expression value across conditions drawn from an
# exponential distribution
true_q0middle <- rexp( ngenes, 1/250 )

# These values might be typical size factors as could be found
# in a real experiment
sf <- c( .5, 1.7, 1.4, .9 )

# Two replicates for each condition:
conds <- c( "A", "A", "B", "B" )

# Mark around 30% of the genes as differentially expressed:
true_isDE <- runif( ngenes ) < .3

# For the DE genes, draw a log fold change from a normal
true_lfc <- rnorm( ngenes, sd=.7 ) * true_isDE

# Go up and down from the true_q0middle according to these log
# fold changes to get the values q for the conditions
true_q0 <- data.frame(
   A = true_q0middle * 2^(  true_lfc/2 ),
   B = true_q0middle * 2^( -true_lfc/2 ) )

# For the variance-mean relation, let's assume that the shape 
# assumed by edgeR is correct, i.e., variance = mean + alpha * mean^2
# and use this alpha:   
alpha <- .015

# Parametrise the NB to mean and variance:
rnbinomMV <- function( n, mu, v ) 
  rnbinom( n, prob = mu/v, size = mu^2/(v-mu) )

# Now draw counts
countsTable <- 
   sapply( 1:4, function(j)
      sapply( true_q0[[ conds[j] ]], function(q)
         rnbinomMV( 1, sf[j]*q, sf[j]*q + sf[j]^2 * q^2 * alpha ) ) )

# The good rows:
nonzero <- rowSums(countsTable)>0 
@

Now, we analyse this count table:

<<>>=
# A standard analysis with DESeq:
cds <- newCountDataSet( countsTable, conds )
cds <- estimateSizeFactors( cds )
cds <- estimateVarianceFunctions( cds )
res <- nbinomTest( cds, "A", "B" )

# The same with edgeR. (edgeR gets the advantage of knowing the true
# library sizes)
dgl <- DGEList( counts=countsTable, group=conds, lib.size=sf*1e7 )
dgl <- estimateCommonDisp(dgl)
edgerRes <- exactTest( dgl )
edgerResPadj <- p.adjust( edgerRes$table$p.value, method="BH" )
@

Compare the results.

Type-I error control: Which percentage of genes without true DE have a p value below 5\%?

For DESeq:
<<>>=
length(which( res$pval[nonzero]<.05 & !true_isDE[nonzero] )) / length(nonzero)
@

For edgeR:
<<>>=
length(which( edgerRes$table$p.value<.05 & !true_isDE[nonzero] )) / length(nonzero)
@

Detection power: Which percentage of the true positives is detected at 10\% FDR?

For DESeq:
<<>>=
length(which( res$padj[nonzero]<.1 & true_isDE[nonzero] )) / 
   length(which( true_isDE[nonzero] ))
@

For edgeR:
<<>>=
length(which( edgerResPadj<.1 & true_isDE[nonzero] )) / 
   length(which( true_isDE[nonzero] ))
@

FDR control:

For DESeq:
<<>>=
length(which( res$padj[nonzero]<.1 & !true_isDE[nonzero] )) / 
   length(which( res$padj[nonzero]<.1 ))
@

For edgeR:
<<>>=
length(which( edgerResPadj<.1 & !true_isDE[nonzero] )) / 
   length(which( edgerResPadj<.1 ))
@

For Fig.\ \ref{pvalcomp}a, we used this code:

<<pval_nonDE_B, fig=TRUE, width=4, height=3>>=
# pdf( "pval_DE_B.pdf", width=4, height=4 )

print(ecdfplot( ~ data | which, 
   make.groups( 
      `DESeq` = res$pval[nonzero][ !true_isDE[nonzero] ],
      `edgeR` = edgerRes$table$p.value[ !true_isDE[nonzero] ] ),
   layout=c(2,1), as.table=TRUE, xlab = "p value",
   panel = function( ...) {
      lsegments( 0, 0, 1, 1, col="darkgray" )
      panel.ecdfplot( ... ) },
   xlim=c( -.005, .1 ), ylim=c( -.005, .1 ),
   scales = list( tick.number = 4, cex=1 ) ))

# dev.off()   
@

and for Fig.\ \ref{pvalcomp}b

<<pval_DE, fig=TRUE, width=4, height=4>>=
# pdf( "pval_DE.pdf", width=4, height=4 )

print(xyplot(  
   res$pval[true_isDE & nonzero] ~ edgerRes$table$p.value[true_isDE[nonzero]],
   panel = function( ... ) {
      panel.hexbinplot( ... )
      llines( c(-35, 0), c(-35, 0), col="#30FF5090" ) },
   trans = function(x) x^(1/4), 
   inv = function(x) x^4,
   xbins = 150,
   scales = list(
      x = list(log=TRUE),
      y = list(log=TRUE), apect=1 ),
   xlim = c( 1e-15, 10 ),
   ylim = c( 1e-15, 10 ),
   aspect = 1,
   xlab="p value from edgeR", ylab="p value from DESeq", asp=1
))

# dev.off()
@

\subsection{Non-constant dispersion}

We prepare data as before:

<<>>=
ngenes <- 20000
true_q0middle <- rexp( ngenes, 1/250 )
sf <- c( .5, 1.7, 1.4, .9 )
conds <- c( "A", "A", "B", "B" )
true_isDE <- runif( ngenes ) < .3
true_lfc <- rnorm( ngenes, sd=.7 ) * true_isDE
true_q0 <- data.frame(
   A = true_q0middle * 2^(  true_lfc/2 ),
   B = true_q0middle * 2^( -true_lfc/2 ) )
@

Now, we define a mean-dependent raw SCV value and use it in
drawing the count data.

<<>>=
alpha <- function(mu) .01 + 9/(mu+100)

countsTable <- 
   sapply( 1:4, function(j)
      sapply( true_q0[[ conds[j] ]], function(q)
         rnbinomMV( 1, sf[j]*q, sf[j]*q + sf[j]^2 * q^2 * alpha( q ) ) ) )
@

Then, continue as before.
<<>>=
nonzero <- rowSums(countsTable)>0

cds <- newCountDataSet( countsTable, conds )
cds <- estimateSizeFactors( cds )
cds <- estimateVarianceFunctions( cds )
res <- nbinomTest( cds, "A", "B" )

dgl <- DGEList( counts=countsTable, group=conds, lib.size=sf*1e7 )
dgl <- estimateCommonDisp(dgl)
dgl <- estimateTagwiseDisp(dgl)
edgerRes <- exactTest( dgl, common.disp=FALSE )
edgerResPadj <- p.adjust( edgerRes$table$p.value, method="BH" )
@

Here is a plot showing the simulated (purple) and the 
fitted (solid red and green) raw SCV curves:

<<scvSim2,fig=TRUE>>=
scvPlot( cds )
xg <- 10^seq( -2, 4, length.out=1000 )
lines( xg, alpha( xg ), col="purple" )
@

Finally, the ECDF plot
<<pvalcompsim2, fig=TRUE,  width=8, height=6 >>=
# pdf( file="pvalcompsim2", width=8, height=6 )

baseMeans <- t(t(countsTable)/sf)
lowerHalf <- baseMeans < median( baseMeans )

print(ecdfplot( ~ data | which, 
   make.groups( 
      `DESeq, below median` = res$pval[nonzero][ 
         !true_isDE[nonzero] & lowerHalf[nonzero] ],
      `DESeq, above median` = res$pval[nonzero][ 
         !true_isDE[nonzero] & !lowerHalf[nonzero] ],   
      `DESeq, all`          = res$pval[nonzero][ 
         !true_isDE[nonzero] ],
      `edgeR, below median` = edgerRes$table$p.value[ 
         !true_isDE[nonzero] & lowerHalf[nonzero] ],
      `edgeR, above median` = edgerRes$table$p.value[ 
         !true_isDE[nonzero] & !lowerHalf[nonzero] ],
      `edgeR, all` = edgerRes$table$p.value[ 
         !true_isDE[nonzero] ] ),
   layout=c(3,2), as.table=TRUE, xlab = "p value",
   panel = function( ...) {
      lsegments( 0, 0, 1, 1, col="darkgray" )
      panel.ecdfplot( ... ) },
   xlim=c( -.005, .1 ), ylim=c( -.005, .1 ),
   scales = list( tick.number = 4, cex=1 ) ))

# dev.off()   
@
\section{Other figures}

\subsection{Bias-adjustment function}

The code for Fig.\ \ref{biasAdjust} is:
<<biasAdjust, fig=TRUE, width=4, height=4>>=
# pdf("biasAdjust.pdf", width=4, height=4)

plot( NULL, xlim=c(0,2), ylim=c(0,2), 
   xlab = expression( ( hat(sigma)^2 - hat(mu) ) / hat(mu)^2 ), 
   ylab = expression( gamma ) )
xg <- seq( 0, 2, length.out=1000 )
for( i in 2:15 ) 
   lines( xg, adjustScvForBias( xg, i ), 
      col=colorRampPalette(c("red","black"))(15)[i] )

# dev.off()      
@

See the code of the functions \emph{DESeq::adjustScvForBias} and 
\emph{DESeq:::prepareScvBiasCorrectionFits} for details on
how the variance adjustment is performed by means of fitting a
LOESS regression through results from numerical simulations.

\section{Figure of $p(a,b)$}

The code for Fig.\ \ref{figp1}:

<<p1, fig=TRUE,  width=4, height=4>>=
# pdf("p1.pdf", width=4, height=4)

dnbinomMV <- function( x, mu, v ) 
  dnbinom( x, prob = mu/v, size = mu^2/(v-mu) )

plot( 0:10000, 
   sapply( 0:10000, function(k) 
      dnbinomMV( k, 7000, 7000+.1*7000^2 ) * 
         dnbinomMV( 10000-k, 4000, 4000+.1*4000^2 ) ), 
   type='l', xlab="a", ylab="p(a,b)" )
abline( v = 7000 / (7000+4000) * 10000 )

# dev.off()
@

Finally, save all data:

<<>>=
save.image( "suppl_ii.RData" )
@

\end{document}