### 2012-01-05 package/smutans/inst/doc/smutans.Rnw
smutans package is being developed.

%\VignetteIndexEntry{Analyzing Streptococcus mutans RNA-Seq data}
%\VignettePackage{smutans}

% To compile this document
% library('weaver'); rm(list=ls()); Sweave('smutans.Rnw', driver=weaver()); system('pdflatex smutans')

\documentclass{article}

\usepackage{Sweave}
\usepackage[letterpaper]{geometry}
\usepackage{hyperref,graphicx}
\usepackage{color}

\SweaveOpts{keep.source=TRUE,eps=FALSE,include=FALSE,width=4,height=4.5}
\newcommand{\Robject}[1]{\texttt{#1}}
\newcommand{\Rpackage}[1]{\textit{#1}}
\newcommand{\Rclass}[1]{\textit{#1}}
\newcommand{\Rfunction}[1]{{\small\texttt{#1}}}
\newcommand{\fixme}[1]{{\textbf{Fixme:} \textit{\textcolor{blue}{#1}}}}

\renewcommand{\floatpagefraction}{0.7}

\author{Sang Chul Choi\\[1em]Cornell University,\\ Ithaca, NY\\[1em]
\texttt{schoi@cornell.edu}}

\title{\textsf{\textbf{Analysing \textit{Streptococcus mutans} RNA-Seq data using R packages}}}

\begin{document}

\maketitle

\begin{abstract}
A number of RNA samples from \textit{Streptococcus mutans} are analyzed using R
packages: \Rpackage{DESeq}, \Rpackage{edgeR}, \Rpackage{goseq}, and \Rpackage{baySeq}.
Tasks include the analysis of count data from RNA-Seq for the detection of
differentially expressed genes.
Tests of enrichment for differentially expressed genes are performed using
\Rpackage{goseq}.
The package \Rpackage{smutans} contains the procedures.
This vignette explains the use of the package. 
\end{abstract}

\tableofcontents

%--------------------------------------------------
\section{Input data and preparations} \label{sec:prep}
%--------------------------------------------------

The \Rpackage{smutans} package expects count data in
the form of \Rclass{CountDataSet}.

To obtain such a count table for your own data, you will need
to create it from your sequence alignments and suitable annotation. Within Bioconductor,
you can use the function \Rfunction{summerizeOverlaps} in the
\Rpackage{GenomicRanges} package. See the vignette, Ref.\ \cite{summerizeOverlaps},
for a worked example. Another possibility (outside of Bioconductor)
is the \emph{htseq-count} script distributed with the HTSeq Python framework
\cite{htseq}. (You do not need to know any Python to use
\texttt{htseq-count}.) A third possibility might be given by the Bioconductor
package \Rpackage{easyrnaseq} (by Nicolas Delhomme; in preparation, available
soon; package name may change).

Another easy way to produce such a table from the output
  of the aligner is to use the \texttt{htseq-count} script distributed
  with the \emph{HTSeq} package. Even though \emph{HTSeq} is a Python
  package, you do not need to know any Python to use
  \texttt{htseq-count}. See
  \url{http://www-huber.embl.de/users/anders/HTSeq/doc/count.html}. (If you use
htseq-count, be sure to remove the extra lines with general counters (``ambiguous'' etc.)
when importing the data.)

The count values must be raw counts of
sequencing reads. This is important for \Rpackage{smutans}'s statistical model to hold,
as only raw reads allow to assess the measurement precision correctly. (Hence,
do not supply rounded values of normalized counts, or counts of covered base pairs.)

Furthermore, it is important that each column stems from an independent biological
replicate.  For purely technical replicates (e.\,g. when the same
library preparation was distributed over multiple lanes of the
sequencer in order to increase coverage), please sum up their counts
to get a single column, corresponding to a unique biological
replicate.  This is needed in order to allow \Rpackage{DESEq} to
estimate variability in the experiment correctly.

As an example dataset, we use the gene level read counts from the
\Rpackage{pasilla} data package. This dataset is from an experiment on
\emph{Drosophila melanogaster} cell cultures and
investigated the effect of RNAi knock-down of the splicing factor
\emph{pasilla}~\cite{Brooks2010}. The data are presented in the object
called \Robject{pasillaGenes}. For a description how this data object
was created from the raw data of Ref.~\cite{Brooks2010}, see the vingette
included with the \Rpackage{pasilla} package.

The \Robject{pasillaGenes} object is of class
\Rclass{CountDataSet}, which is the data container used by
\Rpackage{smutans}. We load the needed packages and the data as
follows.

<<options,results=hide,echo=FALSE>>=
options(digits=3)
@

There are 13 RNA-seq samples in \Robject{smutansGenes} that can be loaded
from the package.
<<smutans2,results=hide>>=
library( "smutans" )
data( "smutansGenes" )
@

\Robject{smutansGenes} contains the counts and also metadata about the samples:

<<look_into_smutansGenes>>=
head( counts(smutansGenes) )
pData( smutansGenes )
@

As you can see, the samples differ by experimental condition (glucose or
galactose) and by library type (ua159 or tw1). First, we only
look at the ua159 data for now. We will look at the tw1 data later.

We create an \Rclass{Smutans} object using \Rclass{CountDataSet} object,
smutansGenes.
<<ua159onlySamples>>=
ua159only <- newSmutans( smutansGenes )
ua159only <- smutans.de2type( ua159only, type="ua159", 
                              condA="glucose", condB="galactose" )
@

\Rfunction{smutans.de2type} uses a two-factor data set to test differential
expression of genes.
Note that the \Rclass{Smutans} contains \Rclass{CountDataSet} objects. 
\Rfunction{smutans.de2type} estimates the effective library size, and the
dispersions. Then, it tests for differential expression.

Because it is recommended to check
the dispersion estimates and to make sure that the data quality is as
expected, we plot the per-gene estimates against the normalized mean
expressions per gene, and then overlay the fitted curve in red.

\begin{figure}
\centering
\includegraphics[width=.6\textwidth]{smutans-figFit}
\caption{Empirical (black dots) and fitted (red lines) dispersion values
plotted against mean expression strength.}
\label{figFit}
\end{figure}

%Calling the function produces the plot (Fig.\ \ref{figFit}).

<<figFit,fig=TRUE>>=
smutans.plotDispersionEstimates(ua159only)
@

The plot in Figure~\ref{figFit} is doubly logarithmic; this may be
helpful or misleading, and it is worth experimenting with other
plotting styles.

As we estimated the
dispersion from only two samples, we should expect the estimates to
scatter with quite some sampling variance around their true values.
Hence, we \Rpackage{smutans} should not use the per-gene estimates directly in the test,
because using too low dispersion values leads to false positives.
Many of the values below the red line are likely to be underestimates of the
true dispersions, and hence, it is prudent to instead rather use the fitted value.
On the othe hand, not all of he values above the red line are overestimations, and
hence, the conservative choice is to keep them instead of replacing them with
their fitted values. If you do not like this default behaviour, you can change it
with the option \texttt{sharingMode} of \Rfunction{estimateDispersions}. Note that

In any case, the dispersion values which finally should be used by the subsequent testing are stored in
the feature data slot of \texttt{cds}:

%------------------------------------------------------------
\section{Inference: Calling differential expression} \label{sec:DE}
%------------------------------------------------------------

\begin{figure}
\centering
\includegraphics[width=.6\textwidth]{smutans-figDE}
\caption{Plot of normalised mean versus $log_2$ fold change
(this plot is sometimes also called the ``MA-plot'') for the contrast ``untreated''
versus ``treated''.}
\label{figDE}
\end{figure}

\subsection{Standard comparison between two experimental conditions}

Having estimated the dispersion for each gene, it is now straight-forward to look for
differentially expressed genes. To contrast two conditions, e.g., to see whether there is differential
expression between conditions ``untreated'' and ``treated'', we simply call the function \Rfunction{nbinomTest}.
It performs the tests as described in the paper and returns a data frame with the $p$ values and other
useful information.

For each gene, we get its mean expression level (at the base scale) as a joint estimate from both conditions, and
estimated separately for each  condition, the fold change from
the first to the second condition, the logarithm (to basis 2) of the fold change, and the
$p$ value for the statistical significance of this change. The \texttt{padj} column contains the
$p$ values, adjusted for multiple testing with the Benjamini-Hochberg procedure (see the R function
\Rfunction{p.adjust}), which controls false discovery
rate (FDR).

Let us first plot the $\log_2$ fold changes against the base means, colouring in red those genes
that are significant at 10\% FDR.
<<figDE,fig=TRUE>>=
smutans.plotDiffExp(ua159only)
@ %$
See Figure~\ref{figDE} for the plot. As we will use this plot more often, we have
stored its code in a function.

%------------------------------------------------------------
\section{Inference: Calling differential expression} \label{sec:DE}
%------------------------------------------------------------

<<allSamples>>=
allSample <- newSmutans( smutansGenes )
allSample <- smutans.de2( allSample )
@

\begin{figure}
\centering
\includegraphics[width=.6\textwidth]{smutans-figFit2}
\caption{Empirical (black dots) and fitted (red lines) dispersion values
plotted against mean expression strength.}
\label{figFit}
\end{figure}

Calling the function produces the plot (Fig.\ \ref{figFit2}).

<<figFit2,fig=TRUE>>=
smutans.plotDispersionEstimates(allSample)
@

<<compareUA159onlyAndAllSample>>=
padj.ua159only <- smutans.padj( ua159only )
padj.allSample <- smutans.padj( allSample )
tab = table( "ua159 only" = padj.ua159only < .1, "all samples" = padj.allSample < .1 )
addmargins( tab )
@
Of the \Sexpr{sum(tab)} gene, \Sexpr{tab[2,2]} genes are found in both analyses. The analysis using all
of the sample found \Sexpr{tab[1,2]} genes. The analysis using only ua159 found \Sexpr{tab[2,1]} genes.

\begin{figure}
\centering
\includegraphics[width=.6\textwidth]{smutans-figCompareUA159onlyAndAllSample}
\caption{Comparison of $p$ values (unadjusted) from the test using only the
four paired-end samples and the test using all seven samples. We can see that
the latter tend to be smaller, as expected from the higher power of the test with all samples.}
\label{figCompareUA159onlyAndAllSample}
\end{figure}

<<figCompareUA159onlyAndAllSample,fig=TRUE>>=
pval.ua159only <- smutans.pval( ua159only )
pval.allSample <- smutans.pval( allSample )
bottom = function(x, theta=1e-12) pmax(x, theta)
plot( bottom(pval.ua159only), bottom(pval.allSample), log="xy", pch=20, cex=.3 )
abline(a=0, b=1, col="blue")
@

<<allSampleWithoutType>>=
allSampleWithoutType <- newSmutans( smutansGenes )
allSampleWithoutType <- smutans.de2WithoutType( allSampleWithoutType,
                                                condA="glucose", 
                                                condB="galactose" )
@

<<compareUA159onlyAndAllSample>>=
padj.allSampleWithoutType <- smutans.padj( allSampleWithoutType )
tab2 <- table(
   `all samples simple` = padj.allSampleWithoutType < 0.1,
   `all samples GLM`    = padj.allSample < 0.1 )
addmargins( tab2 )
largeNumber <- 10
@
Of the \Sexpr{sum(tab2)} gene, \Sexpr{tab2[2,2]} genes are found in both analyses. The analysis using all
of the sample found \Sexpr{tab2[1,2]} genes. The analysis using all of the same
sample but without considering type found \Sexpr{tab2[2,1]} genes.

\begin{figure}
\centering
\includegraphics[width=.6\textwidth]{smutans-figLfc}
\caption{Scatterplot of direct (\Robject{lfc}) versus \emph{moderated} log-ratios (\Robject{mod\_lfc}).
  The moderation criterion used is variance stabilisation.
  The red points correspond to values that were infinite in \Robject{lfc}
  and were abitrarily set to $\pm\Sexpr{largeNumber}$ for the purpose of plotting.
  These values vary in a finite range (as shown in the plot) in \Robject{mod\_lfc}.}
\label{figLfc}
\end{figure}

<<figLfc,fig=TRUE>>=
select <- 1:40
smutans.de2TypeLfc( ua159only, condA="glucose", condB="galactose" )
@

\begin{figure}
\centering
\includegraphics[width=.6\textwidth]{smutans-figHeatmap1}
\caption{Heatmaps showing the expression data of the top \Sexpr{length(select)}
  differentially expressed genes. Left, the variance stabilisation
  transformed data,
  right, the original count data.
  The right plot is dominated by a small number of data points with large values.}
\label{figHeatmap1}
\end{figure}

<<figHeatmap1,fig=TRUE>>=
smutans.de2TypeHeatmap( ua159only )
@

\begin{figure}
\centering
\includegraphics[width=.6\textwidth]{smutans-figClust}
\caption{Heatmap showing the Euclidean distances between the samples
  as calculated from the variance-stabilising transformation of the
  count data.}
\label{figClust}
\end{figure}

<<figClust,fig=TRUE>>=
smutans.de2Clust( allSample )
@

\section{Session Info}
<<sessi>>=
sessionInfo()
@

\bibliographystyle{unsrt}
\bibliography{smutans}

\end{document}