### 2012-01-21 package/supplementary.Rnw

\documentclass{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{color} 
\usepackage{setspace} 
\usepackage{rotating}
\usepackage{natbib}
\usepackage{xr}
\usepackage{../../latex/sty/genres}
\usepackage{url}
\usepackage{longtable}

\externaldocument{manuscript}

% Text layout
\oddsidemargin 0in
\evensidemargin 0in
\topmargin -.5in
\textwidth 6.5in
\textheight 9in
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

\SweaveOpts{keep.source=TRUE,eps=FALSE,include=FALSE,width=6,height=6}
\newcommand{\Rpackage}[1]{R package \textit{#1}}
\newcommand{\program}[1]{{\texttt{#1}}}
\newcommand{\fixme}[1]{{\textbf{Fixme:} \textit{\textcolor{blue}{#1}}}}
\newcommand{\comment}[1]{}

\makeatother
\pagestyle{myheadings}
\markright{Transcriptome of Streptococcus mutans}

%

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}

\makeatother

\renewcommand{\thefootnote}{\textit{\alph{footnote}}}

% Leave date blank
\date{}

\pagestyle{myheadings}
%% ** EDIT HERE **
\markright{Transcriptome of Streptococcus mutans}

\begin{document}

\setcounter{secnumdepth}{-1}

\noindent \begin{LARGE}{\bf Defining Transcriptome using RNA Sequencing
 \\[0.7ex] 
of {\em Streptococcus mutans}}\end{LARGE}\\[2ex]
{\bf 
Sang Chul Choi, Charles D.\ Danko, Melissa J.\ Hubisz, \\
Paulina Dmitrievna Pavinski Bitar, \\
Robert A.\ Burne, Michael J.\ Stanhope, and Adam Siepel}%\\[0.5ex]

\tableofcontents

\listoftables

\listoffigures

\doublespacing

\clearpage

<<echo=FALSE,results=hide>>=
# not run when smutans R package is installed
# Use this when smutans R package is being tested.
library(DESeq)
library(GenomicRanges)
library(plotrix)
library(xtable)
source("smutans/R/class_and_slots.R")
source("smutans/R/core.R")
source("smutans/R/methods.R")
load("smutans/data/smutansData.tx.RData")
load("smutans/data/smutansData.txGenes.RData")
load("smutans/data/smutansData.txPileup.RData")
load("smutans/data/smutans.feature.genes.RData")
load("smutans/data/smutans.genes.criteria.RData")
load("smutans/data/smutans.go.genes.RData")
load("smutans/data/smutans.cat.desc.RData")
load("smutans/data/smutansGenes.RData")
@

\section{Supplementary Text}
A
\subsection{Summary statistics for recombinant trees}
\citep{Gorke2008}

\enlargethispage{\baselineskip}

\renewcommand*{\refname}{Literature Cited}
\bibliographystyle{../../latex/bst/mbe}
\bibliography{siepel-mutans}

\clearpage{}\setcounter{figure}{0}
\setcounter{table}{0}

\makeatletter 
\renewcommand{\thefigure}{S\@arabic\c@figure} 
\renewcommand{\thetable}{S\@arabic\c@table} 
\renewcommand{\figurename}{Supplementary Figure}
\renewcommand{\tablename}{Supplementary Table}
\makeatother

%\section{Supplementary Tables}

<<echo=FALSE,results=hide>>=
geneDescription <- read.delim("feature-genome.out-geneonly.txt",head=FALSE,sep="\t")
geneDescription <- geneDescription[,c(4,7)]
colnames(geneDescription) <- c("Gene ID","Gene Description")
@

<<echo=FALSE,results=tex>>=
#glucoseOnly <- newSmutans( smutansGenes, title="Glucose Only" )
#glucoseOnly <- smutans.de2( glucoseOnly, condition="glucose", 
#                            condA="ua159", condB="tw1" )
#save(glucoseOnly, file=file.path("glucoseOnly.RData"))
load("glucoseOnly.RData")
x <- smutans.de2List( glucoseOnly, qval=1e-3 )
x <- x[,c("id","baseMeanA","baseMeanB","log2FoldChange","padj")]
colnames(x) <- c("Gene ID","UA159","TW1","fold", "p-value")
x <- merge(x, geneDescription, by.x = "Gene ID", by.y = "Gene ID")
x.big <- xtable( x[order(x[,"p-value"]),],
                 display=c("s","s","d","d","f","e","s"),
                 digits=0,
                 align=c('l','l','r','r','r','r','p{3.5in}'),
                 label='tabGlucoseOnly',
                 caption='List of genes for the comparison UA159 and TW1 under glocuse-grown cells',
               )
print( x.big,
       tabular.environment='longtable',
       caption.placement="top",
       include.rownames=FALSE,
       floating=FALSE) 
@
\clearpage{}

<<echo=FALSE,results=tex>>=
#galactoseOnly <- newSmutans( smutansGenes, title="Galactose Only" )
#galactoseOnly <- smutans.de2( galactoseOnly, condition="galactose", 
#                              condA="ua159", condB="tw1" )
#save(galactoseOnly, file=file.path("galactoseOnly.RData"))
load("galactoseOnly.RData")
x <- smutans.de2List( galactoseOnly, qval=1e-3 )
x <- x[,c("id","baseMeanA","baseMeanB","log2FoldChange","padj")]
colnames(x) <- c("Gene ID","UA159","TW1","fold", "p-value")
x <- merge(x, geneDescription, by.x = "Gene ID", by.y = "Gene ID")
x.big <- xtable( x[order(x[,"p-value"]),],
                 display=c("s","s","d","d","f","e","s"),
                 digits=0,
                 align=c('l','l','r','r','r','r','p{3.5in}'),
                 label='tabGalactoseOnly',
                 caption='List of genes for the comparison UA159 and TW1 under galactose-grown cells',
               )
print( x.big,
       tabular.environment='longtable',
       caption.placement="top",
       include.rownames=FALSE,
       floating=FALSE) 
@
\clearpage{}

<<echo=FALSE,results=tex>>=
#ua159only <- newSmutans( smutansGenes, title="UA159 Only" )
#ua159only <- smutans.de2( ua159only, type="ua159", 
#                          condA="glucose", condB="galactose" )
#save(ua159only, file=file.path("ua159only.RData"))
load("ua159only.RData")
x <- smutans.de2List( ua159only, qval=1e-3 )
x <- x[,c("id","baseMeanA","baseMeanB","log2FoldChange","padj")]
colnames(x) <- c("Gene ID","Glucose","Galactose","fold", "p-value")
x <- merge(x, geneDescription, by.x = "Gene ID", by.y = "Gene ID")
x.big <- xtable( x[order(x[,"p-value"]),],
                 display=c("s","s","d","d","f","e","s"),
                 digits=0,
                 align=c('l','l','r','r','r','r','p{3.5in}'),
                 label='tabUA159Only',
                 caption='List of genes for the comparison of UA159 wild-type grown with glucose and galactose ',
               )
print( x.big,
       tabular.environment='longtable',
       caption.placement="top",
       include.rownames=FALSE,
       floating=FALSE) 
@
\clearpage{}

<<echo=FALSE,results=tex>>=
#tw1only <- newSmutans( smutansGenes, title="TW1 Only" )
#tw1only <- smutans.de2( tw1only, type="tw1", 
#                          condA="glucose", condB="galactose" )
#save(tw1only, file=file.path("tw1only.RData"))
load("tw1only.RData")
x <- smutans.de2List( tw1only, qval=1e-3 )
x <- x[,c("id","baseMeanA","baseMeanB","log2FoldChange","padj")]
colnames(x) <- c("Gene ID","Glucose","Galactose","fold", "p-value")
x <- merge(x, geneDescription, by.x = "Gene ID", by.y = "Gene ID")
x.big <- xtable( x[order(x[,"p-value"]),],
                 display=c("s","s","d","d","f","e","s"),
                 digits=0,
                 align=c('l','l','r','r','r','r','p{3.5in}'),
                 label='tabTW1Only',
                 caption='List of genes for the comparison of TW1 mutant-type grown with glucose and galactose ',
               )
print( x.big,
       tabular.environment='longtable',
       caption.placement="top",
       include.rownames=FALSE,
       floating=FALSE) 
@
\clearpage{}

%\section{Supplementary Figures}

%=============================================================================
%\begin{figure}[!ht]
%\includegraphics[scale=0.42]{figures/clonalorigin}
%\caption[Illustration of recombinant tree]{
%{\bf Illustration of a recombinant tree in the ClonalOrigin model.}
%(A) A recombinant tree for our five-sample clonal frame with four
%recombinant edges (dashed lines with arrows running from donors to
%recipients).  (B) Genomic 
%regions associated with recombinant edges.  Recombination 
%intensities (number of edge types present) and gene tree labels are shown at
%bottom. (C)~Induced local gene trees.
%Bold lines indicate branches of the clonal frame, while dashed lines indicate
%recombinant edges. See section \textbf{Summary statistics of recombinant trees} in
%\textbf{Supplementary Text} for additional details.}
%\label{fig:clonalorigin}
%\end{figure}
%\clearpage{}%
%=============================================================================

\end{document}