### 2012-03-05 package/smutans/inst/doc/smutans.Rnw

%\VignetteIndexEntry{Analyzing Streptococcus mutans RNA-seq data}
%\VignettePackage{smutans}

% To compile this document
% library('weaver'); rm(list=ls()); Sweave('smutans.Rnw', driver=weaver()); system('pdflatex smutans')

\documentclass{article}

\usepackage{Sweave}
\usepackage[letterpaper]{geometry}
\usepackage{hyperref,graphicx}
\usepackage{color}
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

\SweaveOpts{keep.source=TRUE,eps=FALSE,include=FALSE,width=4,height=4.5}
\newcommand{\Robject}[1]{object \texttt{#1}}
\newcommand{\Rpackage}[1]{R package \textit{#1}}
\newcommand{\Rclass}[1]{class \textit{#1}}
\newcommand{\Rfunction}[1]{{function \small\texttt{#1}}}
\newcommand{\program}[1]{{\texttt{#1}}}
\newcommand{\fixme}[1]{{\textbf{Fixme:} \textit{\textcolor{blue}{#1}}}}

\renewcommand{\floatpagefraction}{0.7}

\author{Sang Chul Choi\\[1em]Cornell University,\\ Ithaca, NY\\[1em]
\texttt{schoi@cornell.edu}}

\title{\textsf{\textbf{Analyzing \textit{Streptococcus mutans} RNA-seq data}}}

\begin{document}

\maketitle

\begin{abstract}
We analyze RNA samples from \textit{Streptococcus mutans} using
RNA-seq to define genes whose expression changes between samples
\cite{Oshlack2010}. Short reads are first mapped to a reference genome using
approaches such as \program{BWA} \cite{Li2009a} or \program{subread}.
Reads fitting into one of the annotated
intervals are counted.  These counts are passed to a statistical package which
detects differences in read counts for statistical significance. For this
purpose, we make use of \Rpackage{DEseq} \cite{Anders2010}.
This prodecure returns a list of differentially expressed
genes, which can be passed to downstream analysis, such as functional category
enrichment test using \Rpackage{goseq} \cite{Young2010}.  The current
\Rpackage{smutans} contains procedures that use functions provided by
the above R packages. 
\end{abstract}

\tableofcontents

%--------------------------------------------------
\section{Data preparations} \label{sec:prep}
%--------------------------------------------------

Analyzing RNA-seq data for determining differentially expressed genes
are well described \cite{Oshlack2010}. We used a short read aligner called
\program{BWA} \cite{Li2009a} to map the sequenced short reads on a reference
genome, \textit{Streptococcus mutans} UA159 (NCBI accession NC\_004350). Because
the aligner allowed a few gaps for efficient alignment of millions of sequences
of size being 100 base pairs long, short reads that consist of RNAs and
sequencing adapter sequences would not be mapped. By using \program{cutadapt}
\cite{Martin2011} we removed parts of adapter sequences from short reads before
mapping on the reference genome.  We also trimmed short read sequences by quality score from
the 3' end with threshold score of 20 \cite{Schmieder2011}. Mapped short read
alignments were converted into readable formats \cite{Li2009b}. We counted short
reads aligned to annotated genes in the reference genome, forming a table of
read counts of the genes. Statistical software R \cite{R2011} with
\Rpackage{DESeq} \cite{Anders2010} was subsequently employed to infer
differentially expressed in different biological conditions. Most of functions
in comparing different experiments were derived from \Rpackage{DESeq}. Readers
should refer to the package for details.  We also generated
tracks for a Streptococcus Genome Browser
(http://strep-genome.bscb.cornell.edu) that summarize gene expression levels
based on the short read alignments. This allowed us to visualize parts of the
reference genome for comparing different RNA samples. Further analyses such as
functional category enrichment test \cite{Young2010} were also employed.  To
obtain count tables of short reads we used scripts available at a website
http://github.com/goshng/RNASeq-Analysis. In the following we will use
the count table to determine differentially expressed genes.

The \Rpackage{smutans} expects count data in the form of \Rclass{CountDataSet},
which is a type supported by \Rpackage{DESeq}.  Note that the count values
are raw counts of sequencing reads.  The \Rpackage{DESeq}'s statistical model,
or negative binomial model, holds for raw reads counts that allows to assess the
measurement precision correctly; we did not have rounded values of normalized
counts, or counts of covered base pairs to form count tables.  Several
exeriments of \textit{S.\ mutans} 
RNA sampling were performed to examine effects of glucose and galactose,
835NP and 835P, OMZ175 and OMZ/HKRR, CSP, and acid tolerance.  The comparison of
glucose and galactose is presented in the \Robject{smutansGenes} as a part of
\Rpackage{smutans}.

<<options,results=hide,echo=FALSE>>=
options(digits=3)
@
In the following chapter, we use \Robject{smutansGenes}, which is of type
\Rclass{CountDataSet} of \Rpackage{DESeq}.  Load the \Rpackage{smutans}, and the
data \Robject{smutansGenes}. 
<<smutans,results=hide>>=
library( "smutans" )
@

Browse the installed directory of \Rpackage{smutans} for several count
table and index files. As we go along the chapters of this vignette, we will use
them.
<<readSmutans>>=
inDir <- system.file("extdata", package="smutans", mustWork=TRUE)
dir(inDir,pattern="count*")
@

Look at two kinds of files: count.txt and count.txt.index. We use
\Rfunction{readSmutans} to read a count table and its index files. You can check
if two factor levels are correctly assigned using \Rfunction{pData}. Refer to
the help documentation of the function in R.
<<eval=FALSE>>=
help(readSmutans)
@
<<readSmutans>>=
smutansGenes <- 
  readSmutans (countsFile=file.path(inDir, "count-ua159-bwa.txt"),
               indexFile =file.path(inDir, "count-ua159.txt.index"),
               condition = c("UA159GLU", "UA159GAL", "TW1GLU", "TW1GAL"),
               firstFactor = c("UA159GLU", "UA159GAL"),
               firstFactorLabel = c("ua159", "tw1"),
               secondFactor = c("UA159GLU", "TW1GLU"),
               secondFactorLabel = c("glucose","galactose"),
               name="S. mutans UA159, TW1, Glucose, and Galactose")
@

The following functions create R data objects. Users are not recommended to use
them because they have been already created and saved together in the package.
They will be replaced by functions of \Rpackage{smutans}. They are documented
for the sake of reminders.
<<samplePreparation,eval=FALSE>>=
# not run
# but create a R scirpt file with the following scripts, and
# run it, i.e., Rscript run.R
library(DESeq)
source("smutans/R/core.R")
smutans.prepareData34( )
smutans.prepareDataUA159TW1( ) 
smutans.prepareGoseq( )
smutans.prepareTranscript( )
q("no")
@
Now we look at the detail of the loaded count table. There are 
\Sexpr{length(colnames(counts(smutansGenes)))}
RNA-seq samples in \Robject{smutansGenes}, which
contains the counts and also metadata about the samples:
<<look_into_smutansGenes>>=
head( counts(smutansGenes) )
@
<<look_into_smutansGenes>>=
pData( smutansGenes )
@
The samples differ by experimental condition (``glucose'' or ``galactose'') and
by library type (``ua159'' or ``tw1'').  Use the following functions to get
helps such as opening the vignette and manual of \Rpackage{smutans}. 
<<eval=FALSE>>=
help.start()
vignette(package="smutans")
openVignette("smutans")
@

%--------------------------------------------------
\section{DE analysis using UA159 and TW1 samples with glucose and galactose}
%--------------------------------------------------

We compared four pairs of experiments using the \Robject{smutansGenes}: two comparisons
between ``glucose'' and ``galactose'' given either type ``ua159'' or ``tw1'',
and another two comparisons between types ``ua159'' and ``tw1'' given either
condition ``glucose'' or ``galactose.'' Before comparing these we clustered
samples using read counts. Figure \ref{figClust} shows the cluster of the
\Sexpr{length(colnames(counts(smutansGenes)))} samples. Refer to
\Rpackage{DESeq} vignette as to how the clustering is conducted.  The clustering
was useful in ascertaining that the nutrient effect was stronger than the mutant
effect: two experimental conditions of ``glucose'' and ``galactose''
distinguished the samples. Within each condition, two different types of
``ua159'' and ``tw1'' separated them.

\begin{figure}
\centering
\includegraphics[width=.9\textwidth]{smutans-figClust}
\caption{Heatmap showing the Euclidean distances between the samples
as calculated from the variance-stabilizing transformation of the
count data.}
\label{figClust}
\end{figure}
<<figClust,fig=TRUE>>=
allSample <- newSmutans( smutansGenes )
smutans.de2Clust( allSample )
@

%------------------------------------------------------------
\subsection{Inference using single-factor data} 
%------------------------------------------------------------

Let us compare ``glucose'' and ``galactose'' conditions given ``ua159'' type. 
\Robject{smutansGenes}
contains short reads counts based on bwa alignment on UA159 genome.  The title
argument would be useful in distinguishing different objects of \Rclass{Smutans}
as we have more R objects of the same type. A subsequent call to
\Rfunction{smutans.de2} uses the two-factor data set to test differential
expression for the annotated genes by estimating
effective library sizes, and dispersions using functions provided by
\Rpackage{DESeq}.
<<ua159only>>=
ua159only <- newSmutans( smutansGenes, title="UA159 Only" )
ua159only <- smutans.de2( ua159only, type="ua159", 
                          condA="glucose", condB="galactose" )
head(smutans.de2List( ua159only ))
@
Descriptions of the columns are in order.
Each gene has its mean expression level (at the base scale) as a joint estimate
from both conditions, and estimated separately for each condition, the fold
change from the first to the second condition, i.e.,
$\frac{baseMeanB}{baseMeanA}$, the logarithm (to basis 2) of the fold change,
and the $p$ value for the statistical significance of this change. The
\texttt{padj} column contains the $p$ values, adjusted for multiple testing with
the Benjamini-Hochberg procedure (see the \Rfunction{p.adjust}),
which controls false discovery rate (FDR). Later, we print the gene list in a
CSV file using option file of \Rfunction{smutans.de2List}. Windows users should
be able to open CSV files with a spreadsheet program such as EXCEL. Users who
are more comfortable with the EXCEL program than R could post-process CSV files
for sorting and finding differentially expressed genes using different q-value
cutoff values.
<<eval=TRUE>>=
smutans.de2List( ua159only, "ua159only.csv" )
@

By fixing ``tw1'' type similarly we compared ``glucose'' and ``galactose''
conditions.
<<tw1onlySamples>>=
tw1only <- newSmutans( smutansGenes, title="TW1 Only" )
tw1only <- smutans.de2( tw1only, type="tw1", 
                        condA="glucose", condB="galactose" )
@
We can print the result in a csv file.
<<eval=TRUE>>=
# not run
smutans.de2List( tw1only, "tw1only.csv" )
@
We compared ``glucose'' and ``galactose'' with either ``ua159'' or ``tw1''
fixed. We computed numbers of differentially expressed genes that were
determined by both or either one of them.
<<>>=
tabUA159onlyTW1only <- smutans.compareResult( ua159only, tw1only )
addmargins( tabUA159onlyTW1only )
@
The table, \Robject{tabUA159onlyTW1only}, shows numbers of genes found to be 
differentially expressed in the two analyses: ``UA159 Only'' and ``TW1 Only.''
Of the total \Sexpr{sum(tabUA159onlyTW1only)} genes, \Sexpr{tabUA159onlyTW1only[2,2]}
genes were found in both analyses. The analysis using TW1-only sample found
\Sexpr{tabUA159onlyTW1only[1,2]} among genes UA159-only sample could not
find. The analysis using UA159-only sample found
\Sexpr{tabUA159onlyTW1only[2,1]} genes among ones TW1-only sample could not
find. UA159-only sample detected more differentially expressed genes.

We now compare sample types on differentially expressed genes. Create two
\Rclass{Smutans} objects: \Robject{glucoseOnly} and \Robject{galactoseOnly}, and
compare them in detecting differentially expressed genes:
<<glucoseOnlySamples>>=
glucoseOnly <- newSmutans( smutansGenes, title="Glucose Only" )
glucoseOnly <- smutans.de2( glucoseOnly, condition="glucose", 
                            condA="ua159", condB="tw1" )
galactoseOnly <- newSmutans( smutansGenes, title="Galactose Only" )
galactoseOnly <- smutans.de2( galactoseOnly, condition="galactose", 
                              condA="ua159", condB="tw1" )
tabGlucoseOnlyGalactoseOnly <- smutans.compareResult( glucoseOnly, 
                                                      galactoseOnly )
addmargins( tabGlucoseOnlyGalactoseOnly )
@
Of the total \Sexpr{sum(tabGlucoseOnlyGalactoseOnly)} genes,
\Sexpr{tabGlucoseOnlyGalactoseOnly[2,2]} genes were found in both analyses at FDR
rate of 10\%. The analysis using galactose-only sample found
\Sexpr{tabGlucoseOnlyGalactoseOnly[1,2]} of the genes that were negative in
glucose-only sample.  The analysis using glucose-only
sample found \Sexpr{tabGlucoseOnlyGalactoseOnly[2,1]} of the genes that were
negative in galactose-only sample. 
Table \ref{tab:de} summarizes the number of differentially expressed genes that
were found using single-factor analyses at FDR of 10\%. We can print them in csv
files using \Rfunction{smutans.de2List}.
<<eval=TRUE>>=
# not run
smutans.de2List( glucoseOnly, "glucoseOnly.csv" )
smutans.de2List( galactoseOnly, "galactoseOnly.csv" )
@

%-----------------------------------------------------------------------------
\begin{table}[!h]
\caption{{\bf Numbers of differentially expressed genes in different analyses}}
\vspace{1ex}
\noindent \begin{centering}
\begin{tabular}{ccccl}
\hline 
Sample A & Sample B & Number of Genes & Result File \\
\hline 
UA159 Glucose & UA159 Galactose & \Sexpr{sum(ua159only@padj < .1,na.rm=T)} & ua159only.csv \\
TW1 Glucose & TW1 Galactose & \Sexpr{sum(tw1only@padj < .1,na.rm=T)} & tw1only.csv \\
UA159 Glucose & TW1 Glucose & \Sexpr{sum(glucoseOnly@padj < .1,na.rm=T)} & glucoseOnly.csv \\
UA159 Galactose & TW1 Galactose & \Sexpr{sum(galactoseOnly@padj < .1,na.rm=T)} & galactoseOnly.csv \\
\hline 
\end{tabular}
\par\end{centering}
\label{tab:de}
\end{table}
%-----------------------------------------------------------------------------

%------------------------------------------------------------
\subsection{Inference using all of the samples} 
%------------------------------------------------------------

Let us determine genes differentially expressed in different conditions of
``glucose'' and ``galactose'' using both types of ``ua159'' and ``tw1'': a
two-factor analysis. Start
with the two-factor analysis by creating a new \Rclass{Smutans} object, called
\Robject{allSample}. Although we already used \Robject{allSample} when
clustering the samples, we recreate it for instructive purposes.  The
\Rfunction{smutans.de2} can use a two-factor data set to perform two-factor
analysis: we are here interested in condition effect. 
<<allSamples,results=hide>>=
allSample <- newSmutans( smutansGenes, 
                         title="All Samples for Comparing Conditions" )
allSample <- smutans.de2( allSample,
                          condA="glucose", condB="galactose" )
tabUA159onlyAndAllSample <- smutans.compareResult( ua159only, 
                                                   allSample )
@
<<>>=
addmargins( tabUA159onlyAndAllSample )
@
Of the total \Sexpr{sum(tabUA159onlyAndAllSample)} gene,
\Sexpr{tabUA159onlyAndAllSample[2,2]} genes were found in both analyses. 
Using all of the samples detected more differentially expressed genes than did
ua159-only sample.
The output from \Robject{allSample} is almost the same except that there are
convergence column.
<<>>=
head( smutans.de2List( allSample ) )
@

We can show that taking the library type into account was important to have
good detection power by doing the analysis againt using the standard workflow,
as outlined earlier, and without informing \Rpackage{DESeq} of the library
types. This anlaysis is not necessary. However, it clearly shows the advantage of the
two-factor analysis.
<<allSampleWithoutType>>=
allSampleWithoutType <- newSmutans( smutansGenes, "All Sample without Type" )
allSampleWithoutType <- smutans.de2WithoutType( allSampleWithoutType,
                                                condA="glucose", 
                                                condB="galactose" )
tabWithoutWithAllSample <- smutans.compareResult( allSampleWithoutType, 
                                                  allSample )
addmargins( tabWithoutWithAllSample )
@
Using all of the sample that considers types found more differentially expressed
genes than did one ignoring the types.

We now proceed to compare the two types by considering all of the samples, and
compare the result with the single-factor analysis.
<<allSamplesType,results=hide>>=
allSampleType <- newSmutans( smutansGenes, title="All samples for type" )
allSampleType <- smutans.de2( allSampleType, compareCondition="no",
                              condA="ua159", condB="tw1" )
tabGlucoseOnlyAllSample <- smutans.compareResult( glucoseOnly, 
                                                  allSampleType )
tabGalactoseOnlyAllSample <- smutans.compareResult( galactoseOnly, 
                                                    allSampleType )
@                                                    
<<>>=
addmargins( tabGlucoseOnlyAllSample )
addmargins( tabGalactoseOnlyAllSample )
@
Of the total \Sexpr{sum(tabGlucoseOnlyAllSample)} genes,
\Sexpr{tabGlucoseOnlyAllSample[2,2]} genes were found in both analyses. The
analysis using all of the sample found \Sexpr{tabGlucoseOnlyAllSample[1,2]}
of the genes that glucose-only sample did not find. 
The analysis using glucose-only sample
found \Sexpr{tabGlucoseOnlyAllSample[2,1]} the genes that all samples did not
find.  In the comparison using galactose-only sample,
of the \Sexpr{sum(tabGalactoseOnlyAllSample)} genes,
\Sexpr{tabGalactoseOnlyAllSample[2,2]} genes were found in both analyses. The
analysis using all of the sample found \Sexpr{tabGalactoseOnlyAllSample[1,2]}
the genes galactose-only sample did not find. 
The analysis using galactose-only sample
found \Sexpr{tabGalactoseOnlyAllSample[2,1]} the genes that all samples did not
find.

One can print the sults in CSV files.
<<eval=TRUE>>=
smutans.de2List( allSample, "allSample.csv" )
smutans.de2List( allSampleType, "allSampleType.csv" )
@

%------------------------------------------------------------
\subsection{Functional category association analyses} 
%------------------------------------------------------------

We find some functional categories enriched with differentially expressed genes
using \Rpackage{goseq} \cite{Young2010} (table \ref{tab:goseq}).
To find functional categories enriched with differentially expressed genes we
first find a list of differentially expressed genes. A set of genes are assigned
to functional gene categories: \Robject{smutans.go.genes}. Features of the genes in
UA159 include gene lengths: \Robject{smutans.feature.genes}.
Functional categories with numbers of genes assigned are in
\Robject{smutans.cat.desc}.
<<goseqUA159only,eval=FALSE>>=
data( "smutans.feature.genes" )
data( "smutans.go.genes" )
data( "smutans.cat.desc" )
smutans.de2Goseq ( ua159only, qval=1e-2, 
                   feature.gene=smutans.feature.genes,
                   go.genes=smutans.go.genes,
                   cat.desc=smutans.cat.desc )
@
%-----------------------------------------------------------------------------
\begin{table}[!h]
\caption{{\bf Gene Ontology (GO) enrichments for differentially expressed
genes}}
\vspace{1ex}
\noindent \begin{centering}
\begin{tabular}{cl}
\hline 
$q^a$ & Description \\
\hline 
%-----------------------------------------------------------------------------
<<goseqUA159only,echo=FALSE,results=tex>>=
data( "smutans.feature.genes" )
data( "smutans.go.genes" )
data( "smutans.cat.desc" )
smutans.de2Goseq ( ua159only, qval=1e-2, 
                   feature.gene=smutans.feature.genes,
                   go.genes=smutans.go.genes,
                   cat.desc=smutans.cat.desc )
@
%-----------------------------------------------------------------------------
\hline 
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$False discovery rate estimated by the
Benjamini-Hochberg method.  All categories having at least ten genes and
$q\leq 0.05$ are displayed.\\ 
\end{flushleft}
\label{tab:goseq}
\end{table}
%-----------------------------------------------------------------------------

%--------------------------------------------------
\section{DE analysis using 835NP, 835P, and UA159 no CSP}
%--------------------------------------------------

Sang-Joon Ahn is leading this experiment. Let us create the input data variable.
We compare three comparisons using three conditions. We do the cluster,
determine differentially expressed genes, and print those genes.  One of the
three conditions, 835P, is distinguished from the others.  We found no
functional categories enriched with differentially expressed genes in the
pairwise comparisons of 835NP, 835P, and UA159 no CSP.  Let's make a list of
differentially expressed genes.  
<<readSmutans>>=
sm835Genes <- 
  readSmutans (countsFile=file.path(inDir, "count-ua159-bwa.txt"),
                indexFile =file.path(inDir, "count-ua159.txt.index"),
                condition = c("835NP", "835P", "UA159noCSP"),
                firstFactor = c("835NP", "835P", "UA159noCSP"),
                firstFactorLabel = c("ua159"),
                secondFactor = c(),
                secondFactorLabel = c("835np","835p", "ua159nocsp"),
                name="S. mutans 835NP, 835P, UA159 no CSP")
@

<<>>=
head( counts(sm835Genes), n=1 )
pData( sm835Genes )
@

<<figClustNpAndP,fig=TRUE>>=
npAndP <- newSmutans( sm835Genes, title="835NP vs. 835P" )
smutans.de2Clust( npAndP )
@
\begin{figure}
\centering
\includegraphics[width=.9\textwidth]{smutans-figClustNpAndP}
\caption{Empirical (black dots) and fitted (red lines) dispersion values
plotted against mean expression strength.}
\label{figClustNpAndP}
\end{figure}
We first cluster the \Sexpr{length(colnames(counts(sm835Genes)))} samples
(fig.\ \ref{figClustNpAndP}).

<<sm835Samples>>=
npAndP <- newSmutans( sm835Genes, title="835NP vs. 835P" )
npAndP <- smutans.de2( npAndP, type="ua159", 
                       condA="835np", condB="835p" )
npAndUA159 <- newSmutans( sm835Genes, title="835NP vs. UA159" )
npAndUA159 <- smutans.de2( npAndUA159, type="ua159", 
                           condA="835np", condB="ua159nocsp" )
pAndUA159 <- newSmutans( sm835Genes, title="835P vs. UA159" )
pAndUA159 <- smutans.de2( pAndUA159, type="ua159", 
                          condA="835p", condB="ua159nocsp" )
@

<<eval=TRUE>>=
# not run
smutans.de2List( npAndP, "npAndP.csv" )
smutans.de2List( npAndUA159, "npAndUA159.csv" )
smutans.de2List( pAndUA159, "pAndUA159.csv" )
@

%--------------------------------------------------
\section{DE analysis using OMZ175 and OMZ/HKRR}
%--------------------------------------------------

We can use either UA159 or OMZ175 genome as a reference genome.  We first
cluster the samples (fig.\ \ref{figClustomz}).  We did not find many genes in
this comparison (see below).

We also mapped short reads on OMZ175 contigs. We wished to find genes in OMZ175 but not in
UA159 to see whether they are enriched with differentially expressed genes. We
also wished to find genes in UA159 but not in OMZ175 to check in the same
fashion. What are the distribution of the genes that unique to either UA159 or
OMZ175 in differentially expressed genes? We could test the independence of
differntially expressed genes and unique genes.
Use 0.1 as the q-value cutoff to have two sets of differentially expressed genes
and not ones. And also divide genes unique to SMU109 and not one. Test the
independence of differential expression and SMU109 unique genes using Fisher's
exact test. This was not useful because of small number of differentially
expressed genes.
<<readSmutans>>=
smomzGenes <- 
  readSmutans (countsFile=file.path(inDir, "count-ua159-bwa.txt"),
               indexFile =file.path(inDir, "count-ua159.txt.index"),
               condition = c("OMZ175", "OMZHKRR"),
               firstFactorLabel = c("ua159"),
               secondFactorLabel = c("OMZ175", "OMZHKRR"),
               name="S. mutans OMZ175, and OMZHKRR based on UA159")
@

<<>>=
head( counts(smomzGenes), n=1 )
pData( smomzGenes )
@

<<figClustomz,fig=TRUE>>=
omz <- newSmutans( smomzGenes, title="OMZ175 vs. OMZ/HKRR" )
smutans.de2Clust( omz )
@

<<smomz,results=hide>>=
omz <- smutans.de2( omz, type="ua159", 
                    condA="OMZ175", condB="OMZHKRR" )
@

<<>>=
smutans.de2List( omz )
@

<<eval=TRUE>>=
# not run
smutans.de2List( omz, "omz.csv" )
@

<<readSmutans>>=
smomzGenesSMU109 <- 
  readSmutans (countsFile=file.path(inDir, "count-omz-bwa.txt"),
               indexFile =file.path(inDir, "count-omz.txt.index"),
               condition = c("OMZ175", "OMZHKRR"),
               firstFactorLabel = c("smu109"),
               secondFactorLabel = c("OMZ175", "OMZHKRR"),
               name="S. mutans OMZ175, and OMZHKRR on SMU109 contigs")
@

<<>>=
head( counts(smomzGenesSMU109), n=1 )
pData( smomzGenesSMU109 )
@

<<figClustomzSMU109,fig=TRUE>>=
omzSMU109 <- newSmutans( smomzGenesSMU109, title="OMZ175 vs. OMZ/HKRR" )
smutans.de2Clust( omzSMU109 )
@

\begin{figure}
\centering
\includegraphics[width=.4\textwidth]{smutans-figClustomz}
\includegraphics[width=.4\textwidth]{smutans-figClustomzSMU109}
\caption{Empirical (black dots) and fitted (red lines) dispersion values
plotted against mean expression strength. Left-hand side is for UA159-based
result, and right-hand side for SMU109 or OMZ175 contigs-based.}
\label{figClustomz}
\end{figure}

<<smomz,results=hide>>=
omzSMU109 <- smutans.de2( omzSMU109, type="smu109", 
                          condA="OMZ175", condB="OMZHKRR" )
@

<<>>=
smutans.de2List( omzSMU109 )
@

<<eval=TRUE>>=
# not run
smutans.de2List( omzSMU109, "omzSMU109.csv" )
@

%--------------------------------------------------
\section{DE analysis using UA159 CSP and SMU86 CSP}
%--------------------------------------------------

We follow the procedure in the previous chapter on the comparison of glucose and
galactose. The cluster of the samples shows that the effect of types ua159 and
smu86 is stronger than that of conditions CSP (fig.\ \ref{figClustCSP}).
Differences due to CSP was more profound in ua159 type than was in smu86
(table \ref{tab:decsp}).

We used 0.01 as the q-value cutoff to have two sets of differentially expressed genes
and not ones. And also divide genes unique to SMU86 and not one. Test the
independence of differential expression and SMU86 unique genes using Fisher's
exact test. We found evidence that the differences were due to genes
unique to SMU86. However, we did not find evidence that differences were due to genes
unique to UA159. The proportion of genes unique to SMU86 was larger than that
of common genes. This shows that genes unique to SMU86 have more profound
effects than common genes under CSP, but genes unique to UA159 does not have
those effects.

<<options,results=hide>>=
csp.qval.cutoff <- 10^-1
@

<<readSmutans>>=
<<readSmutans>>=
cspGenes <-
  readSmutans (countsFile=file.path(inDir, "count-ua159-bwa.txt"),
                indexFile =file.path(inDir, "count-ua159.txt.index"),
                condition = c("UA159noCSP", "UA159CSP", "Smu86noCSP", "Smu86CSP"),
                firstFactor = c("UA159noCSP", "UA159CSP"),
                firstFactorLabel = c("ua159", "smu86"),
                secondFactor = c("UA159noCSP", "Smu86noCSP"),
                secondFactorLabel = c("nocsp","csp"),
                name="S. mutans no CSP and with CSP")
@

\begin{figure}
\centering
\includegraphics[width=.9\textwidth]{smutans-figClustCSP}
\caption{Heatmap showing the Euclidean distances between the samples
as calculated from the variance-stabilizing transformation of the
count data.}
\label{figClustCSP}
\end{figure}
<<figClustCSP,fig=TRUE>>=
allSampleCSP <- newSmutans( cspGenes )
smutans.de2Clust( allSampleCSP )
@

<<>>=
ua159csp <- newSmutans( cspGenes, title="UA159" )
ua159csp <- smutans.de2( ua159csp, type="ua159", 
                         condA="nocsp", condB="csp" )
@

<<eval=TRUE>>=
# not run
smutans.de2List( ua159csp, "ua159csp.csv" )
@

<<smu86csp>>=
smu86csp <- newSmutans( cspGenes, title="SMU86" )
smu86csp <- smutans.de2( smu86csp, type="smu86", 
                             condA="nocsp", condB="csp" )
@

<<eval=TRUE>>=
# not run
smutans.de2List( smu86csp, "smu86csp.csv" )
@

<<>>=
tabUA159SMU86 <- smutans.compareResult( ua159csp, smu86csp,
                                        qval=csp.qval.cutoff )
addmargins( tabUA159SMU86 )
@

<<>>=
nocsp <- newSmutans( cspGenes, title="NO CSP" )
nocsp <- smutans.de2( nocsp, condition="nocsp", 
                        condA="ua159", condB="smu86" )
csp <- newSmutans( cspGenes, title="CSP" )
csp <- smutans.de2( csp, condition="csp", 
                        condA="ua159", condB="smu86" )
tabnocspcsp <- smutans.compareResult( nocsp, csp, 
                                      qval=csp.qval.cutoff )
addmargins( tabnocspcsp )
@

<<eval=TRUE>>=
# not run
smutans.de2List( nocsp, "nocsp.csv" )
smutans.de2List( csp, "csp.csv" )
@

<<allSamplesSmu86,results=hide>>=
allSampleSmu86 <- newSmutans( cspGenes, 
                              title="All Samples for Comparing CSP and no CSP" )
allSampleSmu86 <- smutans.de2( allSampleSmu86,
                               condA="nocsp", condB="csp" )
tabUA159cspAndAllSampleSmu86 <- smutans.compareResult( ua159csp, 
                                                       allSampleSmu86,
                                                       qval=csp.qval.cutoff )
tabSMU86cspAndAllSampleSmu86 <- smutans.compareResult( smu86csp, 
                                                       allSampleSmu86, 
                                                       qval=csp.qval.cutoff )
@
<<>>=
addmargins( tabUA159cspAndAllSampleSmu86 )
addmargins( tabSMU86cspAndAllSampleSmu86 )
@

<<eval=TRUE>>=
# not run
smutans.de2List( allSampleSmu86, "allSampleSmu86.csv" )
@

<<allSamplesCsp,results=hide>>=
allSampleCsp <- newSmutans( cspGenes, 
                              title="All Samples for Comparing SMU86" )
allSampleCsp <- smutans.de2( allSampleCsp, compareCondition="no",
                               condA="ua159", condB="smu86" )
tabnocspAndAllSampleCsp <- smutans.compareResult( nocsp, 
                                                  allSampleCsp,
                                                  qval=csp.qval.cutoff )
tabcspAndAllSampleCsp <- smutans.compareResult( csp, 
                                                allSampleCsp,
                                                qval=csp.qval.cutoff )
@
<<>>=
addmargins( tabnocspAndAllSampleCsp )
addmargins( tabcspAndAllSampleCsp )
@

<<eval=TRUE>>=
# not run
smutans.de2List( allSampleCsp, "allSampleCsp.csv" )
@

%-----------------------------------------------------------------------------
\begin{table}[!h]
\caption{{\bf Numbers of differentially expressed genes in CSP
analyses.} We use multiple testing adjusted p-value of $\Sexpr{csp.qval.cutoff}$.}
\vspace{1ex}
\noindent \begin{centering}
\begin{tabular}{ccccl}
\hline 
Sample A & Sample B & Number of Genes & Result File \\
\hline 
UA159 no CSP & UA159 CSP    & \Sexpr{sum(ua159csp@padj < csp.qval.cutoff,na.rm=T)} & ua159ph.csv \\
SMU86 no CSP & SMU86 CSP    & \Sexpr{sum(smu86csp@padj < csp.qval.cutoff,na.rm=T)} & smu21ph.csv \\
UA159 no CSP & SMU86 no CSP & \Sexpr{sum(nocsp@padj < csp.qval.cutoff,na.rm=T)} & ph7only.csv \\
UA159 CSP & SMU86 CSP       & \Sexpr{sum(csp@padj < csp.qval.cutoff,na.rm=T)} & ph5only.csv \\
\hline 
\end{tabular}
\par\end{centering}
\label{tab:decsp}
\end{table}
%-----------------------------------------------------------------------------

Use SMU86 contigs as a reference genome.

<<readSmutans>>=
smu86cspGenes <-
  readSmutans (countsFile=file.path(inDir, "count-smu86-bwa.txt"),
                indexFile =file.path(inDir, "count-smu86.txt.index"),
                condition = c("Smu86noCSP", "Smu86CSP"),
                firstFactorLabel = c("smu86"),
                secondFactorLabel = c("nocsp","csp"),
                name="S. mutans SMU86 nocsp and csp")
@

<<>>=
smu86cspSMU86 <- newSmutans( smu86cspGenes, title="SMU86" )
smu86cspSMU86 <- smutans.de2( smu86cspSMU86, type="smu86", 
                              condA="nocsp", condB="csp" )
@

<<eval=TRUE>>=
# not run
smutans.de2List( smu86cspSMU86, "smu86cspSMU86.csv" )
@

<<eval=TRUE>>=
x <- scan(file.path(inDir, "SMU86.diff.lst"), what="character")
y <- smutans.de2List( smu86cspSMU86, qval=10^-2)$id
z <- smu86cspSMU86@res$id
x1 <- z[z %in% x]
x2 <- z[!z %in% x]
y1 <- z[z %in% y]
y2 <- z[!z %in% y]
sum(x1 %in% y1)
sum(x1 %in% y2)
sum(x2 %in% y1)
sum(x2 %in% y2)
unique.diff <-
matrix(c(4, 80, 12, 1710),
       nrow = 2,
       dimnames =
       list(c("Diff", "Not diff"),
            c("Unique", "Common")))
fisher.test(unique.diff)
@

<<>>=
x <- scan(file.path(inDir, "UA159-SMU86.diff.lst"), what="character")
y <- smutans.de2List( smu86csp, qval=10^-2)$id
z <- smu86csp@res$id
x1 <- z[z %in% x]
x2 <- z[!z %in% x]
y1 <- z[z %in% y]
y2 <- z[!z %in% y]
sum(x1 %in% y1)
sum(x1 %in% y2)
sum(x2 %in% y1)
sum(x2 %in% y2)
unique.diff <-
matrix(c(1, 216, 12, 1731),
       nrow = 2,
       dimnames =
       list(c("Diff", "Not diff"),
            c("Unique", "Common")))
fisher.test(unique.diff)
@

%--------------------------------------------------
\section{DE analysis using acid tolerance}
%--------------------------------------------------

Sara is leading the experiment. 
The clustering shows that the effect of type, ua159 vs\. smu21, is stronger
than pH effect (fig.\ \ref{figClustpH}).
There seem to be too many differentially expressed genes in the comparison of
different pH (table \ref{tab:deph}). We therefore used a much smaller cutoff value of
multiple testing adjusted $p$-value. 
When comparing two conditions, pH 7 and pH 5, we found more differentially
expressed genes in ua159 than we did in smu21.
When comparing two types, ua159 and smu21, we found more differentially
expressed genes under pH 5 than we found under pH 7.

We used 0.01 as the q-value cutoff to have two sets of differentially expressed genes
and not ones. And also divide genes unique to SMU21 and not one. Test the
independence of differential expression and SMU21 unique genes using Fisher's
exact test. We did not find evidence that the differences were due to genes
unique to SMU21. However, we found evidence that differences were due to genes
unique to UA159. The proportion of genes unique to UA159 was smaller than that
of common genes. This shows that genes unique to SMU21 have similar effects to
common genes under pH 5, but genes unique to UA159 does not have those effects.

<<options,results=hide>>=
ph.qval.cutoff <- 10^-10
@
<<readSmutans>>=
phGenes <-
  readSmutans (countsFile=file.path(inDir, "count-ua159-bwa.txt"),
               indexFile =file.path(inDir, "count-ua159.txt.index"),
               condition = c("UA159PH7", "UA159PH5", "Smu21PH7", "Smu21PH5"),
               firstFactor = c("UA159PH7", "UA159PH5"),
               firstFactorLabel = c("ua159", "smu21"),
               secondFactor = c("UA159PH7", "Smu21PH7"),
               secondFactorLabel = c("ph7","ph5"),
               name="S. mutans UA159 pH7 and pH5")
@

\begin{figure}
\centering
\includegraphics[width=.9\textwidth]{smutans-figClustpH}
\caption{Heatmap showing the Euclidean distances between the samples
as calculated from the variance-stabilizing transformation of the
count data.}
\label{figClustpH}
\end{figure}
<<figClustpH,fig=TRUE>>=
allSamplepH <- newSmutans( phGenes )
smutans.de2Clust( allSamplepH )
@

<<ua159ph>>=
ua159ph <- newSmutans( phGenes, title="UA159" )
ua159ph <- smutans.de2( ua159ph, type="ua159", 
                        condA="ph7", condB="ph5" )
@

<<eval=TRUE>>=
# not run
smutans.de2List( ua159ph, "ua159ph.csv" )
@

<<smu21ph>>=
smu21ph <- newSmutans( phGenes, title="SMU21" )
smu21ph <- smutans.de2( smu21ph, type="smu21", 
                        condA="ph7", condB="ph5" )
@

<<eval=TRUE>>=
# not run
smutans.de2List( smu21ph, "smu21ph.csv" )
@

<<>>=
tabUA159SMU21 <- smutans.compareResult( ua159ph, smu21ph, qval=ph.qval.cutoff )
addmargins( tabUA159SMU21 )
@

<<ph7only>>=
ph7only <- newSmutans( phGenes, title="pH 7" )
ph7only <- smutans.de2( ph7only, condition="ph7", 
                        condA="ua159", condB="smu21" )
ph5only <- newSmutans( phGenes, title="pH 5" )
ph5only <- smutans.de2( ph5only, condition="ph5", 
                        condA="ua159", condB="smu21" )
tabpH7pH5 <- smutans.compareResult( ph7only, ph5only, qval=ph.qval.cutoff )

addmargins( tabpH7pH5 )
@

<<eval=TRUE>>=
# not run
smutans.de2List( ph7only, "ph7only.csv" )
smutans.de2List( ph5only, "ph5only.csv" )
@

Using all of the samples.

<<allSamplesSmu21,results=hide>>=
allSampleSmu21 <- newSmutans( phGenes, 
                              title="All Samples for Comparing pH" )
allSampleSmu21 <- smutans.de2( allSampleSmu21,
                               condA="ph7", condB="ph5" )
tabUA159pHAndAllSampleSmu21 <- smutans.compareResult( ua159ph, 
                                                      allSampleSmu21,
                                                      qval=ph.qval.cutoff )
tabSMU21pHAndAllSampleSmu21 <- smutans.compareResult( smu21ph, 
                                                      allSampleSmu21,
                                                      qval=ph.qval.cutoff )

<<>>=
addmargins( tabUA159pHAndAllSampleSmu21 )
addmargins( tabSMU21pHAndAllSampleSmu21 )
@

<<eval=TRUE>>=
# not run
smutans.de2List( allSampleSmu21, "allSampleSmu21.csv" )
@

<<allSamplesph,results=hide>>=
allSampleph <- newSmutans( phGenes, 
                           title="All Samples for Comparing UA159 and SMU21" )
allSampleph <- smutans.de2( allSampleph, compareCondition="no",
                            condA="ua159", condB="smu21" )
tabpH7AndAllSampleph <- smutans.compareResult( ph7only, 
                                               allSampleph, 
                                               qval=ph.qval.cutoff )
tabpH5AndAllSampleph <- smutans.compareResult( ph5only, 
                                               allSampleph, 
                                               qval=ph.qval.cutoff )
@

<<>>=
addmargins( tabpH7AndAllSampleph )
addmargins( tabpH5AndAllSampleph )
@

<<eval=TRUE>>=
# not run
smutans.de2List( allSampleph, "allSampleph.csv" )
@

%-----------------------------------------------------------------------------
\begin{table}[!h]
\caption{{\bf Numbers of differentially expressed genes in acid tolerance
analyses.} We use multiple testing adjusted p-value of $\Sexpr{ph.qval.cutoff}$.}
\vspace{1ex}
\noindent \begin{centering}
\begin{tabular}{ccccl}
\hline 
Sample A & Sample B & Number of Genes & Result File \\
\hline 
UA159 pH 7 & UA159 pH 5 & \Sexpr{sum(ua159ph@padj < ph.qval.cutoff,na.rm=T)} & ua159ph.csv \\
SMU21 pH 7 & SMU21 pH 5 & \Sexpr{sum(smu21ph@padj < ph.qval.cutoff,na.rm=T)} & smu21ph.csv \\
UA159 pH 7 & SMU21 pH 7 & \Sexpr{sum(ph7only@padj < ph.qval.cutoff,na.rm=T)} & ph7only.csv \\
UA159 pH 5 & SMU21 pH 5 & \Sexpr{sum(ph5only@padj < ph.qval.cutoff,na.rm=T)} & ph5only.csv \\
\hline 
\end{tabular}
\par\end{centering}
\label{tab:deph}
\end{table}
%-----------------------------------------------------------------------------

Use SMU21 contigs as a reference genome.

<<readSmutans>>=
smu21phGenes <-
  readSmutans (countsFile=file.path(inDir, "count-smu21-bwa.txt"),
                indexFile =file.path(inDir, "count-smu21.txt.index"),
                condition = c("Smu21PH7", "Smu21PH5"),
                firstFactor = c(),
                firstFactorLabel = c("smu21"),
                secondFactor = c(),
                secondFactorLabel = c("ph7", "ph5"),
                name="S. mutans UA159 pH7 and pH5 on SMU21")
@

<<smu21phSMU21>>=
smu21phSMU21 <- newSmutans( smu21phGenes, title="SMU21" )
smu21phSMU21 <- smutans.de2( smu21phSMU21, type="smu21", 
                            condA="ph7", condB="ph5" )
@

<<eval=TRUE>>=
# not run
smutans.de2List( smu21phSMU21, "smu21phSMU21.csv" )
@

<<eval=TRUE>>=
x <- scan(file.path(inDir, "SMU21.diff.lst"), what="character")
y <- smutans.de2List( smu21phSMU21, qval=10^-2)$id
z <- smu21phSMU21@res$id
x1 <- z[z %in% x]
x2 <- z[!z %in% x]
y1 <- z[z %in% y]
y2 <- z[!z %in% y]
sum(x1 %in% y1)
sum(x1 %in% y2)
sum(x2 %in% y1)
sum(x2 %in% y2)
unique.diff <-
matrix(c(7, 67, 106, 1562),
       nrow = 2,
       dimnames =
       list(c("Diff", "Not diff"),
            c("Unique", "Common")))
fisher.test(unique.diff)
@

<<>>=
x <- scan(file.path(inDir, "UA159-SMU21.diff.lst"), what="character")
y <- smutans.de2List( smu21ph, qval=10^-2)$id
z <- smu21ph@res$id
x1 <- z[z %in% x]
x2 <- z[!z %in% x]
y1 <- z[z %in% y]
y2 <- z[!z %in% y]
sum(x1 %in% y1)
sum(x1 %in% y2)
sum(x2 %in% y1)
sum(x2 %in% y2)
unique.diff <-
matrix(c(2, 267, 107, 1575),
       nrow = 2,
       dimnames =
       list(c("Diff", "Not diff"),
            c("Unique", "Common")))
fisher.test(unique.diff)
@

%--------------------------------------------------
\section{DE analysis using predicted transcripts}
\label{sec:detranscript}
%--------------------------------------------------

We need ways of supporting the predicted transcripts. We need a way of using
expression levels and counts of starts and ends of short reads to predict
transcripts. For now, we just use an ad hoc method for finding transcript start
and end sites. A likelihood can be constructed by multiplying geometric
probabilities. Given a range of possible transcription boundaries we can find
estimates of positions and geometirc parmeter $p$ that maximize the likelihood
for both ends of a transcript.

\Rpackage{GenomicRanges} is useful in summarizing predicted transcripts.
feature-genome.out-geneonly is stored in \Robject{smutans.feature.genes}.  
File FASTQ001.parsernaseq1 is ParseRNAseq output file for transcript
predictions. File FASTQ001.parsernaseq.pileup is ParseRNAseq input file for
site-wise expression levels.

\Rfunction{smutans.prepareTranscript} creates two objects:
\Robject{smutans.tx} and \Robject{smutans.txPileup}.
<<tx,eval=FALSE>>=
source("http://www.bioconductor.org/biocLite.R")
biocLite("GenomicRanges")
@

%--------------------------------------------------
\section{DE analysis using predicted small RNAs and target genes}
\label{sec:desmallrna}
%--------------------------------------------------

Create a file with intergenic regions sequences of \textit{S.\ mutans} UA159.
Create a file with bacterial genome sequences except \textit{S.\ mutans} NN2025.
BLAST the intergenic regions against the bacterial genomes. Extract sequence
alignments from the BLAST output. Refine the alignment using MUSCLE. Convert the
FASTA-format alignments to MSA-format ones.  Find small RNA candidates using
RNAz.  Find targets of the small RNA candidates using RNAplex. Targets are
scored for each small RNA candidate. The scores were used to do functional
category enrichment tests of targets for each small RNA candidate.

Table \ref{tab:goseq} 
shows functional categories enriched with target genes for predicted small
RNAs. Let's use \Rfunction{smutans.mannwhitney} to the package and other data
objects.
%-----------------------------------------------------------------------------
\begin{table}[!h]
\caption{{\bf Gene Ontology (GO) enrichments for differentially expressed
genes}}
\vspace{1ex}
\noindent \begin{centering}
\begin{tabular}{lcccl}
\hline 
small RNA & $q^a$ & count & GO & Description \\
\hline 
%-----------------------------------------------------------------------------
<<goseqUA159only,echo=FALSE,results=tex>>=
data( "smutans.genes.criteria" )
data( "smutans.go.genes" )
data( "smutans.cat.desc" )
smutans.mannwhitney( qval=0.05,
                   genes.criteria=smutans.genes.criteria,
                   go.genes=smutans.go.genes,
                   cat.desc=smutans.cat.desc )
@
%-----------------------------------------------------------------------------
\hline 
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$False discovery rate estimated by the
Benjamini-Hochberg method.  All categories having at least ten genes and
$q\leq 0.05$ are displayed.\\ 
\end{flushleft}
\label{tab:goseq}
\end{table}
%-----------------------------------------------------------------------------

<<smutansRNA>>=
data( "smutansGenes2" )
ua159only <- newSmutans( smutansGenes2, title="UA159 Only" )
ua159only <- smutans.de2( ua159only, type="ua159", 
                          condA="glucose", condB="galactose" )
allSample <- newSmutans( smutansGenes2, 
                         title="All Samples for Comparing Conditions" )
allSample <- smutans.de2( allSample, 
                          condA="glucose", condB="galactose" )
tabUA159onlyAndAllSample <- smutans.compareResult( ua159only, 
                                                   allSample )
addmargins( tabUA159onlyAndAllSample )
@
Of the total \Sexpr{sum(tabUA159onlyAndAllSample)} gene,
\Sexpr{tabUA159onlyAndAllSample[2,2]} genes were found in both analyses. The
analysis using all of the sample found \Sexpr{tabUA159onlyAndAllSample[1,2]}
genes. The analysis using ua159-only sample found \Sexpr{tabUA159onlyAndAllSample[2,1]}
genes. We found a few predicted small RNAs that appeared to be differentially
expressed.
<<smutansRNA>>=
smutans.de2List( allSample )[grep("locus",rownames(smutans.de2List(allSample))),]
@
We also found a few predicted small RNAs that appeared to be differentially
expressed in the comparison of the two types: ``ua159'' and ``tw1.''
<<smutansRNA>>=
allSampleType <- newSmutans( smutansGenes2, title="All samples for type" )
allSampleType <- smutans.de2( allSampleType, compareCondition="no",
                              condA="ua159", condB="tw1" )
smutans.de2List( allSampleType )[grep("locus",rownames(smutans.de2List(allSampleType))),]
@

%--------------------------------------------------
\section{Miscellaneous R scripts}
%--------------------------------------------------

Because \Rpackage{DESeq} recommended to check
the dispersion estimates and to make sure that the data quality is as
expected, we plot the per-gene estimates against the normalized mean
expressions per gene, and then overlay the fitted curve in red color
(Fig.\ \ref{figFit}).

\begin{figure}
\centering
\includegraphics[width=.6\textwidth]{smutans-figFit}
\caption{Empirical (black dots) and fitted (red lines) dispersion values
plotted against mean expression strength for the data set of UA159-only sample
to compare glucose and galactose conditions.}
\label{figFit}
\end{figure}
<<figFit,fig=TRUE>>=
smutans.plotDispersionEstimates(ua159only)
@
\fixme{The following paragraph is adopted from \Rpackage{DESeq} vignette.}
As we estimated the dispersion from only two samples, we should expect the
estimates to scatter with quite some sampling variance around their true values.
Hence, \Rpackage{DESeq} should not use the per-gene estimates directly in the
test, because using too low dispersion values leads to false positives.  Many of
the values below the red line are likely to be underestimates of the true
dispersions, and hence, it is prudent to instead rather use the fitted value.
On the othe hand, not all of the values above the red line are overestimations,
and hence, the conservative choice is to keep them instead of replacing them
with their fitted values.
Having estimated the dispersion for each gene, we look for differentially
expressed genes. To contrast two conditions, e.g., to see whether there is
differential expression between conditions ``glucose'' and ``galactose'',
\Rfunction{smutans.de2} already performed the 
\Rfunction{nbinomTest} provided by \Rpackage{DESeq}.  
Let's list genes with resulting information from \Rpackage{DESeq}.

%-----------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=.6\textwidth]{smutans-figDE}
\caption{Plot of normalised mean versus $log_2$ fold change
(this plot is sometimes also called the ``MA-plot'') for the contrast ``glucose''
versus ``galactose'' using UA159-only sample.}
\label{figDE}
\end{figure}
%-----------------------------------------------------------------------------
<<figDE,fig=TRUE>>=
smutans.plotDiffExp(ua159only)
@
Figure \ref{figDE} shows a plot of the $\log_2$ fold changes against the base
means, coloring in red those genes that are significant at 10\% of FDR.

We could check 
the dispersion estimates and to make sure that the data quality is as
expected by plotting the per-gene estimates against the normalized mean
expressions per gene, and then overlay the fitted curve.
<<figFitAllSample,eval=FALSE>>=
# not run
smutans.plotDispersionEstimates(allSample)
@
<<figCompareUA159onlyAndAllSample,fig=TRUE>>=
pval.ua159only <- smutans.pval( ua159only )
pval.allSample <- smutans.pval( allSample )
bottom = function(x, theta=1e-12) pmax(x, theta)
plot( bottom(pval.ua159only), bottom(pval.allSample), log="xy", pch=20, cex=.3 )
abline(a=0, b=1, col="blue")
@
Figure \ref{figCompareUA159onlyAndAllSample} shows the raw p-values from
the two-factor analysis versus those from the single-factor analysis. The
p-values from the two-factor analysis are overall smaller than those from the
single-factor analysis. We would expect that this was the case because of higher
power of the two-factor analysis.
%-----------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=.6\textwidth]{smutans-figCompareUA159onlyAndAllSample}
\caption{Comparison of $p$ values (unadjusted) from the test using only the
four paired-end samples and the test using all seven samples. We can see that
the latter tend to be smaller, as expected from the higher power of the test with all samples.}
\label{figCompareUA159onlyAndAllSample}
\end{figure}
%-----------------------------------------------------------------------------

Once we find differentially expressed genes, we could cluster them. This could
be useful to group genes that behave similarly.  These potentially useful
figures can be plotted from \Rpackage{DESeq}.  We could also find function
categories that were enriched with differentially expressed genes. For
clustering differentially expressed genes we need moderate fold changes to make
them amenable to distance calculation.  Figure \ref{figLfc} shows moderated log
fold changes versus the original log fold changes.

<<options,results=hide,echo=FALSE>>=
largeNumber <- 10
@

\begin{figure}
\centering
\includegraphics[width=.6\textwidth]{smutans-figLfc}
\caption{Scatterplot of direct (\Robject{lfc}) versus \emph{moderated} log-ratios (\Robject{mod\_lfc}).
  The moderation criterion used is variance stabilisation.
  The red points correspond to values that were infinite in \Robject{lfc}
  and were abitrarily set to $\pm\Sexpr{largeNumber}$ for the purpose of plotting.
  These values vary in a finite range (as shown in the plot) in \Robject{mod\_lfc}.}
\label{figLfc}
\end{figure}
<<figLfc,fig=TRUE>>=
select <- 1:40
smutans.de2TypeLfc( ua159only, condA="glucose", condB="galactose" )
@
Now we plot the first \Sexpr{length(select)} genes clustered based on the
moderated log fold changes.

%-----------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=.6\textwidth]{smutans-figHeatmap1}
\caption{Heatmaps showing the expression data of the top \Sexpr{length(select)}
  differentially expressed genes. Left, the variance stabilization
  transformed data, right, the original count data.
  The right plot is dominated by a small number of data points with large values.}
\label{figHeatmap1}
\end{figure}
%-----------------------------------------------------------------------------
<<figHeatmap1,fig=TRUE>>=
smutans.de2TypeHeatmap( ua159only )
@

\begin{figure}
\centering
\includegraphics[width=.9\textwidth]{smutans-figHeatmapNpAndP}
\caption{Empirical (black dots) and fitted (red lines) dispersion values
plotted against mean expression strength. \textbf{figHeatmapNpAndP}}
\label{figHeatmapNpAndP}
\end{figure}
<<figHeatmapNpAndP,fig=TRUE>>=
smutans.de2TypeHeatmap( npAndP )
@

\section{Session Info}
<<sessi>>=
sessionInfo()
@

\bibliographystyle{unsrt}
\bibliography{siepel-mutans}

\end{document}