### 2012-04-24 package/tutorial.Rnw
Just an update.

%\VignetteIndexEntry{Analyzing Streptococcus mutans RNA-seq data}
%\VignettePackage{smutans}

% To compile this document
% library('weaver'); rm(list=ls()); Sweave('smutans.Rnw', driver=weaver()); system('pdflatex smutans')

\documentclass{article}

\usepackage{Sweave}
\usepackage[letterpaper]{geometry}
\usepackage{hyperref,graphicx}
\usepackage{color}
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

\SweaveOpts{keep.source=TRUE,eps=FALSE,include=FALSE,width=4,height=4.5}
\newcommand{\Robject}[1]{object \texttt{#1}}
\newcommand{\Rpackage}[1]{R package \textit{#1}}
\newcommand{\Rclass}[1]{class \textit{#1}}
\newcommand{\Rfunction}[1]{{function \small\texttt{#1}}}
\newcommand{\program}[1]{{\texttt{#1}}}
\newcommand{\fixme}[1]{{\textbf{Fixme:} \textit{\textcolor{blue}{#1}}}}

\renewcommand{\floatpagefraction}{0.7}

\author{Sang Chul Choi\\[1em]Cornell University,\\ Ithaca, NY\\[1em]
\texttt{schoi@cornell.edu}}

\title{\textsf{\textbf{Analyzing \textit{Streptococcus mutans} RNA-seq data\\using R packages}}}

\begin{document}

\maketitle

\begin{abstract}
We analyze RNA samples from \textit{Streptococcus mutans}.  We use
RNA-seq to define genes whose expression changes between samples
\cite{Oshlack2010}. Short reads are first mapped to a reference genome using
approaches such as \program{BWA} \cite{Li2009a}.  Reads fitting into one of the annotated
intervals are counted.  These counts are passed to a statistical package which
treats differences in read counts for statistical significance. For this
purpose, we made use of \Rpackage{DEseq} \cite{Anders2010} or \Rpackage{edgeR}
\cite{Robinson2010}. This prodecure returns a list of differentially expressed
genes, which can be passed to downstream analysis, such as functional category
enrichment test using \Rpackage{goseq} \cite{Young2010}.  The current
\Rpackage{smutans} contains procedures that use functions provided by
the above R packages. 
\end{abstract}

\tableofcontents

%--------------------------------------------------
\section{Data preparations} \label{sec:prep}
%--------------------------------------------------

Steps of analyzing RNA-seq data for determining differentially expressed genes
are described in \cite{Oshlack2010}. We used a short read aligner called
\program{BWA} \cite{Li2009a} to map the sequenced short reads on a reference genome,
\textit{Streptococcus mutans} UA159 (NCBI accession NC\_004350). Because the
aligner allowed a few gaps for efficient alignment of millions of sequences of
size being 100 base pairs long, short reads that consist of RNAs and sequencing
adapter sequences would be not mapped. We used \program{cutadapt} \cite{Martin2011} to
remove parts of adapter sequences from short reads before mapping on the
reference genome. Mapped short read alignments were converted into readable
formats \cite{Li2009b}. We counted short reads aligned to annotated genes in the
reference genome, forming a table of read counts of the genes. Statistical
software R \cite{R2011} with \Rpackage{DEseq} \cite{Anders2010}
was subsequently employed to infer differentially expressed in different
biological conditions.  We also generated tracks for our recently released
Streptococcus Genome Browser (http://strep-genome.bscb.cornell.edu) that
summarize gene expression levels based the short read alignments. This allowed
us to visualize parts of the reference genome for comparing different RNA
samples. Further analyses such as functional category enrichment test
\cite{Young2010} were also employed.

The \Rpackage{smutans} expects count data in the form of \Rclass{CountDataSet},
which is a type provided by \Rpackage{DESeq}.  Using \program{BWA} \cite{Li2009a} and
\program{samtools} \cite{Li2009b} we aligned short reads from RNA sequencing. Using gene
annotations in the UA159 genome we counted short reads that overlapped the
genes.  The count values are raw counts of sequencing reads. The
\Rpackage{DESeq}'s statistical model, or negative binomial model, hold for raw
reads that allows to assess the measurement precision correctly; we did not have
rounded values of normalized counts, or counts of covered base pairs.  Several
exeriments of \textit{S.\ mutans} RNA sampling were performed.  \fixme{Find the
papers that are related with the experiments.} One of them was presented in the
\Robject{smutansGenes} as a part of \Rpackage{smutans}. The
\Robject{smutansGenes} is of \Rclass{CountDataSet}, which is the data type used
by \Rpackage{DESeq}. We have a convenient \Rclass{Smutans}, which contains slots
for \Rclass{CountDataSet} objects, and results from analyses thereof.  Before
loading necessary R packages and data sets, let us describe what functions were
used for preparing data objects such as \Robject{smutansGenes}. One should
create an R script file at the directory that contains the source code directory
of the \Rpackage{smutans}. Put the following scripts in the R script file and
run the script at that directory. That should create RData files at the
directory smutans/data. User should not run the R script because the data files
should be found in the \Rpackage{smutans}.

% http://www.oralgen.lanl.gov/oralgen/downloads/smut.csv
% http://www.oralgen.lanl.gov/cgi-bin/gene_id_search.cgi?dbname=smut&gene_id=SMu0001
% \fixme{Where are flat files for genes of S. mutans UA159?}
<<samplePreparation,eval=FALSE>>=
# not run
# but create a R scirpt file with the following scripts, and
# run it, i.e., Rscript run.R
library(DESeq)
source("smutans/R/core.R")
smutans.prepareData34( )
smutans.prepareData835NPP( )
smutans.prepareDataUA159TW1( ) 
smutans.prepareGoseq( )
smutans.prepareDataOMZ175( ) 
smutans.prepareDataUA159CSP( ) 
smutans.prepareDataSMU86CSP( )
q("no")
@
<<options,results=hide,echo=FALSE>>=
options(digits=3)
@
Now, see some of the data objects by loading the \Rpackage{smutans}, and a data
\Robject{smutansGenes}. Note that the object is of type \Rclass{CountDataSet} of
\Rpackage{DESeq} in which users should find \Rfunction{counts} and
\Rfunction{pData}. 
<<smutans,results=hide>>=
library( "smutans" )
data( "smutansGenes" )
@
There are 
\Sexpr{length(colnames(counts(smutansGenes)))}
RNA-seq samples in \Robject{smutansGenes}, which
contains the counts and also metadata about the samples:
<<look_into_smutansGenes>>=
head( counts(smutansGenes) )
@
<<look_into_smutansGenes>>=
pData( smutansGenes )
@
The samples differ by experimental condition (``glucose'' or
``galactose'') and by library type (``ua159'' or ``tw1'').

Samples of 835NP, 835P, and UA159 no CSP should be loaded using the following
command.
<<sm835,results=hide>>=
data( "sm835Genes" )
@

% Resources
% Google charts - Venn Diagram
% http://code.google.com/apis/chart/index.html

\fixme{Describe other data objects as well.
sm34Genes.RData,
sm835Genes.RData,
smutans.cat.desc.RData,
smutans.feature.genes.RData, and
smutans.go.genes.RData.}

%--------------------------------------------------
\section{DE analysis using UA159 and TW1 samples with
glucose and galactose} \label{sec:detw1}
%--------------------------------------------------

We could compare four pairwise comparisons using the two-factor data set of
\Robject{smutansGenes}: a pair of comparisons between ``glucose'' and
``galactose'' given either type ``ua159'' or ``tw1'', and another pair of
comparisons between types ``ua159'' and ``tw1'' given either condition
``glucose'' or ``galactose.'' We focus on one of the four pairwise comparisons
first. Later, we will perform analyses for the rest of the three comparisons.
We look at the ``ua159'' data for now, and will look at the ``tw1'' data later.

We can cluster samples using variance stabilized data. Figure \ref{figClust}
shows the \Sexpr{length(colnames(counts(smutansGenes)))} sampled clustered. 
The clustering should reflect our experimental design, i.e., samples are
more similar when they have the same treatment or the same library type.
To make this conclusion, we estimate the dispersion with
mode ``blind'' (see \Rpackage{DESeq} vignette), as only then,
the variance stabilizing transformation is not informed about the design and
hence not biased towards a result supporting it. Such an analysis is useful
for quality control.  This appears to be useful as a first in any analysis.

\begin{figure}
\centering
\includegraphics[width=.9\textwidth]{smutans-figClust}
\caption{Heatmap showing the Euclidean distances between the samples
  as calculated from the variance-stabilizing transformation of the
  count data.}
\label{figClust}
\end{figure}
<<figClust,fig=TRUE>>=
allSample <- newSmutans( smutansGenes )
smutans.de2Clust( allSample )
@
As expected, two experimental conditions of ``glucose'' and ``galactose'' divide the
samples. Within each condition, two different types of ``ua159'' and ``tw1'' divide
them again. The pair of ``glucose ua159'' and ``galactose tw1'' is the most
distinguished.

%------------------------------------------------------------
\subsection{Inference using single-factor data} 
\label{subsec:ua159only}
%------------------------------------------------------------

We create a \Rclass{Smutans} object using the \Rclass{CountDataSet}
\Robject{smutansGenes}. The title option would be useful in distinguishing
different objects of \Rclass{Smutans} as we have more R objects with the type.  
A subsequent call to
\Rfunction{smutans.de2} uses a two-factor data set to test differential
expression of genes.  The \Rfunction{smutans.de2} estimates effective library
sizes, and dispersions using functions provided by \Rpackage{DESeq}. Then, it
tests genes for differential expression.
<<ua159onlySamples>>=
ua159only <- newSmutans( smutansGenes, title="UA159 Only" )
ua159only <- smutans.de2( ua159only, type="ua159", 
                          condA="glucose", condB="galactose" )
@
Because \Rpackage{DESeq} recommended to check
the dispersion estimates and to make sure that the data quality is as
expected, we plot the per-gene estimates against the normalized mean
expressions per gene, and then overlay the fitted curve in red color
(Fig.\ \ref{figFit}).

\begin{figure}
\centering
\includegraphics[width=.6\textwidth]{smutans-figFit}
\caption{Empirical (black dots) and fitted (red lines) dispersion values
plotted against mean expression strength for the data set of UA159-only sample
to compare glucose and galactose conditions.}
\label{figFit}
\end{figure}
<<figFit,fig=TRUE>>=
smutans.plotDispersionEstimates(ua159only)
@
\fixme{The following paragraph is adopted from \Rpackage{DESeq} vignette.}
As we estimated the dispersion from only two samples, we should expect the
estimates to scatter with quite some sampling variance around their true values.
Hence, \Rpackage{DESeq} should not use the per-gene estimates directly in the
test, because using too low dispersion values leads to false positives.  Many of
the values below the red line are likely to be underestimates of the true
dispersions, and hence, it is prudent to instead rather use the fitted value.
On the othe hand, not all of the values above the red line are overestimations,
and hence, the conservative choice is to keep them instead of replacing them
with their fitted values.
Having estimated the dispersion for each gene, we look for differentially
expressed genes. To contrast two conditions, e.g., to see whether there is
differential expression between conditions ``glucose'' and ``galactose'',
\Rfunction{smutans.de2} already performed the 
\Rfunction{nbinomTest} provided by \Rpackage{DESeq}.  
Let's list genes with resulting information from \Rpackage{DESeq}.
<<ua159onlyDE2List>>=
head(smutans.de2List( ua159only ))
@
For each gene, we get its mean expression level (at the base scale) as a joint
estimate from both conditions, and estimated separately for each condition, the
fold change from the first to the second condition, i.e., 
$\frac{baseMeanB}{baseMeanA}$, the logarithm (to basis 2)
of the fold change, and the $p$ value for the statistical significance of this
change. The \texttt{padj} column contains the $p$ values, adjusted for multiple
testing with the Benjamini-Hochberg procedure (see the R function
\Rfunction{p.adjust}), which controls false discovery rate (FDR). Later, we
print the gene list out in a CSV file using option file of
\Rfunction{smutans.de2List}.
%-----------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=.6\textwidth]{smutans-figDE}
\caption{Plot of normalised mean versus $log_2$ fold change
(this plot is sometimes also called the ``MA-plot'') for the contrast ``glucose''
versus ``galactose'' using UA159-only sample.}
\label{figDE}
\end{figure}
%-----------------------------------------------------------------------------
<<figDE,fig=TRUE>>=
smutans.plotDiffExp(ua159only)
@
Figure \ref{figDE} shows a plot of the $\log_2$ fold changes against the base
means, coloring in red those genes that are significant at 10\% of FDR. 
Let us use ``tw1'' type to compare ``glucose'' and ``galactose'' conditions.
We continue to analyze the data set using type ``tw1'' to compare the two
conditions. We perform the single-factor analysis. 
<<tw1onlySamples>>=
tw1only <- newSmutans( smutansGenes, title="TW1 Only" )
tw1only <- smutans.de2( tw1only, type="tw1", 
                        condA="glucose", condB="galactose" )
tabUA159onlyTW1only <- smutans.compareResult( ua159only, tw1only )
addmargins( tabUA159onlyTW1only )
@
The table, \Robject{tabUA159onlyTW1only}, shows numbers of genes found to be 
differentially expressed in the two analyses: ``UA159 Only'' and ``TW1 Only.''
Of the total \Sexpr{sum(tabUA159onlyTW1only)} genes, \Sexpr{tabUA159onlyTW1only[2,2]}
genes were found in both analyses. The analysis using TW1-only sample found
\Sexpr{tabUA159onlyTW1only[1,2]} genes. The analysis using UA159-only sample found
\Sexpr{tabUA159onlyTW1only[2,1]} genes.

Until now we focused on the effect of condition either using UA159-only or
TW1-only samples. We now change the gear to see effects of sample types on
differentially expressed genes. We create two \Rclass{Smutans} objects:
\Robject{glucoseOnly} and \Robject{galactoseOnly}, and compare them in detecting
differentially expressed genes:
<<glucoseOnlySamples>>=
glucoseOnly <- newSmutans( smutansGenes, title="Glucose Only" )
glucoseOnly <- smutans.de2( glucoseOnly, condition="glucose", 
                            condA="ua159", condB="tw1" )
galactoseOnly <- newSmutans( smutansGenes, title="Galactose Only" )
galactoseOnly <- smutans.de2( galactoseOnly, condition="galactose", 
                              condA="ua159", condB="tw1" )
tabGlucoseOnlyGalactoseOnly <- smutans.compareResult( glucoseOnly, 
                                                      galactoseOnly )
addmargins( tabGlucoseOnlyGalactoseOnly )
@
Of the total \Sexpr{sum(tabGlucoseOnlyGalactoseOnly)} genes,
\Sexpr{tabGlucoseOnlyGalactoseOnly[2,2]} genes were found in both analyses at FDR
rate of 10\%. The analysis using galactose-only sample found
\Sexpr{tabGlucoseOnlyGalactoseOnly[1,2]} genes. The analysis using glucose-only
sample found \Sexpr{tabGlucoseOnlyGalactoseOnly[2,1]} genes. 
Table \ref{tab:de} summarizes the number of differentially expressed genes that
were found using single-factor analyses at FDR of 10\%.

%-----------------------------------------------------------------------------
\begin{table}[!h]
\caption{{\bf Numbers of differentially expressed genes in different analyses}}
\vspace{1ex}
\noindent \begin{centering}
\begin{tabular}{ccccl}
\hline 
Sample A & Sample B & Number of Genes & Result File \\
\hline 
UA159 Glucose & UA159 Galactose & \Sexpr{sum(ua159only@padj < .1,na.rm=T)} & ua159only.csv \\
TW1 Glucose & TW1 Galactose & \Sexpr{sum(tw1only@padj < .1,na.rm=T)} & tw1only.csv \\
UA159 Glucose & TW1 Glucose & \Sexpr{sum(glucoseOnly@padj < .1,na.rm=T)} & glucoseOnly.csv \\
UA159 Galactose & TW1 Galactose & \Sexpr{sum(galactoseOnly@padj < .1,na.rm=T)} & galactoseOnly.csv \\
% UA159 Glucose & UA159 Galactose & \Sexpr{sum(tabUA159onlyTW1only[2,])} & ua159only.csv \\
% TW1 Glucose & TW1 Galactose & \Sexpr{sum(tabUA159onlyTW1only[,2])} & tw1only.csv \\
% UA159 Glucose & TW1 Glucose & \Sexpr{sum(tabGlucoseOnlyGalactoseOnly[2,])} & glucoseOnly.csv \\
% UA159 Galactose & TW1 Galactose & \Sexpr{sum(tabGlucoseOnlyGalactoseOnly[,2])} & galactoseOnly.csv \\
\hline 
\end{tabular}
\par\end{centering}
\label{tab:de}
\end{table}
%-----------------------------------------------------------------------------

%------------------------------------------------------------
\subsection{Inference using all of the samples} 
\label{subsec:allSample}
%------------------------------------------------------------

It is inconvenient to use four different analyses to find differentially
expressed genes under one of two factors. Besides, it would be desirable to use all of
the data to find differentially expressed genes because using more samples might
be more powerful in detecting differentially expressed genes.  Let us find genes
differentially expressed in different conditions of ``glucose'' and
``galactose'' using both types of ``ua159'' and ``tw1.'' We will compare these
genes and those found in the single-factor analysis that we did previously,
i.e., \Robject{ua159only}.
Let's start with multi-factor the analysis by creating a new \Rclass{Smutans} object, called
\Robject{allSample}. Although we already used \Robject{allSample} when
clustering the samples, we recreate it for instructive purposes.
<<allSamples,results=hide>>=
allSample <- newSmutans( smutansGenes, 
                         title="All Samples for Comparing Conditions" )
allSample <- smutans.de2( allSample )
@
The \Rfunction{smutans.de2} uses a two-factor data set to
perform two-factor analysis. The two factors are type and condition. We are here
interested in condition effect. We could check 
the dispersion estimates and to make sure that the data quality is as
expected by plotting the per-gene estimates against the normalized mean
expressions per gene, and then overlay the fitted curve.
<<figFitAllSample,eval=FALSE>>=
# not run
smutans.plotDispersionEstimates(allSample)
@
%-----------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=.6\textwidth]{smutans-figCompareUA159onlyAndAllSample}
\caption{Comparison of $p$ values (unadjusted) from the test using only the
four paired-end samples and the test using all seven samples. We can see that
the latter tend to be smaller, as expected from the higher power of the test with all samples.}
\label{figCompareUA159onlyAndAllSample}
\end{figure}
%-----------------------------------------------------------------------------
<<figCompareUA159onlyAndAllSample,fig=TRUE>>=
pval.ua159only <- smutans.pval( ua159only )
pval.allSample <- smutans.pval( allSample )
bottom = function(x, theta=1e-12) pmax(x, theta)
plot( bottom(pval.ua159only), bottom(pval.allSample), log="xy", pch=20, cex=.3 )
abline(a=0, b=1, col="blue")
@
<<compareUA159onlyAndAllSample>>=
tabUA159onlyAndAllSample <- smutans.compareResult( ua159only, 
                                                   allSample )
addmargins( tabUA159onlyAndAllSample )
@
Of the total \Sexpr{sum(tabUA159onlyAndAllSample)} gene,
\Sexpr{tabUA159onlyAndAllSample[2,2]} genes were found in both analyses. The
analysis using all of the sample found \Sexpr{tabUA159onlyAndAllSample[1,2]}
genes. The analysis using ua159-only sample found \Sexpr{tabUA159onlyAndAllSample[2,1]}
genes.  Figure \ref{figCompareUA159onlyAndAllSample} shows the raw p-values from
the two-factor analysis versus those from the single-factor analysis. The
p-values from the two-factor analysis are overall smaller than those from the
single-factor analysis. We would expect that this was the case because of higher
power of the two-factor analysis.

We can show that taking the library type into account was important to have
good detection power by doing the analysis again using the standard workflow,
as outlined earlier, and without informing \Rpackage{DESeq} of the library
types. This anlaysis is not necessary. However, it clearly shows the advantage of the
two-factor analysis.
<<allSampleWithoutType>>=
allSampleWithoutType <- newSmutans( smutansGenes, "All Sample without Type" )
allSampleWithoutType <- smutans.de2WithoutType( allSampleWithoutType,
                                                condA="glucose", 
                                                condB="galactose" )
tabWithoutWithAllSample <- smutans.compareResult( allSampleWithoutType, 
                                                  allSample )
addmargins( tabWithoutWithAllSample )
@
Of the total \Sexpr{sum(tabWithoutWithAllSample)} genes,
\Sexpr{tabWithoutWithAllSample[2,2]} genes were found in both analyses. The
analysis using all of the sample found \Sexpr{tabWithoutWithAllSample[1,2]}
genes. The analysis using all of the same sample but without considering type
found only \Sexpr{tabWithoutWithAllSample[2,1]} genes.

As we mentioned above, we can have two analyses using the two-factor analysis.
Until now we compared conditions.
We proceed to compare the two types by considering all of the samples.
<<allSampleType,results=verbatim>>=
allSampleType <- newSmutans( smutansGenes, title="All samples for type" )
allSampleType <- smutans.de2( allSampleType, compareCondition="no" )
tabGlucoseOnlyAllSample <- smutans.compareResult( glucoseOnly, 
                                                  allSampleType )
addmargins( tabGlucoseOnlyAllSample )
tabGalactoseOnlyAllSample <- smutans.compareResult( galactoseOnly, 
                                                    allSampleType )
addmargins( tabGalactoseOnlyAllSample )
@
Of the total \Sexpr{sum(tabGlucoseOnlyAllSample)} genes,
\Sexpr{tabGlucoseOnlyAllSample[2,2]} genes were found in both analyses. The
analysis using all of the sample found \Sexpr{tabGlucoseOnlyAllSample[1,2]}
genes. The analysis using glucose-only sample
found \Sexpr{tabGlucoseOnlyAllSample[2,1]} genes.  
In the comparison using galactose-only sample,
of the \Sexpr{sum(tabGalactoseOnlyAllSample)} genes,
\Sexpr{tabGalactoseOnlyAllSample[2,2]} genes were found in both analyses. The
analysis using all of the sample found \Sexpr{tabGalactoseOnlyAllSample[1,2]}
genes. The analysis using galactose-only sample
found \Sexpr{tabGalactoseOnlyAllSample[2,1]} genes.

We print out the result of differentially expressed genes in CSV files. A
spreadsheet program such as EXCEL should be able to open CSV files. Users who
are more comfortable with the EXCEL program than R could post-process CSV
files for sorting and finding differentially expressed genes using different
q-value cutoff values.
<<smutanSamples>>=
smutans.de2List( allSample, "allSample.csv" )
smutans.de2List( allSampleType, "allSampleType.csv" )
smutans.de2List( glucoseOnly, "glucoseOnly.csv" )
smutans.de2List( galactoseOnly, "galactoseOnly.csv" )
smutans.de2List( ua159only, "ua159only.csv" )
smutans.de2List( tw1only, "tw1only.csv" )
head( smutans.de2List( allSample ) )
@
The output from \Robject{allSample} is different from that of
\Robject{ua159only} or \Robject{tw1only}.
The first column of ``allSample.csv'' is for genes.
The next three columns show the fitted coefficients, converted to a logarithm
base 2 scale: ``X.Intercept.'', ``typeua159'', and ``conditionglucose.''
The log2 fold change due to the condition is shown in the 4th column. As indicated
by the column name, it is the effect of ``glucose'', i.e., the log ratio of
``glucose'' versus ``galactose'' or $\log(glucose/galactose)$. 
Note that the library type also
had noticeable influence on the expression, and hence would have increased the dispersion
estimates (and so reduced power) if we had not fitted an effect for it.
The column \emph{deviance}
is the deviance of the fit. (Comparing the deviances with a $\chi^2$ likelihood ratio test is
how \Rfunction{nbinomGLMTest} calculates the $p$ values.) The column, \emph{converged},
indicates whether the calculation of coefficients and deviance has fully converged.
(If it is false too often, you can try to change the GLM control parameters, as
explained in the help page to \Rfunction{fitNbinomGLMs}.)
The second to the last column, \emph{pval}, is the raw p-values from the test
\Rfunction{nbinomGLMTest} provided by \Rpackage{DESeq}. The last column,
\emph{padj}, is the multiple testing adjusted p-values.

%------------------------------------------------------------
\subsection{Functional category association analyses} 
\label{subsec:functional}
%------------------------------------------------------------

Once we find differentially expressed genes, we could cluster them. This could
be useful to group genes that behave similarly.  These potentially useful
figures can be plotted from \Rpackage{DESeq}.  We could also find function
categories that were enriched with differentially expressed genes. For
clustering differentially expressed genes we need moderate fold changes to make
them amenable to distance calculation.  Figure \ref{figLfc} shows moderated log
fold changes versus the original log fold changes.

<<options,results=hide,echo=FALSE>>=
largeNumber <- 10
@

\begin{figure}
\centering
\includegraphics[width=.6\textwidth]{smutans-figLfc}
\caption{Scatterplot of direct (\Robject{lfc}) versus \emph{moderated} log-ratios (\Robject{mod\_lfc}).
  The moderation criterion used is variance stabilisation.
  The red points correspond to values that were infinite in \Robject{lfc}
  and were abitrarily set to $\pm\Sexpr{largeNumber}$ for the purpose of plotting.
  These values vary in a finite range (as shown in the plot) in \Robject{mod\_lfc}.}
\label{figLfc}
\end{figure}
<<figLfc,fig=TRUE>>=
select <- 1:40
smutans.de2TypeLfc( ua159only, condA="glucose", condB="galactose" )
@
Now we plot the first \Sexpr{length(select)} genes clustered based on the
moderated log fold changes.

%-----------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=.6\textwidth]{smutans-figHeatmap1}
\caption{Heatmaps showing the expression data of the top \Sexpr{length(select)}
  differentially expressed genes. Left, the variance stabilization
  transformed data, right, the original count data.
  The right plot is dominated by a small number of data points with large values.}
\label{figHeatmap1}
\end{figure}
%-----------------------------------------------------------------------------
<<figHeatmap1,fig=TRUE>>=
smutans.de2TypeHeatmap( ua159only )
@

<<goseqUA159only,eval=FALSE>>=
data( "smutans.feature.genes" )
data( "smutans.go.genes" )
data( "smutans.cat.desc" )
smutans.de2Goseq ( ua159only, qval=1e-2, 
                   feature.gene=smutans.feature.genes,
                   go.genes=smutans.go.genes,
                   cat.desc=smutans.cat.desc )
@
We find some functional categories enriched with differentially expressed genes
(table \ref{tab:goseq})
We find functional categories that are enriched with differentially expressed genes.
To find functional categories enriched with differentially expressed genes we
first find a list of differentially expressed genes. A set of genes are assigned
to functional gene categories: \Robject{smutans.go.genes}. Features of the genes in
UA159 include gene lengths: \Robject{smutans.feature.genes}.
Functional categories with numbers of genes assigned are in
\Robject{smutans.cat.desc}.
%-----------------------------------------------------------------------------
\begin{table}[!h]
\caption{{\bf Gene Ontology (GO) enrichments for differentially expressed
genes}}
\vspace{1ex}
\noindent \begin{centering}
\begin{tabular}{cl}
\hline 
$q^a$ & Description \\
\hline 
%-----------------------------------------------------------------------------
<<goseqUA159only,echo=FALSE,results=tex>>=
data( "smutans.feature.genes" )
data( "smutans.go.genes" )
data( "smutans.cat.desc" )
smutans.de2Goseq ( ua159only, qval=1e-2, 
                   feature.gene=smutans.feature.genes,
                   go.genes=smutans.go.genes,
                   cat.desc=smutans.cat.desc )
@
%-----------------------------------------------------------------------------
\hline 
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$False discovery rate estimated by the
Benjamini-Hochberg method.  All categories having at least ten genes and
$q\leq 0.05$ are displayed.\\ 
\end{flushleft}
\label{tab:goseq}
\end{table}
%-----------------------------------------------------------------------------

%--------------------------------------------------
\section{DE analysis using 835NP, 835P, and UA159 no CSP}
\label{sec:de835}
%--------------------------------------------------

We first cluster the \Sexpr{length(colnames(counts(sm835Genes)))} samples.
\begin{figure}
\centering
\includegraphics[width=.9\textwidth]{smutans-figClustNpAndP}
\caption{Empirical (black dots) and fitted (red lines) dispersion values
plotted against mean expression strength.}
\label{figClustNpAndP}
\end{figure}
<<figClustNpAndP,fig=TRUE>>=
data( "sm835Genes" )
npAndP <- newSmutans( sm835Genes, title="835NP vs. 835P" )
smutans.de2Clust( npAndP )
@

We found no functional categories enriched with differentially expressed
genes in the pairwise comparisons of 835NP, 835P, and UA159 no CSP.
Let's make a list of differentially expressed genes. 
Another kind of heatmap plots for
grouping genes could be interesting. 
But, basically we have a list of genes with
some adjusted p-values for detecting differentially expressed genes.

<<look_into_smutansGenes>>=
head( counts(sm835Genes) )
pData( sm835Genes )
@

<<sm835Samples>>=
npAndP <- newSmutans( sm835Genes, title="835NP vs. 835P" )
npAndP <- smutans.de2( npAndP, type="ua159", 
                       condA="835np", condB="835p" )
npAndUA159 <- newSmutans( sm835Genes, title="835NP vs. UA159" )
npAndUA159 <- smutans.de2( npAndUA159, type="ua159", 
                           condA="835np", condB="ua159nocsp" )
pAndUA159 <- newSmutans( sm835Genes, title="835P vs. UA159" )
pAndUA159 <- smutans.de2( pAndUA159, type="ua159", 
                          condA="835p", condB="ua159nocsp" )
smutans.de2List( npAndP, "npAndP.csv" )
smutans.de2List( npAndUA159, "npAndUA159.csv" )
smutans.de2List( pAndUA159, "pAndUA159.csv" )
@

\begin{figure}
\centering
\includegraphics[width=.9\textwidth]{smutans-figHeatmapNpAndP}
\caption{Empirical (black dots) and fitted (red lines) dispersion values
plotted against mean expression strength. \textbf{figHeatmapNpAndP}}
\label{figHeatmapNpAndP}
\end{figure}
<<figHeatmapNpAndP,fig=TRUE>>=
smutans.de2TypeHeatmap( npAndP )
@

%--------------------------------------------------
\section{DE analysis using OMZ175 and OMZ/HKRR}
\label{sec:deomz}
%--------------------------------------------------

<<look_into_smomzGenes>>=
data( "smomzGenes" )
head( counts(smomzGenes) )
pData( smomzGenes )
@

<<smomz,results=hide>>=
omz <- newSmutans( smomzGenes, title="OMZ175 vs. OMZ/HKRR" )
omz <- smutans.de2( omz, type="ua159", 
                    condA="OMZ175", condB="OMZHKRR" )
smutans.de2List( omz, "omz.csv" )
@

%--------------------------------------------------
\section{DE analysis using UA159 CSP}
\label{sec:deua159csp}
%--------------------------------------------------

<<look_into_smua159cspGenes>>=
data( "smua159cspGenes" )
head( counts(smua159cspGenes) )
pData( smua159cspGenes )
@

<<smua159csp,results=hide>>=
ua159csp <- newSmutans( smua159cspGenes, title="UA159 no CSP vs. CSP" )
ua159csp <- smutans.de2( ua159csp, type="ua159", 
                         condA="UA159noCSP", condB="UA159CSP" )
smutans.de2List( ua159csp, "ua159csp.csv" )
@

%--------------------------------------------------
\section{DE analysis using SMU86 CSP}
\label{sec:desmu86csp}
%--------------------------------------------------

<<look_into_smsmu86cspGenes>>=
data( "smsmu86cspGenes" )
head( counts(smsmu86cspGenes) )
pData( smsmu86cspGenes )
@

<<smsmu86csp,results=hide>>=
smu86csp <- newSmutans( smsmu86cspGenes, title="UA159 no CSP vs. CSP" )
smu86csp <- smutans.de2( smu86csp, type="ua159", 
                         condA="Smu86noCSP", condB="Smu86CSP" )
smutans.de2List( smu86csp, "smu86csp.csv" )
@

%--------------------------------------------------
\section{DE analysis using predicted transcripts}
\label{sec:detranscript}
%--------------------------------------------------

We need ways of supporting the predicted transcripts. We need a way of using
expression levels and counts of starts and ends of short reads to predict
transcripts. For now, we just use an ad hoc method for finding transcript start
and end sites. A likelihood can be constructed by multiplying geometric
probabilities. Given a range of possible transcription boundaries we can find
estimates of positions and geometirc parmeter $p$ that maximize the likelihood
for both ends of a transcript.

\Rpackage{GenomicRanges} might be useful in 
<<tx,eval=FALSE>>=
source("http://www.bioconductor.org/biocLite.R")
biocLite("GenomicRanges")
@

%--------------------------------------------------
\section{DE analysis using predicted small RNAs and target genes}
\label{sec:desmallrna}
%--------------------------------------------------

Create a file with intergenic regions sequences of \textit{S.\ mutans} UA159.
Create a file with bacterial genome sequences except \textit{S.\ mutans} NN2025.
BLAST the intergenic regions against the bacterial genomes. Extract sequence
alignments from the BLAST output. Refine the alignment using MUSCLE. Convert the
FASTA-format alignments to MSA-format ones.  Find small RNA candidates using
RNAz.  Find targets of the small RNA candidates using RNAplex. Targets are
scored for each small RNA candidate. The scores were used to do functional
category enrichment tests of targets for each small RNA candidate.

Table \ref{tab:goseq} 
shows functional categories enriched with target genes for predicted small
RNAs. Let's use \Rfunction{smutans.mannwhitney} to the package and other data
objects.
%-----------------------------------------------------------------------------
\begin{table}[!h]
\caption{{\bf Gene Ontology (GO) enrichments for differentially expressed
genes}}
\vspace{1ex}
\noindent \begin{centering}
\begin{tabular}{lcccl}
\hline 
small RNA & $q^a$ & count & GO & Description \\
\hline 
%-----------------------------------------------------------------------------
<<goseqUA159only,echo=FALSE,results=tex>>=
data( "smutans.genes.criteria" )
data( "smutans.go.genes" )
data( "smutans.cat.desc" )
smutans.mannwhitney( qval=0.05,
                   genes.criteria=smutans.genes.criteria,
                   go.genes=smutans.go.genes,
                   cat.desc=smutans.cat.desc )
@
%-----------------------------------------------------------------------------
\hline 
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$False discovery rate estimated by the
Benjamini-Hochberg method.  All categories having at least ten genes and
$q\leq 0.05$ are displayed.\\ 
\end{flushleft}
\label{tab:goseq}
\end{table}
%-----------------------------------------------------------------------------

<<smutansRNA>>=
data( "smutansGenes2" )
ua159only <- newSmutans( smutansGenes2, title="UA159 Only" )
ua159only <- smutans.de2( ua159only, type="ua159", 
                          condA="glucose", condB="galactose" )
allSample <- newSmutans( smutansGenes2, 
                         title="All Samples for Comparing Conditions" )
allSample <- smutans.de2( allSample )
tabUA159onlyAndAllSample <- smutans.compareResult( ua159only, 
                                                   allSample )
addmargins( tabUA159onlyAndAllSample )
@
Of the total \Sexpr{sum(tabUA159onlyAndAllSample)} gene,
\Sexpr{tabUA159onlyAndAllSample[2,2]} genes were found in both analyses. The
analysis using all of the sample found \Sexpr{tabUA159onlyAndAllSample[1,2]}
genes. The analysis using ua159-only sample found \Sexpr{tabUA159onlyAndAllSample[2,1]}
genes. We found a few predicted small RNAs that appeared to be differentially
expressed.
<<smutansRNA>>=
smutans.de2List( allSample )[grep("locus",rownames(smutans.de2List(allSample))),]
@
We also found a few predicted small RNAs that appeared to be differentially
expressed in the comparison of the two types: ``ua159'' and ``tw1.''
<<smutansRNA>>=
allSampleType <- newSmutans( smutansGenes2, title="All samples for type" )
allSampleType <- smutans.de2( allSampleType, compareCondition="no" )
smutans.de2List( allSampleType )[grep("locus",rownames(smutans.de2List(allSampleType))),]
@

\section{Prewriting about UA159 versus TW1}

\subsection{Abstract}
A bacterial transcriptome of the primary etiological agent of human dental caries,
\textit{Stretptococcus mutans}, is described using RNA-seq data. Putative
transcripts boundaries are defined. Small RNAs are predicted by computational
methods, and confirmed by RNA-seq and Southern blot experiments. We refine the
previously established importance of CcpA regulation as central metabolism and
virulence gene expression using RNA-seq data.
CcpA is known to play a direct role in carbon catabolite repression (CCR).
We describe genes that are differentially expressed in glucose-grown wild-type
(UA159) and CcpA-deficient (TW1) strains. We found that more
genes were differentially expressed in the comparison between wild-type and
CcpA-deficient strains in galactose-grown strains than in
glucose-grown strains.
We also found that more genes differentially expressed in wild-type than in
CcpA-deficient strain when comparing glucose- and galactose-grown cells. These
two findings were different from those of \cite{Abranches2008}. \fixme{We may
want to see effects of different p-value cutoff.} We confirmed that CcpA
modulates the pathogenic potential of \textit{S.\ mutans} through global control
of gene expression.
We predicted a few small RNAs based on sequence alignments and RNA secondary
structures. We confirmed them experimentally. Target genes of predicted small
RNAs were predicted. The target genes were enriched in some functional
categories. We listed highly potential small RNAs with their target genes
functional categories. 
Transcripts were predicted using RNA-seq short reads mapped on the wild-type
UA159 genome.

\subsection{Reviews of papers that cited \cite{Abranches2008}}

CCR in bacteria involves regulatory networks that activate or silence genes in
response to carbon source and availability.

\subsection{Goal: Refinement of \cite{Abranches2008}}

Comparisons of microarray and RNA-seq are not good way to go.

\subsection{Methods and Materials}

\subsubsection{Bacterial strains and growth conditions}

\subsubsection{RNA isolation}

\subsubsection{RNA sequencing}

\subsubsection{Statistical analysis for differential expressions}
We used an R package, DESeq, \cite{Anders2010} to determine differentially
expressed genes based on negative binomial model originally developed by \cite{Robinson2007}.
Steps of analyzing RNA-seq data for determining differentially expressed genes
are described in \cite{Oshlack2010}. We used a short read aligner called
\program{BWA} \cite{Li2009a} to map the sequenced short reads on a reference genome,
\textit{Streptococcus mutans} UA159 (NCBI accession NC\_004350). Because the
aligner allowed a few gaps for efficient alignment of millions of sequences of
size being 100 base pairs long, short reads that consist of RNAs and sequencing
adapter sequences would be not mapped. We used \program{cutadapt} \cite{Martin2011} to
remove parts of adapter sequences from short reads before mapping on the
reference genome. Mapped short read alignments were converted into readable
formats \cite{Li2009b}. We counted short reads aligned to annotated genes in the
reference genome, forming a table of read counts of the genes. Statistical
software R \cite{R2011} with \Rpackage{DEseq} \cite{Anders2010}
was subsequently employed to infer differentially expressed in different
biological conditions.  We also generated tracks for our recently released
Streptococcus Genome Browser (http://strep-genome.bscb.cornell.edu) that
summarize gene expression levels based the short read alignments. This allowed
us to visualize parts of the reference genome for comparing different RNA
samples. Further analyses such as functional category enrichment test
\cite{Young2010} were also employed.

\subsubsection{small RNAs}

RNAz is used to find small RNA candidates. Sequence alignments were a
critical step for finding small RNAs in the approach. Phylogenies of streptococcus species could
be considered to choose species related with Streptococcus mutans.
Alternatively, I find sequences similar to intergenic regions, and refine the
alignments using a multiple sequence alignment tool such as MUSCLE
\cite{Edgar2004a}. Steps for
achieving this goal include: 1. Exract intergenic regions including up- and
down-stream, 2. BLAST the intergenic regions against non-redundant DB to find
similar sequences \cite{Altschul1990}, 3. Refine each alignment using MUSCLE \cite{Edgar2004a}, 4. Apply RNAz to the
alignments for scoring regions for small RNAs
\cite{Washietl2005,Gruber2010}, and 5. Find transcripts with high
score of small RNAs. The transcripts must be defined in advance.

RNAplex
\cite{Tafer2008}

A review paper about small RNA predictions and targets
\cite{Tafer2008a}

RNAplfold
\cite{Bernhart2006}

\subsubsection{Transcripts}

``-c 10 -b 25 -force_gp''

\subsubsection{Gene category associations and virulence genes}

We assigned genes to Gene Ontology (GO) categories by comparing the {\em
  Streptococcus} genes to bacterial proteins from the Uniref90 database
using BLASTP, and then assigning the same GO classification as the target
gene of the uniProt GOA database if the match had an $E$-value of
$<1.0\times10^{-5}.$  
A gene family was assigned a given
classification if any of its genes was assigned that classification.  To
test for associations with replacing gene transfers we used the
recombination intensity estimated for each gene, and for additive gene
transfers we used the number of additive transfers inferred for each
family.  To test for significance, we performed a Mann-Whitney $U$ test of
the values (recombination intensities or numbers of transfers) associated
with a given category vs.\ the values for the all other genes/families.  We
used the \cite{Benjamini1995} method to correct for multiple comparisons.

\subsection{Results}

CcpA affects CCR of \textit{fruA}. 
\textit{fruA} fructan hydrolase
When Wen2002 deleted CRE sequencs in the \textit{fruA} promoter region, CCR was
almost completely alleviated. 
What are CCR-sensitive genes? 
What are the mechanism of CcpA on CCR of the \textit{fruA} operon? 
Expression levels of \textif{fruA} in \textit{S.\ mutans} strains that were
grown to early exponential phase (OD_600, 0.2) in TV medium supplemented with
0.5\% glucose or galactose. \fixme{What does this mean?} 
Deletion of \textit{ccpA} resulted in substantial relief of the CCR of
\textit{fruA} expression exerted by both glucose and fructose. 
CcpA could directly interact with the \textit{fruA} promoter region.

The role of CcpA in gene expression during growth on glucose. 
\textit{S.\ mutans} UA159 and TW1 grown under catabolite-repressing conditions
(glucose), and catabolite-derepressing conditions (galactose). Under
glucose-only condition I found xx number of differentially expressed genes. What
are the functional categories of those differentially expressed genes? What are
the genes that are most profoundly up- or down-regulated (Smu.1421 to Smu.1424,
PDH enzyme complex), ... and more? ... This regulatory control was
short-circuited in the strain growing on glucose and lacking CcpA. I showed that
some of KEGG pathways were affected by TW1 grown under glucose. We found
down-regulated genes, which would have something to do with pathogenicity. 
Upto here, I compared UA159 and TW1 under glucose.

Growth in derepressing conditions induces major changes in the UA159
transcriptome. Here, I compared glucose and galactose conditions for growing
UA159. Energy metabolism was enriched with differentially expressed genes. There
were more up-regulated genes in galactose condition compared with glucose
condition. Some genes are described in detail.

Now, I compared glucose and galactose for growing TW1.

Now, I compared UA159 and TW1 under galactose.

What genes are differentially expressed?

Name the transcripts.
I found xx transcripts. Among them xx \% were mono-cistronic operons. xx \% were
poly-cistronic operons. xx transcripts could not defined because RNA-seq short
reads were not precise. Transripts were assigned with functional categories, and
differentially expressed transcripts were found. Transcripts and functional
categories: how can I assess that transcripts consist of genes with similar
functional cateogires.

Name the small RNAs.
RNAz predicted a few small RNAs. Some of them were differentially expressed.
Some functional categories were enriched with target genes of the predicted
genes.

\subsection{Discussion}
We used RNA-seq data to delineate transcript boundaries. Although this map of
transcripts does not precisely tell us where transcripts start and end, it
informs us of which open reading frames can be transcribed together. Potentially
polycistronic operons are defined. We focused on interpretations of operon in
light of CcpA regulations.

\subsection{Abbreviations}

\begin{enumerate}
\item CCR Carbon catabolite repression
\item CcpA ?
\item IPS intracellular polysaccharide
\end{enumerate}

\section{Session Info}
<<sessi>>=
sessionInfo()
@

\bibliographystyle{unsrt}
\bibliography{siepel-mutans}

\end{document}