### 2012-09-12 package-source/sagalactiae.Rnw
Manual html.

\documentclass{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{color} 
\usepackage{setspace} 
\usepackage{rotating}
\usepackage{natbib}
\usepackage{xr}

% Text layout
\oddsidemargin 0in
\evensidemargin 0in
\topmargin -.5in
\textwidth 6.5in
\textheight 9in
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

\SweaveOpts{keep.source=TRUE,eps=FALSE,include=FALSE,width=6,height=6}
\newcommand{\Rpackage}[1]{R package \textit{#1}}
\newcommand{\program}[1]{{\texttt{#1}}}
\newcommand{\fixme}[1]{{\textbf{Fixme:} \textit{\textcolor{blue}{#1}}}}
\newcommand{\comment}[1]{}

\makeatother
\pagestyle{myheadings}
\markright{Transcriptome of Streptococcus agalactiae}

\begin{document}

\begin{titlepage}

\title{Defining Transcriptome using RNA Sequencing\\
of {\em Streptococcus agalactiae} }

\author{Vince P.\ Richards$^{1}$, Sang Chul Choi$^{2}$,
and Adam Siepel$^{2}$, and Michael J.\ Stanhope$^{1}$}

\date{ }
\maketitle
\thispagestyle{empty}

\begin{footnotesize}
\begin{center}
$^1$Department of Population Medicine and Diagnostic Sciences,\\
College of Veterinary Medicine, Cornell University, Ithaca, NY 14853, USA
\\[1ex]
$^2$Department of Biological Statistics and Computational Biology,\\
Cornell University, Ithaca, NY 14853, USA
\end{center}
\end{footnotesize}

\vspace{1in}

\begin{tabular}{lp{4.5in}}
{\bf Submission type:}& Research Article
\vspace{1ex}\\
{\bf Keywords:}&RNA-seq, non-coding RNAs
\vspace{1ex}\\
{\bf Running Head:}&Transcriptome of {\em Streptococcus agalactiae}
\vspace{1ex}\\ 
{\bf Corresponding Author:}&
\begin{minipage}[t]{4in}
 Adam Siepel\\
 102E Weill Hall, Cornell University\\
 Ithaca, NY 14853\\
 Phone: +1-607-254-1157\\
 Fax: +1-607-255-4698\\
 Email: acs4@cornell.edu
\end{minipage}
\end{tabular}

\end{titlepage}

\setcounter{page}{1}

\doublespacing

\section*{ABSTRACT}

A bacterial strain NEM316 of {\em Streptococcus agalactiae} has been known to
originate in humans. We provide a series of evidence for the hypothesis that it
is originated from cows.

\section*{INTRODUCTION}

\section*{RESULTS}

@
<<echo=FALSE,results=hide>>=
library(smutans)

for (i in c("NC_004368")) {
file.1 <- paste("../output/agalactiae/1/bwa/count-",i,".txt", sep="")              
file.out.1 <- paste(i,"-sag56-A-milk-B-broth-clust.pdf", sep="")                         
file.out.2 <- paste(i,"-sag56-A-milk-B-broth.csv", sep="")                               
file.out.3 <- paste(i,"-sag56-A-milk-B-broth.ps", sep="")

agGenes <-
  readSmutans(countsFile=file.1,
              indexFile="../output/agalactiae/1/run-analysis/count.txt.index",
              condition = c("sag56hm", "sag56hb"),
              firstFactorLabel = c("NEM316"),
              secondFactorLabel = c("sag56hm", "sag56hb"),
              name="S. agalactiae NEM316 in milk versus broth")

ag <- newSmutans( agGenes , title="S. agalactiae NEM316 in milk versus broth")

pdf(file.out.1)                                                                 
smutans.de2Clust( ag )                                                          
dev.off()

ag <- smutans.de2( ag, type="NEM316", condA="sag56hm", condB="sag56hb")
smutans.de2List( ag, file=file.out.2 )                                          
smutans.plotDiffExp(ag,file=file.out.3,qval=0.05)                               
}
@

\subsection*{Differential expression of RNA transcripts}

\section*{METHODS}

\subsection*{Bacterial strains and growth conditions}

\subsection*{RNA isolation}

\subsection*{RNA sequencing}

\subsection*{Short reads alignments}
About 20 millions of short reads were sequenced for each sample                 
(table \ref{count}).  We used a                                                 
short read aligner, \program{BWA} \citep{Li2009a}, to map the sequenced short   
reads on the reference genome, \textit{S.\ mutans} UA159 (NCBI accession        
NC\_004350). Because the aligner allowed a few gaps for efficient alignments of 
millions of sequences of size being 100 base pairs long, short reads that       
consisted of RNAs with sequencing adapter sequences would not be mapped.  We    
used \program{cutadapt} \citep{Martin2011} to remove parts of adapter sequences 
from short reads before mapping on the reference genome.  Mapped short read     
alignments were converted into readable formats using \program{samtools}        
\citep{Li2009b}.

\subsection*{Transcript predictions}

\subsection*{Prediction of small RNAs and targets}

\subsection*{Experimental validation of small RNAs}

\subsection*{Statistical analysis for differential expressions}

\subsection*{Gene category associations}

\section*{DATA ACCESS}

\section*{ACKNOWLEDGMENTS}

\fixme{Add more peoples or fundings if needed.}
This study was funded by the National Institute of Allergy and Infectious
Disease, US National Institutes of Health, under grant number AI073368-01A2
(to M.J.S.\ and A.S).  Additional support was provided by National Science
Foundation CAREER Award DBI-0644111 and a David and Lucile Packard
Fellowship for Science and Engineering (to A.S.).  
We thank Christopher Browngardt for assistance with labelling functional classes
of \textit{S.\ mutans} genes.

\section*{FIGURE LEGENDS}

\section*{FIGURES}

\section*{TABLES}

\clearpage{}

\bibliographystyle{mbe}
\bibliography{siepel-mutans}
\clearpage{}

\end{document}