### 2012-09-12 package-source/tutorial.Rnw

%\VignetteIndexEntry{Analyzing Streptococcus mutans RNA-seq data}
%\VignettePackage{smutans}

% To compile this document
% library('weaver'); rm(list=ls()); Sweave('smutans.Rnw', driver=weaver()); system('pdflatex smutans')

\documentclass{article}

\usepackage{Sweave}
\usepackage[letterpaper]{geometry}
\usepackage{hyperref,graphicx}
\usepackage{color}
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

\SweaveOpts{keep.source=TRUE,eps=FALSE,include=FALSE,width=4,height=4.5}
\newcommand{\Robject}[1]{object \texttt{#1}}
\newcommand{\Rpackage}[1]{R package \textit{#1}}
\newcommand{\Rclass}[1]{class \textit{#1}}
\newcommand{\Rfunction}[1]{{function \small\texttt{#1}}}
\newcommand{\program}[1]{{\texttt{#1}}}
\newcommand{\fixme}[1]{{\textbf{Fixme:} \textit{\textcolor{blue}{#1}}}}

\renewcommand{\floatpagefraction}{0.7}

\author{Sang Chul Choi\\[1em]Cornell University,\\ Ithaca, NY\\[1em]
\texttt{schoi@cornell.edu}}

\title{\textsf{\textbf{A Toturial of BioConductor R packages}}}

\begin{document}

\maketitle

\begin{abstract}
BioConductor R packages have been useful in analyzing RNA-seq data sets. I wish
to record for the instructional purposes.
\end{abstract}
\tableofcontents

\section{Genomes}

\subsection{Sample sequences from a genome}

\section{Short reads}

Steps of analyzing RNA-seq data for determining differentially expressed genes
are described in \cite{Oshlack2010}.

\subsection{Create short reads}

\subsection{Read short reads}

\Rpackage{ShortRead}

<<eval=FALSE,echo=TRUE>>=
library(ShortRead)

fastqFileOrg <- "/v4scratch/sc2265/rnaseq/output/ua159/1/data/FASTQ001.fq.gz"
fastqFile <- "/v4scratch/sc2265/rnaseq/output/ua159/1/bwa/FASTQ001.prinseq.fq.gz"

fastqFile <- "/v4scratch/sc2265/rnaseq/output/ua159/1/bwa/FASTQ001.recovered.fq.gz"
fq <- readFastq(fastqFile, withIds=TRUE)
fqOrg <- readFastq(fastqFileOrg, withIds=TRUE)

# Number of reads
length(fq)
# The first 3 Reads
head(sread(fq), 3)
# The first 3 Read quality scores
head(quality(fq), 3)
# The first 3 Read IDs
head(id(fq), 3)
# Tally reads by length
table(width(fq))
# Remove reads with N
fqN <- fq
fq <- fqN[grep("N",sread(fqN),invert=TRUE)]
# 
sampler <- FastqSampler(fastqFile, 100000)
fqSample <- yield(sampler)
report(qa(fqSample),dest="/v4scratch/sc2265/rnaseq/output/ua159/1/bwa/report001")

qa2 <- qa(fq, basename(fastqFile)) 
qa3 <- qa(fqSample, basename(fastqFile)) 
report(qa2,dest="/v4scratch/sc2265/rnaseq/output/ua159/1/bwa/report001")

indexBam(reads_fn)
@
X11 error in report

Reweight each read.

\subsection{Sample short reads}

\subsection{Correct bias in short reads}

<<eval=FALSE,echo=TRUE>>=
library(seqbias) 
library(Rsamtools) 
ref_fn <- "/v4scratch/sc2265/rnaseq/output/ua159/1/data/NC_004350.fna"
ref_f <- FaFile( ref_fn ) 
open.FaFile( ref_f )
reads_fn <- "/v4scratch/sc2265/rnaseq/output/ua159/1/bwa/FASTQ001.sorted.bam"
indexBam(reads_fn)
ref_seqs <- scanFaIndex( ref_f )

ref_seq <- getSeq(ref_f)
I.all <- GRanges(seqnames=Rle(c("NC_004350.1"),c(2)),
                 ranges=IRanges(c(1,1),width=rep(width(ref_seq),2)),
                 strand=Rle(strand(c("+","-")),c(1,1)))
seqlengths(I.all) <- c(width(ref_seq))

I <- random.intervals( ref_seqs, n = 5, m = 100000 )
seqs <- scanFa( ref_f, I )
neg_idx <- as.logical( I@strand == '-' ) 
seqs[neg_idx] <- reverseComplement( seqs[neg_idx] )
counts <- count.reads( reads_fn, I, binary = T )
freqs <- kmer.freq( seqs, counts )
sb <- seqbias.fit( ref_fn, reads_fn, L = 5, R = 20 )
bias <- seqbias.predict( sb, I )
bias <- seqbias.predict( sb, I.all )
bias <- seqbias.predict( sb5, I.all )

bias[[2]][99996]
seqbias.save(sb,"/v4scratch/sc2265/rnaseq/output/ua159/1/bwa/FASTQ001.yml")
sb <- seqbias.load( ref_fn, "/v4scratch/sc2265/rnaseq/output/ua159/1/bwa/FASTQ001.yml")
sb5 <- seqbias.load( ref_fn, "/v4scratch/sc2265/rnaseq/output/ua159/1/bwa/FASTQ001-500000.yml")
sb5 <- seqbias.fit( ref_fn, reads_fn, n=5e5, L = 5, R = 20 )

N <- width(ref_seq) # genome length
cnt <- count.reads(reads_fn, I.all, sb = sb5)
cnt <- count.reads(reads_fn, GRanges("seq1", IRanges(start = 1, width = N)), sb = sb5)

readlen <- 100
cov <- coverage(GRanges("NC_004350.1", IRanges(start = 1:N, width = readlen)), weight = cnt[[1]])
@

\subsection{Load biasseq}

<<eval=FALSE,echo=TRUE>>=
library(seqbias) 
library(Rsamtools) 
ref_fn <- "/v4scratch/sc2265/rnaseq/output/ua159/1/data/NC_004350.fna"
ref_f <- FaFile( ref_fn ) 
open.FaFile( ref_f )
reads_fn <- "/v4scratch/sc2265/rnaseq/output/ua159/1/bwa/FASTQ001.sorted.bam"
indexBam(reads_fn)
ref_seqs <- scanFaIndex( ref_f )

ref_seq <- getSeq(ref_f)
I.all <- GRanges(seqnames=Rle(c("NC_004350.1"),c(2)),
                 ranges=IRanges(c(1,1),width=rep(width(ref_seq),2)),
                 strand=Rle(strand(c("+","-")),c(1,1)))
seqlengths(I.all) <- c(width(ref_seq))

sb5 <- seqbias.load( ref_fn, "/v4scratch/sc2265/rnaseq/output/ua159/1/bwa/FASTQ001-500000.yml")
bias <- seqbias.predict( sb5, I.all )
@

\subsection{Compute the coverage of short reads}

<<eval=FALSE,echo=TRUE>>=
bamFile <- "/v4scratch/sc2265/rnaseq/output/ua159/1/bwa/FASTQ001.sorted.bam"
aln <- readGappedAlignments(bamFile)
aln <- as(aln, "GRanges")
coverage(aln)
cov <- coverage(aln, weight = cnt[[1]])
bias[[2]][99996]

cvg <- rep(0,width(ref_seq))
s1 <- start(aln)
s2 <- end(aln)
s3 <- strand(aln)
for (i in 1:length(aln)) {
  if (i %% 100000 == 0) {
    print(i)
  }
  if (runValue(s3[i]) == "+") {
    cvg[s1[i]:s2[i]] <- cvg[s1[i]:s2[i]] + 1/bias[[1]][s1[i]]
  } else {
    cvg[s1[i]:s2[i]] <- cvg[s1[i]:s2[i]] + 1/bias[[2]][s2[i]]
  }
}
cvgFile <- "/v4scratch/sc2265/rnaseq/output/ua159/1/bwa/FASTQ001.cvg"
save(cvg,file=cvgFile)
@

\subsection{Export coverage to a wiggle file}

\section{BWA}

\subsection{Count non-unique reads}

bamFile=/v4scratch/sc2265/rnaseq/output/ua159/1/bwa/FASTQ001.sorted.bam
samtools view \$bamFile | grep -v XT:A:U | wc -l

The automatic annotation of bacterial genomes
Emily J. Richardson and Mick Watson

rtracklayer
export()

\section{Parallel running}

\subsection{Multicore}

<<eval=FALSE,echo=TRUE>>=
install.packages(c("multicore", "snow", "foreach", "doMC", "doSNOW"))
install.packages(c("Rmpi", "doMPI"))
library(multicore) 
multicore:::detectCores()
@

\section{Apply functions}

\subsection{lapply}

/home/fs01/sc2265/Downloads/r-devel/b/bin/Rscript 1.R 8 1
<<eval=FALSE,echo=TRUE>>=
library(seqbias)
library(Rsamtools) 
library(ShortRead)

args <- commandArgs(trailingOnly = TRUE)
if (length(args) != 2)
{
  cat ("Rscript 1.R 8 1\n")
  quit("no")
}

bamFile <- "/v4scratch/sc2265/rnaseq/output/ua159/1/bwa/FASTQ001.sorted.bam"
aln <- readGappedAlignments(bamFile)
aln <- as(aln, "GRanges")

n <- as.integer(args[1]) + 1L
a <- as.integer(seq(1L,length(aln),length.out=n))
a[9] <- a[9] + 1
s.1 <- a[as.integer(args[2])]
s.2 <- a[as.integer(args[2]) + 1] - 1
quit("no")

ref_fn <- "/v4scratch/sc2265/rnaseq/output/ua159/1/data/NC_004350.fna"
ref_f <- FaFile( ref_fn ) 
open.FaFile( ref_f )
ref_seqs <- scanFaIndex( ref_f )
ref_seq <- getSeq(ref_f)
I.all <- GRanges(seqnames=Rle(c("NC_004350.1"),c(2)),
                 ranges=IRanges(c(1,1),width=rep(width(ref_seq),2)),
                 strand=Rle(strand(c("+","-")),c(1,1)))
seqlengths(I.all) <- c(width(ref_seq))

sb5 <- seqbias.load( ref_fn, "/v4scratch/sc2265/rnaseq/output/ua159/1/bwa/FASTQ001-500000.yml")
bias <- seqbias.predict( sb5, I.all )

s1 <- start(aln)
s2 <- end(aln)
s3 <- strand(aln)

# foreach (i=1:length(aln)) %do% {
cvg <- rep(0,width(ref_seq))
# for (i in 1:length(aln)) {
for (i in s.1:s.2) {
  if (runValue(s3[i]) == "+") {
    cvg[s1[i]:s2[i]] <- cvg[s1[i]:s2[i]] + 1/bias[[1]][s1[i]]
  } else {
    cvg[s1[i]:s2[i]] <- cvg[s1[i]:s2[i]] + 1/bias[[2]][s2[i]]
  }
}
cvgFile <- paste("/v4scratch/sc2265/rnaseq/output/ua159/1/bwa/FASTQ001.cvg",args[2],sep=".")
save(cvg,file=cvgFile)
close.FaFile( ref_f )
@

track type=wiggle_0 name="RNA-seq" description="RNA-seq 1" visibility=full
autoScale=on color=0,200,100 priority=30
fixedStep chrom=chr1 start=1 step=1
1
...
1

\subsection{doMC}

<<eval=FALSE,echo=TRUE>>=
library(doMC)
registerDoMC()
a <- foreach(i=1:1000000) %dopar% sqrt(i)
getDoParWorkers()
getDoParName()
getDoParVersion()
options(cores=8)
@

<<eval=FALSE,echo=TRUE>>=
cvg <- foreach (i=1:length(aln),.combine="+") %dopar% {
  if (runValue(s3[i]) == "+") {
    cvg[s1[i]:s2[i]] <- cvg[s1[i]:s2[i]] + 1/bias[[1]][s1[i]]
  } else {
    cvg[s1[i]:s2[i]] <- cvg[s1[i]:s2[i]] + 1/bias[[2]][s2[i]]
  }
  cvg
}
@

\section{Session Info}
<<session>>=
sessionInfo()
@

\bibliographystyle{unsrt}
\bibliography{siepel-mutans}

\end{document}