### 2012-10-01 package-source/mss-lin.Rnw

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Load smutans package and data
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
<<eval=FALSE,echo=FALSE,results=hide>>=
# not run when smutans R package is installed
# Use this when smutans R package is being tested.
library(DESeq)
library(goseq)
library(GenomicRanges)
library(IRanges)
library(plotrix)
library(xtable)
library(RColorBrewer)
library(lattice)
library(rtracklayer)
source("smutans/R/class_and_slots.R")
source("smutans/R/core.R")
source("smutans/R/methods.R")
load("smutans/data/smutansData.tx.RData")
load("smutans/data/smutansData.txGenes.RData")
load("smutans/data/smutansData.txGrCombined.RData")
load("smutans/data/smutansData.txAnnotation.RData")
load("smutans/data/smutansData.txRNAz.RData")
load("smutans/data/smutansData.txPileup.RData")
load("smutans/data/smutans.feature.genes.RData")
load("smutans/data/smutans.genes.criteria.RData")
load("smutans/data/smutans.go.genes.RData")
load("smutans/data/smutans.cat.desc.RData")
load("smutans/data/smutans.keg.genes.RData")
load("smutans/data/smutans.keg.desc.RData")
load("smutans/data/smutansGenes.RData")
@
<<load,echo=FALSE,results=hide>>=
# Run with smutans R package version 0.1.5 installed with R 2.14.2
# Remove eval=FALSE when the R package is installed
# R package version 0.2.1 is a new branch from 0.1.5.
# Have a separate branch from 0.1.9; say, 0.3.1 for further development.
library(smutans)
#library(lattice)
#library(rtracklayer)
data(smutansData.tx)
data(smutansData.txGenes)
data(smutansData.txGrCombined)
data(smutansData.txAnnotation)
data(smutansData.txRNAz)
data(smutansData.txPileup)
data(smutans.feature.genes)
data(smutans.genes.criteria)
data(smutans.go.genes)
data(smutans.cat.desc)
data(smutans.keg.genes)
data(smutans.keg.desc)
data(smutansGenes)
data(smutansGenes2)
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Functional association and gene description
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
<<global,echo=FALSE,results=hide>>=
inDir <- system.file("extdata", package="smutans", mustWork=TRUE)
#inDir <- "extdata"

trim <- function (x) gsub("^\\s+|\\s+$", "", x)

# geneDescription <- read.delim("feature-genome.out-geneonly.txt",head=FALSE,sep="\t")
# Note that feature-genome.out-geneonly.txt contains annotations.
# feature-genome.out-geneonly does not.
geneDescription <- read.delim(file.path(inDir,"feature-genome.out-geneonly.txt"),head=FALSE,sep="\t")
geneDescription <- geneDescription[,c(4,7)]
colnames(geneDescription) <- c("Gene ID","Gene Description")

# functionalDescription <- read.csv("feature-genome.out-geneonly-oralgen-functional.txt",head=FALSE)
functionalDescription <- read.csv(file.path(inDir,"feature-genome.out-geneonly-oralgen-functional.txt"),head=FALSE)
functionalDescription <- functionalDescription[,c(2,12)]
colnames(functionalDescription) <- c("Gene ID","Functional Class")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Transcript analyses R scripts
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
<<transcript,echo=FALSE,result=hide>>=
notExpressedTx <- smutansData.txGrCombined[elementMetadata(smutansData.txGrCombined)$zero > 0.1 & elementMetadata(smutansData.txGrCombined)$pileup < 5]
grTx <- smutansData.txGrCombined[!(elementMetadata(smutansData.txGrCombined)$zero > 0.1 & elementMetadata(smutansData.txGrCombined)$pileup < 5)]
# We query genes and annotations (genes and RNAs) againt tx's.
matchGenesTx <- findOverlaps(smutansData.txGenes,grTx)
matchAnnotationTx <- findOverlaps(smutansData.txAnnotation,grTx)
# I use gaps with IRanges because gaps with GenomicRanges does not work as expected.
grNoAnnotation <- gaps(ranges(smutansData.txAnnotation))
grNoncoding <- gaps(ranges(smutansData.txGenes))
grRNAz <- ranges(smutansData.txRNAz)
# 
matchUnannotatedRNA <- findOverlaps(grRNAz,grNoAnnotation,type="within")
txRNAz.unannotated <- grRNAz[matchMatrix(matchUnannotatedRNA)[,1]]

# A bug in GenomicRanges: the last segment is the while genome. I removed it
# from the noncoding ranges.
# matchNoncoding <- findOverlaps(grNoncoding,grTx)

# Expressed transcripts without annotated genes
grTxWGenes <- grTx[unique(matchMatrix(matchGenesTx)[,2])]
grTxWOGenes <- grTx[-unique(matchMatrix(matchGenesTx)[,2])]
grTxWOAnnotation <- grTx[-unique(matchMatrix(matchAnnotationTx)[,2])]

# Find transcripts with genes with conflicted strands
grGenesPlus <- smutansData.txGenes[strand(smutansData.txGenes)=='+']
grGenesMnus <- smutansData.txGenes[strand(smutansData.txGenes)=='-']
#mtchPlus <- findOverlaps(grGenesPlus,grTxWGenes)
#mtchMnus <- findOverlaps(grGenesMnus,grTxWGenes)
#txWithConflictedGenes <- intersect(matchMatrix(mtchPlus)[,2],matchMatrix(mtchMnus)[,2])

# Tx with genes strands
grTx2 <- grTxWGenes[!strand(grTxWGenes)=='*']

# Find the number of genes in segments.
matchGenesTx2 <- findOverlaps(smutansData.txGenes,grTx2,type="within")
number.genes.tx <- runLength(Rle(matchMatrix(matchGenesTx2)[,"subject"]))

# Length distribution of UTR regions
# grTx2 <- grTxWGenes[-txWithConflictedGenes] 
# Lengths of 5' and 3' UTRs
mtchPlus <- findOverlaps(grGenesPlus,grTx2,type="within")
mtchMnus <- findOverlaps(grGenesMnus,grTx2,type="within")
length5UTR <- c()
length3UTR <- c()
for (i in unique(matchMatrix(mtchPlus)[,"subject"])) {
  x <- matchMatrix(mtchPlus)[matchMatrix(mtchPlus)[,"subject"]==i,"query"]
  y <- min(start(grGenesPlus[x])) - start(grTx2[i])
  z <- end(grTx2[i]) - max(end(grGenesPlus[x])) 
  length5UTR <- c(length5UTR,y)
  length3UTR <- c(length3UTR,z)
}
for (i in unique(matchMatrix(mtchMnus)[,"subject"])) {
  x <- matchMatrix(mtchMnus)[matchMatrix(mtchMnus)[,"subject"]==i,"query"]
  y <- min(start(grGenesMnus[x])) - start(grTx2[i])
  z <- end(grTx2[i]) - max(end(grGenesMnus[x])) 
  length5UTR <- c(length5UTR,z)
  length3UTR <- c(length3UTR,y)
}

# Correlation of expression levels of adjacent genes.
pileup.x <- lapply(ranges(smutansData.txGenes), function(x) seqselect(smutansData.txPileup, x))
pileup.y <- mapply(function(x,y) sum(x)/y, x=pileup.x, y=width(smutansData.txGenes))
elementMetadata(smutansData.txGenes)["score"] <- pileup.y
mtch3 <- findOverlaps(smutansData.txGenes,grTx2)
x <- rle(matchMatrix(mtch3)[,"subject"])
comp.exp.two.genes <- c()
for (i in x$values[x$lengths > 1]) {
  y <- combn(matchMatrix(mtch3)[,"query"][matchMatrix(mtch3)[,"subject"]==i],2)
  for (j in 1:dim(y)[2]) {
    comp.exp.two.genes <- cbind(comp.exp.two.genes,as.matrix(elementMetadata(smutansData.txGenes)["score"][y[,j],]))
  }
}
# Correlation of expression levels of adjacent genes between transcripts.
comp.exp.two.genes.between.transcript <- c()
for (i in 1:length(x$values)) {
  if (i == length(x$values)) {
    break
  }
  y1 <- matchMatrix(mtch3)[,"query"][matchMatrix(mtch3)[,"subject"]==x$values[i]]
  i2 <- i + 1
  y2 <- matchMatrix(mtch3)[,"query"][matchMatrix(mtch3)[,"subject"]==x$values[i2]]
  l <- as.matrix(do.call(expand.grid, list(y1,y2)))
  for (j in 1:dim(l)[1]) {
    comp.exp.two.genes.between.transcript <- cbind(comp.exp.two.genes.between.transcript,as.matrix(elementMetadata(smutansData.txGenes)["score"][l[j,],]))
  }
}
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Prediction of small RNAs and targets
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
<<rna,echo=FALSE,results=hide>>=
# Read Infernal's result
#x <- import.bed("/Users/goshng/Documents/Projects/RNASeq-Analysis/test/rfam/ua159-v1/cm.bed")
#irCm <- ranges(x)$chr1
irCm <- ranges(smutansData.txAnnotation[1:274])
# Read RNAz's result
ir1 <- gaps(ranges(smutansData.txAnnotation))
ir2 <- ranges(smutansData.txRNAz)
irRNAz <- ir2
mtchUnannotatedRNA <- findOverlaps(ir2,ir1,type="within")
txRNAz.unannotated <- ir2[matchMatrix(mtchUnannotatedRNA)[,1]]
# Read Pichon's result
# x <- import.bed("/Users/goshng/Documents/Projects/RNASeq-Analysis/test/pichon/ua159-v1/pichon.bed")
x <- import.bed(file.path(inDir,"pichon.bed"))
irPichon <- ranges(x)$chr1
mtchUnannotatedPichon <- findOverlaps(irPichon,ir1,type="within")
txPichon.unannotated <- irPichon[matchMatrix(mtchUnannotatedPichon)[,1]]
txPichon.reduced <- reduce(txPichon.unannotated)
elementReducedPichon <- findOverlaps(txPichon.reduced, txPichon.unannotated, select="first")
# Read sRNAscanner's result
# x <- import.bed("/Users/goshng/Documents/Projects/RNASeq-Analysis/test/sRNAscanner/ua159-v1/srnascanner.bed")
x <- import.bed(file.path(inDir,"srnascanner.bed"))
irScanner <- ranges(x)$chr1
mtchUnannotatedScanner <- findOverlaps(irScanner,ir1,type="within")
txScanner.unannotated <- irScanner[matchMatrix(mtchUnannotatedScanner)[,1]]
txScanner.reduced <- reduce(txScanner.unannotated)
elementReducedScanner <- findOverlaps(txScanner.reduced, txScanner.unannotated, select="first")
# Merge three results of RNAz, Pichon, and Scanner.
irUnion <- reduce(c(irRNAz,irPichon,irScanner))
names(irUnion) <- paste("PredSmallRNA",seq(length(irUnion)),sep="-")
# Read genes in ranges
irGene <- ranges(smutansData.txGenes)
# Find neighboring genes
left.right <- c()
for (k in seq(length(irUnion))) {
  x <- irUnion[k]
  irGeneUnion <- sort(c(irGene,x))
  i <- grep("PredSmallRNA",names(irGeneUnion))
  j.left <- NA
  j.left.distance <- NA
  j.left.description <- NA
  if (i > 1) {
    j.left <- names(irGeneUnion)[i-1]
    j.left.distance <- start(irGeneUnion)[i] - end(irGeneUnion)[i-1]
    j.left.description <- as.character(geneDescription[geneDescription[,1]==j.left,2])
    if (nchar(j.left.description) > 30) {
      j.left.description <- paste(substr(j.left.description,1,25),"...")
    }
  }
  j.right <- NA
  j.right.distance <- NA
  j.right.description <- NA
  if (i < length(names(irGeneUnion))) {
    j.right <- names(irGeneUnion)[i+1]
    j.right.distance <- start(irGeneUnion)[i+1] - end(irGeneUnion)[i]
    j.right.description <- as.character(geneDescription[geneDescription[,1]==j.right,2])
    if (nchar(j.right.description) > 30) {
      j.right.description <- paste(substr(j.right.description,1,25),"...")
    }
  }
  left.right <- rbind(left.right, c(j.left.description,j.left,j.left.distance,names(x),j.right.distance,j.right,j.right.description))
}
rm(irGeneUnion,k,x,i,j.left,j.left.distance,j.right,j.right.distance)
rm(j.left.description,j.right.description)
colnames(left.right) <- c("Description","Left gene","Distance","Predicted sRNA",
                          "Distance", "Right gene","Description")
xtable.tabNeighborSmallRNA <- xtable( left.right,
  display=c("s","s","s","d","s","d","s","s"),
  digits=c(0,0,0,1,0,1,0,0),
  align=c('l','r','r','l','r','l','l','l'),
  label='tabNeighborSmallRNA',
  caption='{\\bf Neighboring genes of predicted small RNAs}' 
)
################################################################################
# Rename predicted RNAs (determined by RNAz) using irUnion or the union set of
# predicted regions.
data(smutans.genes.criteria)
matchRNAz <- findOverlaps(irRNAz,irUnion)
x <- irUnion[matchMatrix(matchRNAz)[,2]]
smutans.genes.criteria <- smutans.genes.criteria[c(TRUE, colnames(smutans.genes.criteria)[-1] %in% names(irRNAz))]
colnames(smutans.genes.criteria)[-1] <- names(x)[match(colnames(smutans.genes.criteria)[-1],names(irRNAz))]

# DE small RNAs
library(smutans)

countsFile <- file.path(inDir,"count.nocds")
indexFile <- file.path(inDir,"count.nocds.txt.index")
################################################################################
# Comparison of UA159GLU and UA159GAL in small RNAs
sampleA <- "UA159GLU"
sampleB <- "UA159GAL"
csvFile = paste("srna-",sampleA,"-",sampleB,".csv",sep="")
clustFile = paste("srna-",sampleA,"-",sampleB,".pdf",sep="")
sampleTitle <- paste(sampleA, sampleB)
#####################
# Read in count data.
#cutoff.count=10
#conds <- scan(file=indexFile,what="character",quiet=TRUE)
#condition = c(sampleA, sampleB)
#subconds <- conds %in% condition
#conds <- conds[subconds]
#firstFactor <- condition
#countsTable <- read.delim (countsFile, header=TRUE, stringsAsFactors=TRUE)       
#rownames(countsTable) <- countsTable$gene
#countsTable <- countsTable[,-1]
#countsTable <- countsTable[,subconds]
#countsTable <- countsTable[(rowSums(countsTable) >= cutoff.count),]
#sGenes <- newCountDataSet( countsTable, data.frame(condition=conds,type="ua159") )
# Read in count data.
#####################
sGenes <-                                                                       
  readSmutans (countsFile=countsFile,
               indexFile =indexFile,
               condition = c(sampleA, sampleB),                                 
               firstFactorLabel = c("reference"),                               
               secondFactorLabel = c(sampleA, sampleB))
omz <- newSmutans( sGenes, title=sampleTitle )                                  
omz <- smutans.de2( omz, type="reference",
                    condA=sampleA, condB=sampleB )
# smutans.de2List( omz, csvFile )
x <- smutans.de2List( omz, qval=1e-1 )                                                 
x <- x[,c("id","log2FoldChange","padj")]
colnames(x) <- c("Predicted small RNA ID","$\\log_2$(gal/glc)", "$P$-value")
x <- x[abs(x[,2])>1,]
xtable.tabRNAua159Only <- xtable( x[order(x[,2],decreasing=TRUE),],
  display=c("s","s","f","E"),
  digits=c(0,0,1,0),
  align=c('l','l','r','r'),
  label='tabRNAua159Only',
  caption='{\\bf List of predicted small RNA regions differentially expressed in UA159 cells growing in TV-glucose (glc) and TV-galactose (gal).}'
)

################################################################################
# Comparison of TW1GLU and TW1GAL in small RNAs
sampleA <- "TW1GLU"
sampleB <- "TW1GAL"
csvFile = paste("srna-",sampleA,"-",sampleB,".csv",sep="")
clustFile = paste("srna-",sampleA,"-",sampleB,".pdf",sep="")
sampleTitle <- paste(sampleA, sampleB)
#####################
# Read in count data.
#cutoff.count=10
#conds <- scan(file=indexFile,what="character",quiet=TRUE)
#condition = c(sampleA, sampleB)
#subconds <- conds %in% condition
#conds <- conds[subconds]
#firstFactor <- condition
#countsTable <- read.delim (countsFile, header=TRUE, stringsAsFactors=TRUE)       
#rownames(countsTable) <- countsTable$gene
#countsTable <- countsTable[,-1]
#countsTable <- countsTable[,subconds]
#countsTable <- countsTable[(rowSums(countsTable) >= cutoff.count),]
#sGenes <- newCountDataSet( countsTable, data.frame(condition=conds,type="tw1") )
# Read in count data.
#####################
sGenes <-
  readSmutans (countsFile=countsFile,
               indexFile =indexFile,
               condition = c(sampleA, sampleB),
               firstFactorLabel = c("tw1"),
               secondFactorLabel = c(sampleA, sampleB))
omz <- newSmutans( sGenes, title=sampleTitle )                                  
omz <- smutans.de2( omz, type="tw1",
                    condA=sampleA, condB=sampleB )
# smutans.de2List( omz, csvFile )
x <- smutans.de2List( omz, qval=1e-1 )                                                 
x <- x[,c("id","log2FoldChange","padj")]
colnames(x) <- c("Predicted small RNA ID","$\\log_2$(gal/glc)", "$P$-value")
x <- x[abs(x[,2])>1,]
xtable.tabRNAtw1Only <- xtable( x[order(x[,2],decreasing=TRUE),],
  display=c("s","s","f","E"),
  digits=c(0,0,1,0),
  align=c('l','l','r','r'),
  label='tabRNAtw1Only',
  caption='{\\bf List of predicted small RNA regions differentially expressed in TW1 cells growing in TV-glucose (glc) and TV-galactose (gal).}' 
)

################################################################################
# Comparison of UA159GLU and TW1GLU in small RNAs
sampleA <- "UA159GLU"
sampleB <- "TW1GLU"
csvFile = paste("srna-",sampleA,"-",sampleB,".csv",sep="")
clustFile = paste("srna-",sampleA,"-",sampleB,".pdf",sep="")
sampleTitle <- paste(sampleA, sampleB)
#####################
# Read in count data.
#cutoff.count=10
#conds <- scan(file=indexFile,what="character",quiet=TRUE)
#condition = c(sampleA, sampleB)
#subconds <- conds %in% condition
#conds <- conds[subconds]
#firstFactor <- condition
#countsTable <- read.delim (countsFile, header=TRUE, stringsAsFactors=TRUE)       
#rownames(countsTable) <- countsTable$gene
#countsTable <- countsTable[,-1]
#countsTable <- countsTable[,subconds]
#countsTable <- countsTable[(rowSums(countsTable) >= cutoff.count),]
#sGenes <- newCountDataSet( countsTable, data.frame(condition=conds,type="glucose") )
# Read in count data.
#####################
sGenes <-
  readSmutans (countsFile=countsFile,
               indexFile =indexFile,
               condition = c(sampleA, sampleB),
               firstFactorLabel = c("glucose"),
               secondFactorLabel = c(sampleA, sampleB))
omz <- newSmutans( sGenes, title=sampleTitle )                                  
omz <- smutans.de2( omz, type="glucose",
                    condA=sampleA, condB=sampleB )
# smutans.de2List( omz, csvFile )
x <- smutans.de2List( omz, qval=1e-1 )                                                 
x <- x[,c("id","log2FoldChange","padj")]
colnames(x) <- c("Predicted small RNA ID","$\\log_2$(TW1/UA159)", "$P$-value")
x <- x[abs(x[,2])>1,]
xtable.tabRNAGlucoseOnly <- xtable( x[order(x[,2],decreasing=TRUE),],
  display=c("s","s","f","E"),
  digits=c(0,0,1,0),
  align=c('l','l','r','r'),
  label='tabRNAGlucoseOnly',
  caption='{\\bf List of predicted small RNA regions differentially expressed in UA159 and TW1 when growing in TV-glucose.}' 
)

################################################################################
# Comparison of UA159GAL and TW1GAL in small RNAs
sampleA <- "UA159GAL"
sampleB <- "TW1GAL"
csvFile = paste("srna-",sampleA,"-",sampleB,".csv",sep="")
clustFile = paste("srna-",sampleA,"-",sampleB,".pdf",sep="")
sampleTitle <- paste(sampleA, sampleB)
#####################
# Read in count data.
#cutoff.count=10
#conds <- scan(file=indexFile,what="character",quiet=TRUE)
#condition = c(sampleA, sampleB)
#subconds <- conds %in% condition
#conds <- conds[subconds]
#firstFactor <- condition
#countsTable <- read.delim (countsFile, header=TRUE, stringsAsFactors=TRUE)       
#rownames(countsTable) <- countsTable$gene
#countsTable <- countsTable[,-1]
#countsTable <- countsTable[,subconds]
#countsTable <- countsTable[(rowSums(countsTable) >= cutoff.count),]
#sGenes <- newCountDataSet( countsTable, data.frame(condition=conds,type="galactose") )
# Read in count data.
#####################
sGenes <-
  readSmutans (countsFile=countsFile,
               indexFile =indexFile,
               condition = c(sampleA, sampleB),
               firstFactorLabel = c("galactose"),
               secondFactorLabel = c(sampleA, sampleB))
omz <- newSmutans( sGenes, title=sampleTitle )                                  
omz <- smutans.de2( omz, type="galactose",
                    condA=sampleA, condB=sampleB )
# smutans.de2List( omz, csvFile )
x <- smutans.de2List( omz, qval=1e-1 )                                                 
x <- x[,c("id","log2FoldChange","padj")]
colnames(x) <- c("Predicted small RNA ID","$\\log_2$(TW1/UA159)", "$P$-value")
x <- x[abs(x[,2])>1,]
xtable.tabRNAGalactoseOnly <- xtable( x[order(x[,2],decreasing=TRUE),],
  display=c("s","s","f","E"),
  digits=c(0,0,1,0),
  align=c('l','l','r','r'),
  label='tabRNAGalactoseOnly',
  caption='{\\bf List of predicted small RNA regions differentially expressed in UA159 and TW1 when growing in TV-galactose.}' 
)
@
<<glucoseOnly,echo=FALSE,results=hide>>=
glucoseOnly <- newSmutans( smutansGenes, title="Glucose Only" )
glucoseOnly <- smutans.de2( glucoseOnly, condition="glucose", 
                            condA="ua159", condB="tw1" )
x <- smutans.de2List( glucoseOnly, qval=1e-3 )
x <- x[,c("id","log2FoldChange","padj")]
colnames(x) <- c("Gene ID","$\\log_2$(TW1/UA159)", "$P$-value")
x <- merge(x, geneDescription, by.x = "Gene ID", by.y = "Gene ID")
x <- x[abs(x[,2])>1,]
xtable.tabGlucoseOnly <- xtable( x[order(x[,2],decreasing=TRUE),],
  display=c("s","s","f","E","s"),
  digits=c(0,0,1,0,0),
  align=c('l','l','r','r','p{3.5in}'),
  label='tabGlucoseOnly',
  caption='{\\bf List of genes differentially expressed in UA159 and TW1 when growing in TV-glucose.}' 
)
x <- merge(x, functionalDescription, by.x = "Gene ID", by.y = "Gene ID")
tableGlucoseOnly <- x
@
<<ua159only,echo=FALSE,results=tex>>=
ua159only <- newSmutans( smutansGenes, title="UA159 Only" )
ua159only <- smutans.de2( ua159only, type="ua159", 
                          condA="glucose", condB="galactose" )
#save(ua159only, file=file.path("ua159only.RData"))
#load("ua159only.RData")
x <- smutans.de2List( ua159only, qval=1e-3 )
x <- x[,c("id","log2FoldChange","padj")]
colnames(x) <- c("Gene ID","$\\log_2$(gal/glc)", "$P$-value")
x <- merge(x, geneDescription, by.x = "Gene ID", by.y = "Gene ID")
x <- x[abs(x[,2])>1,]
xtable.tabUA159Only <- xtable( x[order(x[,2],decreasing=TRUE),],
  display=c("s","s","f","E","s"),
  digits=c(0,0,1,0,0),
  align=c('l','l','r','r','p{3.5in}'),
  label='tabUA159Only',
  caption='{\\bf List of genes differentially expressed in UA159 cells growing in TV-glucose (glc) and TV-galactose (gal).}'
)
x <- merge(x, functionalDescription, by.x = "Gene ID", by.y = "Gene ID")
tableUA159Only <- x
@
<<tw1only,echo=FALSE,results=tex>>=
tw1only <- newSmutans( smutansGenes, title="TW1 Only" )
tw1only <- smutans.de2( tw1only, type="tw1", 
                          condA="glucose", condB="galactose" )
#save(tw1only, file=file.path("tw1only.RData"))
#load("tw1only.RData")
x <- smutans.de2List( tw1only, qval=1e-3 )
x <- x[,c("id","log2FoldChange","padj")]
colnames(x) <- c("Gene ID","$\\log_2$(gal/glc)", "$P$-value")
x <- merge(x, geneDescription, by.x = "Gene ID", by.y = "Gene ID")
x <- x[abs(x[,2])>1,]
xtable.tabTW1Only <- xtable( x[order(x[,2],decreasing=TRUE),],
  display=c("s","s","f","E","s"),
  digits=c(0,0,1,0,0),
  align=c('l','l','r','r','p{3.5in}'),
  label='tabTW1Only',
  caption='{\\bf List of genes differentially expressed in TW1 cells growing in TV-glucose (glc) and TV-galactose (gal).}' 
)
x <- merge(x, functionalDescription, by.x = "Gene ID", by.y = "Gene ID")
tableTW1Only <- x
@
<<galactoseOnly,echo=FALSE,results=tex>>=
galactoseOnly <- newSmutans( smutansGenes, title="Galactose Only" )
galactoseOnly <- smutans.de2( galactoseOnly, condition="galactose", 
                              condA="ua159", condB="tw1" )
#save(galactoseOnly, file=file.path("galactoseOnly.RData"))
#load("galactoseOnly.RData")
x <- smutans.de2List( galactoseOnly, qval=1e-3 )
x <- x[,c("id","log2FoldChange","padj")]
colnames(x) <- c("Gene ID","$\\log_2$(TW1/UA159)", "$P$-value")
x <- merge(x, geneDescription, by.x = "Gene ID", by.y = "Gene ID")
x <- x[abs(x[,2])>1,]
xtable.tabGalactoseOnly <- xtable( x[order(x[,2],decreasing=TRUE),],
  display=c("s","s","f","E","s"),
  digits=c(0,0,1,0,0),
  align=c('l','l','r','r','p{3.5in}'),
  label='tabGalactoseOnly',
  caption='{\\bf List of genes differentially expressed in UA159 and TW1 when growing in TV-galactose}.'
)
x <- merge(x, functionalDescription, by.x = "Gene ID", by.y = "Gene ID")
tableGalactoseOnly <- x
@
<<save,echo=FALSE,results=hide>>=
save.image(file="mss-lin.RData")
@
<<sequence,eval=FALSE,echo=FALSE,results=hide>>=
library(Biostrings)
x <- readFASTA("NC_004350.fna")
y <- DNAString(x[[1]]$seq)
irUnion[35]
i <- 35;toString(subseq(y,start(irUnion)[i],end(irUnion)[i]))
i <- 71;toString(subseq(y,start(irUnion)[i],end(irUnion)[i]))
i <- 116;toString(subseq(y,start(irUnion)[i],end(irUnion)[i]))
i <- 117;toString(subseq(y,start(irUnion)[i],end(irUnion)[i]))
i <- 204;toString(subseq(y,start(irUnion)[i],end(irUnion)[i]))
i <- 205;toString(subseq(y,start(irUnion)[i],end(irUnion)[i]))
rm(x,y)
# PredSmallRNA-35, PredSmallRNA-71, PredSmallRNA-116, PredSmallRNA-117, and 
# PredSmallRNA-204.
# In the other three sets of pairwise comparison, we found only a few
# differentially expressed regions:
# the region, PredSmallRNA-35, (Glucose/Galactose - TW1 background);
# PredSmallRNA-204, and PredSmallRNA-205 (UA159/TW1 - Glucose condition); and
# PredSmallRNA-116, and PredSmallRNA-205 (UA159/TW1 - Galactose condition).
@

\documentclass[11pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{color} 
\usepackage{setspace} 
\usepackage{rotating}
\usepackage{xr}
\usepackage{url}
\usepackage{longtable}
\usepackage{lscape}
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
\usepackage[round,comma,numbers]{natbib}

\usepackage{xcolor}
\usepackage{titlesec}
\titleformat{\section}[runin]{\bfseries\filcenter}{}{1em}{}
\titleformat{\subsection}[runin]{\bfseries}{}{1em}{}
\setcounter{secnumdepth}{0}
%\titlespacing*{\section}{0pt}{0pt}{0pt}
%\titlespacing*{\subsection}{0pt}{0pt}{0pt}

% \usepackage{../../latex/sty/genres}
\def\@cite#1#2{(#1\if@tempswa , #2\fi)}
\def\@biblabel#1{}

% Supplementary material
\externaldocument{mss-lin-supp}

% Text layout
\oddsidemargin 0in
\evensidemargin 0in
\topmargin -.5in
\textwidth 6.5in
\textheight 9in
\parskip 0.25in

\makeatletter
\SweaveOpts{keep.source=TRUE,eps=FALSE,include=FALSE,width=6,height=6}
\newcommand{\fixme}[1]{\textbf{(COMMENT BY: }\textit{\textcolor{red}{#1}})}
\newcommand{\change}[1]{\textit{\textcolor{blue}{#1}}}
\newcommand{\comment}[1]{}
\makeatother

% \pagestyle{myheadings}
\pagestyle{empty}
\markright{RNA-seq analysis in S. mutans}

\begin{document}

\begin{titlepage}

\title{CcpA and Catabolite Repression Explored by RNA-Seq in\\
{\em Streptococcus mutans}}

\author{Lin Zeng$^{1,4}$, Sang Chul Choi$^{2,3,4}$, Charles D.\ Danko$^{3}$, \\
Adam Siepel$^{3}$, Michael J.\ Stanhope$^{2}$, and Robert A.\ Burne$^{1}$}

% Melissa J.\ Hubisz$^{3}$, \\
% Paulina Dmitrievna Pavinski Bitar$^{2}$, \\

\date{ }
\maketitle
\thispagestyle{empty}

\begin{footnotesize}
\begin{center}
$^1$Department of Oral Biology,\\
College of Dentistry, University of Florida, Gainesville, Florida 32610 USA
\\[1ex]
$^2$Department of Population Medicine and Diagnostic Sciences,\\
College of Veterinary Medicine, Cornell University, Ithaca, NY 14853, USA
\\[1ex]
$^3$Department of Biological Statistics and Computational Biology,\\
Cornell University, Ithaca, NY 14853, USA
\\[1ex]
$^4$LZ and SCC contributed equally to this work.
\\
\end{center}
\end{footnotesize}

\vspace{1in}

\begin{tabular}{lp{4.5in}}
{\bf Submission type:}& Research Article
\vspace{1ex}\\
{\bf Keywords:}&Carbon catabolite repression, CcpA, RNA deep sequencing,
biofilm, dental caries
\vspace{1ex}\\
{\bf Running Head:}&RNA-seq analysis in {\em S.\ mutans}
\vspace{1ex}\\ 
{\bf Corresponding Author:}&
\begin{minipage}[t]{4in}
 Robert Burne\\
 Department of Oral Biology, University of Florida\\
 College of Dentistry, P. O.  Box 100424\\
 Gainesville, FL 32610\\
 Phone: +1-352-273-8850\\
 Fax: +1-352-273-8829\\
 Email: rburne@dental.ufl.edu
\end{minipage}
\end{tabular}

\end{titlepage}

\setcounter{page}{2}

\doublespacing
\raggedright
\parindent 2em
\parskip 0pt

\section*{ABSTRACT}

A bacterial transcriptome of the primary etiological agent of human dental
caries, {\em Streptococcus mutans}, is described here using RNA deep sequencing.
In comparison with previously published data generated using microarray
assays, differential expression profiles of the transcriptome in the context of
a strain lacking the CcpA protein and carbohydrates revealed good agreement with
DNA microarrays, yet expanded considerably the repertoire of DNA sequences that
showed statistically-significant changes in expression as a function of the
presence of CcpA and growth carbohydrate. Some of novel mRNAs and small RNAs
appear by virtue of their location or identity to influence the regulation
and prioritization of carbohydrates.
Our study demonstrated the validity of
RNA-seq as an alternative to DNA microarrays in studying gene regulation in {\em
S.\ mutans} by yielding a more precise landscape of transcriptomic changes in
response to specific mutations and growth conditions.

\clearpage{}

\section*{INTRODUCTION}

About one fifth of children between the ages of 2 and 19 were reported to have
untreated dental caries in the United States (National Center for Health
Statistics, 2010). It is generally accepted that the presence of acid-tolerant
bacteria, a carbohydrate-rich diet and a susceptible host are all required for
development of dental caries. In fact, carbohydrates introduced into the oral
cavity provide the preferred energy sources for the members of the oral
microbiome that are regarded as significant contributors to the caries process:
aciduric streptococci and {\em Actinomyces} subspecies, lactobacilli,
bifidobacteria and {\em Scardovia} species. These organisms are particularly
effective at converting host- and diet-derived carbohydrates into organic acids
that drive demineralization of the tooth.

Cariogenic bacteria, including the primary etiological agent of human dental
caries, {\em Streptococcus mutans}, are usually equipped with multiple pathways
for the internalization and catabolism of carbohydrates \citep{Ajdic2002}. The
development of these repertoires of carbohydrate catabolic pathways likely
reflects adaptation to the complex combination of carbohydrates that are
secreted by the host in glycoproteins and other glycoconjugates, as well as to
the variety of simple and complex carbohydrates that became part of the human
diet a few thousand years ago. Because of the complexities of the repertoire of
carbohydrates to which oral biofilms are exposed and the intermittent feeding
patterns of humans, it is reasonable to conclude that many of the most abundant
members of the oral microbiota have evolved sophisticated pathways to rapidly
and efficiently prioritize the assimilation, catabolism and storage of
carbohydrates. Likewise, when the diet becomes enriched for carbohydrates, the
microorganisms \change{that} must adapt their utilization of substrates as the
physico-chemical and microbial composition of the biofilms acquires the
characteristics associated with enhanced cariogenic potential. This includes
enrichment for the aforementioned aciduric organisms, lower biofilm pH and other
changes in microenvironments, e.g. redox that are well documented to alter
bacterial gene expression and physiology
\citep{Ahn2007,Burne2009,Kolenbrander2000,Marquis1995}.

Carbon catabolite repression (CCR) allows bacteria to utilize carbon sources in
a selective fashion, turning off non-essential catabolic functions while
activating pathways required for utilization of preferred carbohydrates and
other carbon sources \citep{Deutscher2008,Gorke2008}. In a number of low GC
content Gram-positive pathogenic bacteria, CCR has also been shown to affect the
expression of a variety of virulence factors in response to the source and
amount of carbohydrate \citep{Gorke2008}. CCR in these organisms is primarily
exerted through the catabolite control protein A, CcpA. When preferred
carbohydrate sources are present, as sensed through the accumulation of
glycolytic intermediates, a protein kinase is activated (HprK) that can
phosphorylate the general sugar:phosphotransferase protein HPr at serine 46
(HPr-Ser46-P). Hpr-Ser46-P functions as a co-factor for the binding of CcpA to
conserved catabolite response elements (CRE) found near the start sites of
target genes, which can activate or repress gene expression depending on the
gene, other factors and the position of the CRE \citep{Deutscher2008}.

CcpA was shown to regulate carbohydrate metabolism and virulence expression in
{\em S.\ mutans} in a transcriptomic study using the microarray technique based
on DNA hybridization \citep{Abranches2008}. A RegPrecise search
(http://regprecise.lbl.gov) of the {\em S.\ mutans} UA159 genome yielded 99
genes in 48 operons with potential CcpA CREs detected in their promoter regions
\citep{Novichkov2010}. Although only a small portion of these genes have been
confirmed to be subject to CCR regulation in {\em S.\ mutans}, many more have
homologous genes whose regulation in related bacterial species have been
demonstrated to be regulated by CcpA.  A comparison with the results of our
previous Microarray analysis indicated that at a twofold cutoff, near half of
the operons suggested by the RegPrecise website were absent in the genes found
differentially expressed by the study \citep{Abranches2008}. Therefore, we
intend to take advantage of this {\em in silico} work and compare it to genes
identified by both Microarray and RNA-seq analyses as having significantly
differential expression.  This comparison could be used as one of the criteria
for validating the two technologies. Furthermore, another {\em S.\ mutans} study
using microarrays that included all intergenic regions (IGR) of UA159 has
reported a number of IGRs having differential expression in cell subpopulations
sorted according to the expression levels of a {\em comX-GFP} fusion, a reporter
that responds to extracellular bacterial quorum-sensing signal CSP
(competence-stimulating peptide) \citep{Lemme2011}. Some of these IGRs likely
encode as yet unknown, regulatory RNA or transcripts that may help regulate the
development of competence, biofilm formation and stress tolerance in {\em S.\
mutans}. We speculate that some other IGRs in the genome of {\em S.\ mutans}
could play regulatory roles in carbohydrate metabolism.

As the experiment of choice to many who studied bacterial transcriptome in the
last decade or so, microarray analysis has proven to be reliable, rapid and
comparatively economical \citep{Conway2003}. However, in standard arrays only
genes or transcripts that are included in a predetermined set of probes can be
detected in any given assay, therefore the scope of microarray analysis became a
limiting factor when a function was mediated by unannotated transcripts or
genes. Very recently, RNA deep sequencing (RNA-seq) has facilitated
annotation-free \change{probe-free?} detection of transcripts, with greater
sensitivity and dynamic ranges in RNA expression levels than microarrays
\citep{Croucher2010,Filiatrault2011}. In addition, RNA-seq allows for the
analysis of RNAs in non-coding regions, small RNAs (sRNAs) and antisense
transcripts. However as a relatively new tool in modern bacteriology, cautions
and considerations are due while embracing this technology. These include the
inherent bias due to steps that require inverse \change{reverse?} transcription
and ligation, and particularly for bacterial samples, relatively low
signal-to-noise ratios partly due to the presence of a large ribosomal RNA
population. In order to test the applicability of RNA-seq in transcriptome study
in {\em S.\ mutans}, we used standard RNA deep sequencing technique to sequence
RNAs in 13 {\em S.\ mutans} samples, focusing on messenger and small RNA
expression levels in response to different carbohydrates (glucose and galactose)
in both the wild-type strain (UA159) and a {\em ccpA} mutant (TW1)
\citep{Abranches2008}. As we have previously conducted Microarray analysis with
the same set of strains growing in identical conditions, these new data allowed
us to validate the RNA-seq technology while expanding the scope of RNAs that may
play a role in regulation of carbohydrate metabolism, a central factor in {\em
S.\ mutans} virulence.

\section*{MATERIALS AND METHODS}

\subsection*{Bacterial strains and growth conditions}

{\em Streptococcus mutans} strains UA159 and TW1 \citep{Abranches2008} were
maintained on BHI (Difco Laboratories, Detroit, MI) agar plates, and bacterial
cultures used for extraction of RNA were prepared with Tryptone-vitamin
\citep{Burne1999} based medium supplemented with 0.5\% of glucose or galactose
(Sigma, St. Louis, MO). Four repeats, each in the volume of 15 mL, were included
for the culture of strain UA159 growing in TV-galactose, while 3 repeats were
used for each of the other 3 cultures. Bacterial cultures were incubated
statically in the presence of 5\% of CO$_2$ at $37^\circ$C until they reached
mid-exponential phase (OD$_{600}$ $\approx$ 0.5), harvested by centrifugation at
$4^\circ$C for 10 min, treated with bacterial RNAprotect Bacteria Reagent
(Qiagen, Germantown, MD), and immediately stored at $-80^\circ$C.

\subsection*{RNA isolation, enrichment, and sequencing}

Total RNA was extracted from frozen bacterial cells using the RNeasy Mini kit
(Qiagen) according to previously published protocols \citep{Ahn2005}. To remove
16S and 23S rRNAs, 10 $\mu$g of high-quality total RNA was processed using a
MICROB{\em Express}{\texttrademark}  Bacterial mRNA Enrichment Kit (Ambion of Life
Technologies, Grand Island, NY), twice, before precipitating with ethanol and
resuspended in 25 $\mu$L of nuclease-free water. The final quality of enriched
mRNA samples was analyzed using an Agilent Bioanalyzer (Agilent Technologies,
Santa Clara, CA).
\change{
We generated cDNA libraries from the enriched mRNA samples using the TruSeq
Illumina kit (Illumina, San Diego, CA), following instructions from the
supplier. Deep sequencing of the RNA samples were performed at the Cornell
University Life Sciences Core Laboratories Center (Ithaca, NY). 
}
\fixme{RAB:more details please: e.g., adapter sequences? Can the protocols be easily obtained using info provided here?}
\fixme{SCC:The raw RNA-seq data set can tell what adapter sequences we used.
Mike and Paulina should know how precise the protocol is.
I need to ask Mike and Paulina about this part. 
I removed a subsection ``RNA sequencing'' becuase it is redundant.}

\subsection*{Short reads alignments}

Approximately 20 million short-reads were obtained for each sample.  Because the
aligner, BWA \citep{Li2009a}, allowed a few gaps for efficient alignments of
millions of reads of size being around 100 base pairs long, shorter reads that
consisting of partially sequencing adapters would not be mapped.  We, therefore,
removed adapter sequences of each short-read \citep{Martin2011}, and trimmed the
resulting sequences by quality score from the 3'-end \citep{Schmieder2011}
before mapping onto the reference genome of strain UA159 (GenBank accession
AE014133) using the short-read aligner.  Mapped short-read alignments were then
converted into readable formats using SAMTOOLS \citep{Li2009b}.

\subsection*{Transcript predictions}

RNA transcripts were inferred by applying a hidden Markov model to site-wise
expression levels \citep{Martin2010}. A pileup command of BWA was used to
convert the short-read alignments into pileup values, which were taken as the
site-wise expression levels along the genome. Site-wise expression levels are a
list of non-negative integers that represent numbers of short-reads mapped at a
particular genomic position. The transcript inference program, ParseRNAseq,
with the following options was used: ``-c 10 -b 25 -force gp'' to bin expression
levels into 25 parts, and to allow 10 emission states for relative expression
intensity. Genes that were annotated in close approximation were used as the
predicted parts of a transcript, while no information regarding the precise
transcription initiation- or stop-sites were pursued in this study due to
limited resources. For similar reasons, the orientation of each transcript could
not be verified solely based on RNA-seq data; instead we used the information of
annotated genes to determine the strands of predicted transcripts.

\subsection*{Prediction of small RNAs and targets}

Although our RNA-seq protocol did not specifically enrich non-coding small RNAs
in the cDNA preparation, small RNAs were retained in the RNA samples as only
ribosomal RNAs were depleted using specific oligonucleotides. Consequently, it
was difficult to discriminate cDNA originated from non-coding RNAs small from
that of mRNAs, as non-coding RNAs expressions were often masked by the expression
of neighboring background mRNA. Therefore, we sought a way of predicting small
RNAs using a computer program called RNAz \citep{Gruber2010}, which used homologous sequences
and RNA secondary structures to predict putative non-coding small RNAs. Because
sequence alignment was a critical step for finding small RNAs via this approach,
intergenic regions that also included up- and down-stream sequences were
extracted, BLAST-searched against a database of bacterial genomes
\citep{Altschul1990}, and the resultant sequence alignments were further refined
using a program, MUSCLE \citep{Edgar2004a}. Subsequently, RNAz was applied
to the alignments for scoring intergenic regions for putative small RNAs
\citep{Washietl2005}. Target genes for each candidate small RNA were also
predicted using RNAplex \citep{Tafer2008} and RNAplfold \citep{Bernhart2006},
and the resultant target genes were used to perform functional category
enrichment tests based on their scores by these two programs.
\change{
In addition to the above procedure for predicting small RNAs, we also employed a
method of rho-independent terminators (RITs) identification
\citep{Kingsford2007}, to prepare candidate small RNAs, that were subsequently
scored using RNAz, and a transcriptional signal-based method to identify
intergenic sRNA transcriptional units (TUs) \citep{Sridhar2010}. 
}

\subsection*{Statistical analysis for differential expressions}

The R package DESeq \citep{Anders2010} was used to determine differentially
expressed genes on the basis of the negative binomial model \citep{Robinson2007}.
Detailed steps of analyzing RNA-seq data for identifying differentially
expressed genes were well described \citep{Oshlack2010}; short-reads aligned to
a particular annotated gene in the reference genome were counted, generating a
table of read counts of all the protein coding genes. Statistical software R
\citep{R2011} with the R package DESeq was then employed to infer differentially
expressed genes in the four biological conditions from two factors of UA159/TW1
and glucose/galactose. \change{For normalizing expression levels among samples,
total sequencing depths for each sample were estimated as the median of the
ratios of the sample's counts to geometric mean across samples
\citep{Anders2010,Robinson2010}.}

\subsection*{Gene functional category associations}

Three sets of gene categories were compiled for testing functional associations
of differentially expressed genes. First, genes were each assigned to Gene
Ontology (GO) categories by comparing to bacterial proteins from the Uniref90
database using the BLASTP program, and a GO classification was assigned if the
match had an $E$-value of $<1.0\times10^{-5}.$ A gene family was assigned a
given classification if any of its genes was assigned that classification.
Second, we collated functional classes of genes in the Oral Pathogen Sequence
Databases available at \url{http://oralgen.lanl.gov}. Third, genes were mapped
onto a set of metabolite pathways for {\em S.\ mutans} available at Kyoto
Encyclopedia of Genes and Genomes (KEGG) \citep{Kanehisa2012}. Using the sets of
gene categories for {\em S.\ mutans}, we performed a Mann-Whitney {\em U} test of the
values of target genes of small RNAs in association with a given category versus
those for the other categories, and also conducted a variation of Fisher's exact
test accounting for gene lengths to test enrichments of differentially expressed
genes \citep{Young2010}. Both tests were corrected for multiple testing
hypotheses \citep{Benjamini1995}.
\fixme{RAB:do we need more explanation here or is the
reference sufficient?}
\fixme{SCC:These are enough for reproducing the results.}

\subsection*{UCSC genome browser tracks and data availability}

We created tracks for recently released Streptococcus Genome Browser
(\url{http://strep-genome.bscb.cornell.edu}) that summarized the results of our
{\em S.\ mutans} transcriptomic analysis with known genes, gene expression
levels based on the short-read alignments, predicted putative transcript
segments, and predicted non-coding small RNAs. The results can be accessed by
clade ``Streptococcus,'' genome ``S. mutans UA159,'' and \change{assembly ``January
2006.''} These tracks can be used to inspect loci of interest and to compare the
results of different RNA-seq data sets. They can also be queried and intersected
with other tracks using the UCSC Table Browser.

\section*{RESULTS and DISCUSSION}

\subsection*{Clustering of RNA samples}

Based on the results of read counts of all annotated genes, a total of 13
RNA-seq samples were clustered without supervision. As illustrated in Figure
\ref{figClust}, the effect of carbohydrate source appeared to be significantly
stronger than that of the loss of the {\em ccpA} gene. Nevertheless, the
replicates of the same bacterial strain growing in the same carbohydrate
conditions clearly clustered together, indicating that the transcriptomic shifts
were the results of both sugar specificity and regulation by CcpA.

\subsection*{Predicted transcripts}

We obtained a set of transcripts using the RNA-seq data for each of the 13
RNA-seq samples. In order to summarize the 13 sets of predicted transcripts,
all of the transcripts of each set were scored based on their
average site-wise expression levels in relation to transcripts in the rest of
set. More highly-ranked transcripts were placed first on the reference genome
based on their scores, while lower-ranked transcripts were purged from the
transcriptome map. In doing so, the final set of transcripts was formed as
non-overlapping transcripts. The distribution of expression levels
of the predicted transcripts are shown in Figure \ref{figGrCombinedExpression}.
We designated transcripts as expressed whose average site-wise expression levels
were greater than 5 and proportion of sites of zero site-wise expression levels
were less than 10\%. Of 
\Sexpr{length(smutansData.txGrCombined)} predicted transcripts,
\Sexpr{length(notExpressedTx)} 
ones were either not expressed or very poorly expressed.  Genes that belonged to
expressed transcripts were designated as expressed.  Of the total 
\Sexpr{length(smutansData.txGenes)}
of genes, 
\Sexpr{length(as.matrix(matchGenesTx)[,1])}
(\Sexpr{as.integer(length(as.matrix(matchGenesTx)[,1])/length(smutansData.txGenes)*100)}\%) 
were designated as expressed (Figure \ref{figLength}). 
Genes that were not designated as expressed could
be because of incorrect gene annotations, or no expression under the
experimental conditions. Short reads might be not be alignable on duplicate
genes because reads could be aligned to multiple places, and those were not
mapped unequivocally \citep{Li2009a}.
Of the 
\Sexpr{length(grTx)} 
expressed transcripts, 
\Sexpr{length(grTxWOAnnotation)} 
contained no annotated genes or RNAs
\fixme{SCC: I might need to find those expressed transcripts on unannotated regions in a table}.
To assess the tranascript map by measuring how constant expression levels
between genes within polycistronic transcripts, gene expression levels for
adjacent genes within expressed transcripts were more correlated than those
between adjacent transcripts (Figure \ref{figExpressionTwoGene}).  Among the 
\Sexpr{length(unique(matchMatrix(matchGenesTx2)[,"subject"]))} 
final segments, 
\Sexpr{sum(number.genes.tx==1)} 
segments contained a single gene while 
\Sexpr{sum(number.genes.tx > 1)} 
contained more than one gene.

Figure \ref{figTranscriptLength} shows the length distribution of expressed
transcripts with annotated
coding genes and unannotated regions. Unannotated regions could include
potential non-coding regions and novel coding genes. Transcripts of unannotated
regions appeared to be shorter than those with coding genes. Most of the
unannotated regions with length near 360 bp could be non-coding regions, and the
rest of the unannotated regions with length near 1000 bp could be new coding
genes. Some of the unannotated genes may represent pseudogene regions. We used
strandnesses of known genes to find strands of segments although we could not
tell strands of the predicted transcripts solely based on our RNA-seq data
because a strand-specific RNA-Seq technique was not employed here. Of the 
\Sexpr{length(grTx)} 
expressed transcripts, 
\Sexpr{sum(strand(grTx)=='*')} 
included annotated genes of both strands 
(\Sexpr{as.integer(sum(strand(grTx)=='*')/length(grTx)*100)}\%).
We then filtered out the transcripts containing genes with conflicting
strandness, leaving 
\Sexpr{length(grTx2)}
expressed transcripts. Figure \ref{figUTRLength} shows length distributions of
5' and 3' UTR of our finally determined  
\Sexpr{length(grTx2)}
transcripts.

\subsection*{Prediction of small RNAs and targets}

\change{
We merged the three sets of regions that were predicted to contain 
small RNAs from the three different prediction methods. 
While the three sets of predicted regions contained
\Sexpr{length(irRNAz)} (RNAz),
\Sexpr{length(irPichon)} (rho-independent terminators based identification), and
\Sexpr{length(irScanner)} (transcriptional signal-based method),
the number of elements in the union set was \Sexpr{length(irUnion)}.
Of the \Sexpr{length(irUnion)} predicted regions of the union set,
\Sexpr{sum(countOverlaps(irUnion,c(irPichon,irRNAz,irScanner)) > 1)}
regions were constructed from multiple predictions. 
Among the predicted regions of the union set,
\Sexpr{sum(countOverlaps(irUnion,irCm) > 0)}
regions overlapped regions annotated by Infernal (v1.0)
using default options for models in RFAM 9.1 \citep{Gardner2009,Nawrocki2009}
(\Sexpr{length(irCm)} structural RNA regions). 
Because we focused our search of small RNAs on non-genic regions, only 
\Sexpr{sum(countOverlaps(irUnion,irGene) > 0)}
predicted small RNA regions overlapped known genes.  We surveyed neighboring
genes of regions that could contain small RNAs, showing that some of predictions
were clustered together, and others were lonely placed in between known genes
(Table \ref{tabNeighborSmallRNA}).  The functional categories of predicted
target genes of RNAz predicted small RNAs are shown in Table \ref{tabGoseq}.  We
also used the union set of predicted small RNAs containing regions to determine
which regions were differentially expressed.  Because of a smaller set of the
union regions than the gene set of {\em S.\ mutans}, we found only a few regions
that were differentially expressed. For example, when comparing UA159 wild-type
cells grown in glucose and galactose, among 243 predicted regions, 129 regions
were removed from the DE analysis because of their low or zero expression
levels. The following five regions were expressed differentially:
PredSmallRNA-35, PredSmallRNA-71, PredSmallRNA-116, PredSmallRNA-117, and 
PredSmallRNA-204. In the other three sets of pairwise comparison, we found only
a few differentially expressed regions:
the region, PredSmallRNA-35, (Glucose/Galactose - TW1 background);
PredSmallRNA-204, and PredSmallRNA-205 (UA159/TW1 - Glucose condition); and
PredSmallRNA-116, and PredSmallRNA-205 (UA159/TW1 - Galactose condition).
We obtained the MFE structure drawing with encoding base-pair probabilities
for PredSmallRNA-204 regions (Figure \ref{figRNA204}), which shows high
probability of base-pairing among bases in the region.
}

\fixme{RAB:Why these sRNA do not appear to match that in Table S1? If possible,
can we combine these two tables into one?.}
\fixme{SCC:Table \ref{tabNeighborSmallRNA} is a table for regions that were
predicted to contain small RNAs, while table \ref{tabGoseq} is table for targets
of predicted small RNAs.}

\fixme{RAB:Also please make sure these structures are of the ones we want to
present.} 
\fixme{SCC:Now, we have only one RNA secondary structure in the supp.}

\fixme{SCC:I remove the section of differentially expressed small RNAs. We now
have only one section for small RNAs. I deleted most of the text of the previous
analyses of DE study of some small RNA predictions by RNAz.}
% \subsection*{Differentially expressed small RNAs and transcripts}

\subsection*{Differential expression of RNA transcripts}

Two independent factors of the biological processes being studied here included
the effect of CcpA and the treatment growth in glucose versus galactose,
resulting in four pair-wise comparisons. We compared two conditions of one
factor with the other factor being fixed at one condition. The cut-off for
designating a gene as being differentially expressed was a change in transcript
levels of at least 2-fold \change{and $P$-value of $10^{-3}$}.
\fixme{SCC:I have updated tables and values mentioned in the main text using
these criteria of 2-fold change and $10^{-3}$ $P$-value.}

\paragraph{Comparison between wild-type and ccpA mutant in glucose-grown cells:}
Table \ref{tabGlucoseOnly} lists differentially expressed genes in the
comparison of wild-type UA159 and the {\em ccpA} mutant strain TW1 grown with
glucose. 
Of 
\Sexpr{nrow(counts(smutansGenes))}
genes, 
\Sexpr{sum(tableGlucoseOnly[,2]>0)}
genes
were found to have at least twofold increase in expression levels in the mutant
relative to the wild-type strain. At the same time, 
\Sexpr{sum(tableGlucoseOnly[,2]<0)}
genes showed at least
twofold reduction in expression in the loss of CcpA. Among these 
\Sexpr{nrow(tableGlucoseOnly)}
differentially
expressed genes, 
\Sexpr{length(grep("Energy metabolism",tableGlucoseOnly[,"Functional Class"]))}
% \change{Used to be 20}
were involved in energy metabolism including glycolysis,
fermentation and sugar utilization; 
\Sexpr{length(grep("PTS",tableGlucoseOnly[,"Functional Class"]))}
% \change{Used to be 10}
encoded Enzyme II components of the PTS 
(see Tables \ref{tabGoseqGlucoseOnly} and \ref{tabKeggGlucoseOnly} for gene
ontology and KEGG terms that were enriched with differentially expressed genes
in this comparison)
and were required for transporting glucose, mannose, sucrose and cellobiose; and
\Sexpr{length(grep("Two-component",tableGlucoseOnly[,"Functional Class"])) + length(grep("Regulatory",tableGlucoseOnly[,"Functional Class"]))}
were classified as regulatory or two-component system genes
(Figures \ref{figDotplotGlucoseOnly}, and \ref{figPieGlucoseOnly}).
We also found                                                                   
\Sexpr{length(grep("Hypothetical|Unknown|Unassigned",tableGlucoseOnly[,"Functional Class"]))}
% \change{Used to be 8}
genes annotated as unknown functions.  We confirmed that the most up-regulated genes
in TW1 encoded the components of the pyruvate dehydrogenase (PDH) enzyme complex
(SMU.1421c$\sim$1424c) as reported previously \citep{Abranches2008}.
Down-regulated genes included a cytoplasmic alpha-amylase (SMU.1590) that was
located downstream of {\em ccpA} \citep{Simpson1998}, and a fructosyltransferase
gene {\em ftf} (SMU.2028) encoding the enzyme involved in converting sucrose to
a fructan homopolymer \citep{Shiroza1988}. These results appeared to be
generally consistent with those of the microarray analysis
\citep{Abranches2008}.

In agreement with the {\em in silico} prediction provided by the RegPrecise
website, results from the RNA-seq study (but not that of microarray study)
identified the following as differentially expressed: genes encoding enzymes
pyruvate kinase (SMU.1190c; decreased by twofold in TW1) and pyruvate-formate
lyase (SMU.402c; increased by twofold in TW1), a putative thiamine biosynthesis
lipoprotein (SMU.1088), SMU.1125c of a putative ribonucleoside-metabolism
operon, a ribosome associated protein (SMU.500), the sucrose-6-phosphate
hydrolase (SMU.1843) and sucrose-PTS EII (SMU.1841c) \citep{Sato1989}, the major
glucose-PTS EII$^{\mathrm{Man}}$ (SMU.1877$\sim$1879) \citep{Abranches2003}, the
cellobiose-PTS operon (SMU.1596c$\sim$1600c) \citep{Zeng2009} and another
$\beta$-glucoside-PTS EII (SMU.980) \citep{Cote2000}. On the other hand, genes
identified only in the microarray study that matched the RegPrecise predictions
included: glucose-6-phosphate isomerase (SMU.307), a putative
ribulose-monophosphate-PTS EIIC component (SMU.270) and a hypothetical protein
(SMU.799c).

There also appeared to be evidence suggesting improved consistency using RNA-seq
technology. For example, a proven CCR-sensitive operon, the glycogen
biosynthesis {\em glg} cluster that spans SMU.1535c$\sim$1539c
\citep{Abranches2008,Harris1992}, was shown here by RNA-seq to be uniformly
derepressed in strain TW1 growing on glucose. The microarray analysis of the
same strains grown in the same condition, however, indicated that the expression
of only one of the genes, SMU.1537c, was altered in a statistically significant
way. The same was true with an inducible fructose-PTS operon (SMU.870$\sim$872).
Furthermore, RNA-seq analysis, at a twofold cutoff, showed disproportionally
fewer genes 
(\Sexpr{sum(tableGlucoseOnly[,2]<0)} out of \Sexpr{nrow(tableGlucoseOnly)}) 
% (11 out of 62) 
with decreased expression in TW1 as compared to UA159, whereas
microarray analysis showed 
% RESULT NOT FROM THIS RESULT
110 
genes with decreased expression and 
% RESULT NOT FROM THIS RESULT
61 
genes with
increased expression in TW1 at the same cutoff, when both were growing in
TV-glucose. Considering the collective evidence regarding CCR in Gram-positive
bacteria, it was conceivable that CcpA exerted its function predominantly through
negative regulations \citep{Deutscher2008}.

\paragraph{Comparison between glucose- and galactose-grown UA159 cells:}
Figure \ref{figPieUA159Only} shows proportions of functional categories that
included differentially expressed genes between glucose-grown and
galactose-grown UA159 cells. Among the four pairwise comparisons this showed the
largest number 
\Sexpr{nrow(tableUA159Only)} 
% \change{(used to be 139)}
of genes being
differentially expressed (Table \ref{tabUA159Only}). Again, energy metabolism
had the most
(\Sexpr{as.integer(length(grep("Energy metabolism",tableUA159Only[,"Functional Class"]))/nrow(tableUA159Only)*100)}\%)
% \change{used to be (33\%)}
differentially expressed genes and as many as 
% \change{used to be 13}
\Sexpr{length(grep("PTS",tableUA159Only[,"Functional Class"]))}
PTS component genes were
affected, reflecting potential changes in uptake of lactose, glucose, fructose,
mannose, cellobiose, sucrose and trehalose
(see Table \ref{tabGoseqUA159Only} to see that carbohydrate metabolic process is
enriched with genes expressed differentially in the change of energy source
carbohydrate).
The most highly enriched KEGG category in the effect of galactose on UA159 was
galactose metabolism (Table \ref{tabKeggUA159Only}).
Clearly, glucose is a potent
activator of CCR, negatively affecting the expression of various glycolytic
components and PTS operons in {\em S.\ mutans} genome. 
Notably in {\em S.\ mutans}, the
effects of glucose on some of these PTS operons, including EII$^{\mathrm{Lac}}$
\citep{Zeng2010} EII$^{\mathrm{Cel}}$
\citep{Zeng2009} and EII$^{\mathrm{Fru}}$ (SMU.1956c$\sim$1961c)
\citep{Zeng2008} have been shown to be mediated primarily
through the major glucose-PTS EII$^{\mathrm{Man}}$, despite the ability of CcpA
to exert its
regulation indirectly by regulating the expression of EII$^{\mathrm{Man}}$
\citep{Zeng2008}. Therefore,
the transcriptomic alteration shown here is likely a compounded result of these
two tiers of regulation. Notwithstanding, this profound change in gene
expression was also partly due to the fact that galactose and its derivatives
could serve as inducers for activating expression of the Leloir pathway and
tagatose pathway genes that are responsible for assimilating galactose and
lactose \citep{Ajdic1998,Zeng2010}. These include consecutive genes SMU.1488c to SMU.1496c
(tagatose pathway) and genes SMU.886 to SMU.889 (Leloir pathway). Interestingly,
over a quarter 
(\Sexpr{as.integer(length(grep("Hypothetical|Unknown|Unassigned",tableUA159Only[,"Functional Class"]))/nrow(tableUA159Only)*100)}\%)
of all differentially expressed genes were considered
hypothetical or of unknown function,
and most 
(\Sexpr{as.integer(sum(tableUA159Only[grep("Hypothetical|Unknown|Unassigned",tableUA159Only[,"Functional Class"]),2]<0)/length(grep("Hypothetical|Unknown|Unassigned",tableUA159Only[,"Functional Class"]))*100)}\%)
% (\Sexpr{as.integer(sum(tableUA159Only[,2]<0)/nrow(tableUA159Only)*100)}\%)
% \change{used to be 26} genes
of these were
up-regulated in the presence of glucose (Figure \ref{figDotplotUA159Only}). Notably, other genes
suppressed by glucose include the {\em glg} operon (SMU.1535c$\sim$1539c), the
glucan-synthesizing enzymes {\em gtfBC} (SMU.1004$\sim$1005)
\citep{Shiroza1987,Ueda1988}, a glucan-binding protein {\em gbpC} (SMU.1396c)
\citep{Sato1997} and a NAD-dependent aldehyde dehydrogenase {\em gabD}
(SMU.2127).

Our previous microarray analysis of the same set of samples showed similar
patterns of change in gene expression, with significant portion (28 out of 90)
of the affected genes encoding hypothetical proteins \citep{Abranches2008}. In comparison with the
predictions at RegPrecise website, the RNA-seq analysis identified 
12 
\fixme{SCC:Can Lin change this number 12 using the updated Table
\ref{tabUA159Only}?}
presumably
CCR-sensitive operons, 
8 
\fixme{SCC:Can Lin change this number 8 as well using the updated Table
\ref{tabUA159Only}?}
of which were missing from the microarray study
performed under identical conditions: SMU.113$\sim$114 (fructose-1-phosphate kinase
and a fructose-PTS EIIBC) \citep{Wen2001}, SMU.1088 (a putative thiamine biosynthesis
lipoprotein), SMU.1843 ({\em scrB}), SMU.2127 ({\em gabD}), SMU.1596$\sim$1600 (cellobiose
operon), SMU.1877$\sim$1879 (EII$^{\mathrm{Man}}$), SMU.980 ($\beta$-glucoside PTS EII) and the
sucrose-PTS ({\em scrA}). Conversely, 8 operons from the list of RegPrecise were
identified by Microarray study and 4 exclusively: SMU.575c ({\em lrgA})
\citep{Ahn2010}, SMU.396
(a glycerol-uptake facilitator protein, {\em glpF}), SMU.877$\sim$883 (the multiple-sugar
metabolism pathway, {\em msm}) \citep{Russell1992} and SMU.402 (pyruvate-formate lyase).

\paragraph{Comparison between glucose- and galactose-grown cells in the
absence of CcpA:} 
As mentioned above, the effect of sugar appears dominant to that of loss of
CcpA.  Similar to observations made with UA159, when transcriptomes from TW1
grown in glucose or galactose were contrasted, energy metabolism 
(\Sexpr{as.integer(length(grep("Energy metabolism",tableTW1Only[,"Functional Class"]))/nrow(tableTW1Only)*100)}\%)
% \change{used to be (36\%)}
was the largest category of genes differentially expressed
(Figure \ref{figPieTW1Only}).
Similarly to the comparison of UA159 cells grown in glucose and galactose,
carbohydrate metabolic process was highly ranked in the functional categories
enriched with differentially expressed genes (Table \ref{tabGoseqTW1Only}),
and a KEGG category, called galactose metabolism, was found to be enriched with
the differentially expressed genes (Table \ref{tabKeggTW1Only}).
In contrast to
wild-type cells however, only
\Sexpr{length(grep("PTS",tableTW1Only[,"Functional Class"]))}
% \change{used to be 7}
PTS genes were identified that encoded PTS components for lactose, fructose and
mannitol. Notably up-regulated by galactose in TW1 were the tagatose and Leloir
pathways for galactose/lactose metabolism, the {\em glg} operon for glycogen
metabolism, {\em gbpC}, {\em gtfB} and the {\em msm} pathway.  A total of 
\Sexpr{sum(tableTW1Only[,2]<0)} 
% \change{used to be 57}
genes were found down-regulated 
in galactose-grown TW1 cells (Table \ref{tabTW1Only}), with majority 
(\Sexpr{as.integer(sum(tableTW1Only[grep("Energy metabolism",tableTW1Only[,"Functional Class"]),2]<0)/sum(tableTW1Only[,2]<0)*100)}\%)
of
which being designated as hypothetical proteins (Figure \ref{figDotplotTW1Only}).

Different from our previous Microarray study that saw as many as 515 genes
differentially expressed by at least twofold in TW1, with more than 50\% having
lower expression in galactose conditions, RNA-seq showed only 
\Sexpr{nrow(tableTW1Only)} 
\change{used to be 130}
\fixme{SCC: Where do we find this number 130? The number 42 is the number of rows in Table 4.}
in total
affected in TW1 as a function of the growth carbohydrate. As the Microarray
study was carried out using the total RNA and the RNA samples used in RNA-seq
analysis were enriched for mRNA and other non-rRNA populations, this discrepancy
perhaps implies an improved signal-to-noise ratio for the RNA-seq technology,
since the difference between TW1 cells grown in glucose and galactose conditions
is expected to be greatly reduced due to the loss of CcpA.

\paragraph{Comparison between the wild-type and ccpA mutant growing in
TV-galactose:} 
As
reported previously, galactose is less effective than glucose at triggering CCR
in {\em S.\ mutans} \citep{Zeng2008}, so the finding that there were fewer differences in the
transcriptomes of UA159 and TW1 in galactose-grown cells was expected. Although
energy metabolism was still the predominant category 
(\Sexpr{as.integer(length(grep("Energy metabolism",tableGalactoseOnly[,"Functional Class"]))/nrow(tableGalactoseOnly)*100)}\%)
% \change{used to be (27\%)}
among all
differentially expressed genes 
(Figures \ref{figPieGalactoseOnly}, and \ref{figDotplotGalactoseOnly}).
Similar to what was observed in glucose-grown cells, a
few genes including those for PDH components (SMU.1421c$\sim$1424c) were up-regulated
in TW1, and alpha-amylase (SMU.1590c) and $\beta$-D-fructosyltransferase {\em
ftf}
(SMU.2028c) were down-regulated. In contrast, no major PTS genes were
differentially expressed in the mutant during growth in galactose. Because both
strains were grown in galactose, genes involved in galactose and lactose
utilization were likely being expressed similarly and therefore no differences
were noted. Consequently, no gene ontology terms were found significantly
associated with differentially expressed genes. These results are largely
consistent with our previous microarray analysis \citep{Abranches2008}.

\section*{CONCLUDING REMARKS}

Our current knowledge of the {\em S.\ mutans} transcriptome is limited to annotated or
predicted genes. Whole transcriptome sequencing using high throughput sequencing
technologies, or deep RNA sequencing (RNA-seq) allowed us to reveal the
transcriptomic topography of {\em S.\ mutans}. Instead of aiming at pin-pointing
transcript start or end sites we focused on differential expression of genes in
response to nutrient availability and deletion of {\em ccpA} from UA159. Although we
did not use strand-specific RNA-seq that could be important in gene-dense
transcriptomes of species like bacteria or archaea, our application of a hidden
Markov model for predicting transcripts to RNA-seq data in conjunction with
known gene annotations allowed for inference of segments of putatively
co-transcribed regions. In doing so, we attempted to ensure a high-quality
transcriptome assembly. Although this map of transcripts did not precisely tell
us where transcripts start and end, it does provide information as to which open
reading frames could be co-transcribed under the conditions tested. 
In comparison with our previous Microarray study of differential gene expression
using the same set of samples, we observed good agreement between these two
methodologies, as well as identification of genes by each method alone that
appear to match an {\em in silico} analysis published online at RegPrecise website,
and consistent with other independent studies. Therefore, we have identified
additional genes or transcripts that are differentially regulated in {\em S.\ mutans}
in response to carbohydrate change or loss of CcpA. We also found discrepancies
of annotated genes and their transcription, for example a gene containing two
transcripts and non-coding regions producing significant number of transcripts.
We hope that our work could serve as a foundation for a comprehensive study of
the {\em S.\ mutans} transcriptome.

\section*{ACKNOWLEDGMENTS}

\fixme{SCC:I wish to add A.S. and C.D.D. as coauthors. I am not sure whether
this is okay. Mike would know better than I do about who should be in the list
of coauthors.}
This study was funded by the National Institute of Allergy and Infectious
Disease, US National Institutes of Health, under grant number AI073368-01A2 (to
M.J.S. and A.S). Additional support was provided by National Science Foundation
Career Award DBI-0644111 and a David and Lucile Packard Fellowship for Science
and Engineering (to A.S). We
thank Christopher Browngardt for assistance with labeling functional classes of
{\em S.\ mutans} genes.

\renewcommand*{\refname}{REFERENCES}
\bibliographystyle{jbactas}
\bibliography{burne}
\clearpage{}

\section*{TABLES}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Table 1. Functional category enrichments of target genes
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%-----------------------------------------------------------------------------
\begin{table}[!h]
\caption{{\bf Gene Ontology enrichments for target genes of putative small RNAs.}}
\vspace{1ex}
\noindent \begin{centering}
\begin{tabular}{lcccl}
\hline 
small RNA & $q^a$ & count & GO & Description \\
\hline 
%-----------------------------------------------------------------------------
<<tabGoseq,echo=FALSE,results=tex>>=
smutans.mannwhitney( qval=0.05,
                     genes.criteria=smutans.genes.criteria,
                     go.genes=smutans.go.genes,
                     cat.desc=smutans.cat.desc )
@
%-----------------------------------------------------------------------------
\hline 
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$False discovery rate estimated by the
Benjamini-Hochberg method.  All categories having at least ten genes and
$q\leq 0.05$ are displayed.\\ 
\end{flushleft}
\label{tabGoseq}
\end{table}
%-----------------------------------------------------------------------------

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Table 2. UA159 vs. TW1 with glucose fixed
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage{}
<<tabGlucoseOnly,echo=FALSE,results=tex>>=
print( xtable.tabGlucoseOnly,
       tabular.environment='longtable',
       caption.placement="top",
       include.rownames=FALSE,
       floating=FALSE,
       sanitize.text.function = function(x){x}) 
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Table 3. Glucose vs. Galactose with UA159 fixed
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage{}
<<tabUA159Only,echo=FALSE,results=tex>>=
print( xtable.tabUA159Only, 
       tabular.environment='longtable',
       caption.placement="top",
       include.rownames=FALSE,
       floating=FALSE,
       sanitize.text.function = function(x){x}) 
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TAble 4. Glucose vs. Galactose with TW1 fixed
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage{}
<<tabTW1Only,echo=FALSE,results=tex>>=
print( xtable.tabTW1Only,
       tabular.environment='longtable',
       caption.placement="top",
       include.rownames=FALSE,
       floating=FALSE,
       sanitize.text.function = function(x){x}) 
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% UA159 vs. TW1 with galactose fixed
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage{}
<<tabGalactoseOnly,echo=FALSE,results=tex>>=
print( xtable.tabGalactoseOnly, 
       tabular.environment='longtable',
       caption.placement="top",
       include.rownames=FALSE,
       floating=FALSE, 
       sanitize.text.function = function(x){x}) 
@

\clearpage{}
\section*{FIGURE LEGENDS}
\setlength{\parindent}{0pt}
\setlength{\parskip}{2ex}

{\bf Figure \ref{figClust}.
Cluster of the 13 RNA-seq samples.} 
Heatmap shows the Euclidean distances between pairs of the samples as calculated
from the variance-stabilizing transformation of the count data. One factor is
the kind of growth carbohydrate (glucose or galactose), and another is the
presence or absense of CcpA from the UA159 strain of {\em S.\ mutans}. The {\em
ccpA} deletion mutant type is designated tw1.

{\bf Figure \ref{figGrCombinedExpression}.
Distribution of expression levels of the predicted transcripts.}
The last bin sums from the expression level ranging from 1000 to 50,000. The
expression level is measured as the average number of reads mapped on predicted
transcripts.

{\bf Figure \ref{figExpressionTwoGene}.
Scatter plot of expression levels of pairs of adjacent genes located in the same
transcripts.}
(Top) Two genes in the same tranascripts are plotted using their expression
levels in log10 scale.
(Bottom) Two genes in different tranascripts are plotted using their expression
levels in log10 scale.

{\bf Figure \ref{figDotplotGlucoseOnly}. 
Distribution of functional classes of differntially expressed genes
in comparing UA159 wild-type and TW1 mutant-type grown in glucose.} 
Functional class ``Energy metabolism; Sugars'' contains 8 although 7 points are
shown, because two values of the 8 points are not different enough to be
displayed in two points.

{\bf Figure \ref{figDotplotUA159Only}. 
Distribution of functional classes of differntially expressed genes
in comparing UA159 wild-type grown in glucose and galactose.}

\clearpage{}

\section*{FIGURES}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figure 1. Cluster of 13 RNA-seq samples using genes expression
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
<<figClust,echo=FALSE,result=hide,fig=TRUE>>=
allSample <- newSmutans( smutansGenes )
smutans.de2Clust( allSample )
@
\begin{figure}[!h]
\centering
\includegraphics[width=.9\textwidth]{mss-lin-figClust}
\caption{}
\label{figClust}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figure 2. Distribution of transcript expression 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
<<figGrCombinedExpression,fig=TRUE,echo=FALSE>>=
l <- list(elementMetadata(smutansData.txGrCombined)$pileup)
bin <- c(as.integer(seq.int(0, 1000, length.out=15)), 50000)
multhist(l,breaks=bin,ylab="Frequency of Transcripts",xlab="Expression Level")
@
\begin{figure}
\centering
\includegraphics[width=.9\textwidth]{mss-lin-figGrCombinedExpression}
\caption{}
\label{figGrCombinedExpression}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figure 3. Correlation of gene expression within and among transcripts
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
<<figExpressionTwoGene,fig=TRUE,echo=FALSE>>=
plot( log10(comp.exp.two.genes[1,]), log10(comp.exp.two.genes[2,]),
      xlab="Expression of first Gene in log10",
      ylab="Expression of second Gene in log10",
      pch=19,cex=0.5)
abline(b=1,lty=2)
@
<<figExpressionTwoGeneBetweenTranscript,fig=TRUE,echo=FALSE>>=
plot( log10(comp.exp.two.genes.between.transcript[1,]),
      log10(comp.exp.two.genes.between.transcript[2,]),
      xlab="Expression of first Gene in log10",
      ylab="Expression of second Gene in log10",
      pch=19,cex=0.5)
abline(b=1,lty=2)
@
\begin{figure}
\centering
\includegraphics[width=.6\textwidth]{mss-lin-figExpressionTwoGene}
\includegraphics[width=.6\textwidth]{mss-lin-figExpressionTwoGeneBetweenTranscript}
\caption{}
\label{figExpressionTwoGene}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figure 4. UA159 and TW1 in Glucose
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
<<figDotplotGlucoseOnly,fig=TRUE,echo=FALSE,width=12,height=6>>=
#par(mar=c(4.1, 3.1, 0.1, 0.1))
par(mai=c(0, 0, 0, 0))
y <- tableGlucoseOnly[,"Functional Class"]
z <- sapply(y, function(x) {x1=strsplit(as.character(x),"\\|\\|"); lapply(x1, trim)})
xm <- tableGlucoseOnly[rep(seq(length(z)),sapply(z,length)),]
xm[,"Functional Class"] <- unlist(z)
# Use xm to plot a dotplot.
dotplot(xm[,"Functional Class"]~xm[,2],
        main="",xlab="Log2 fold change",col="black",pch=88)
@
\begin{sidewaysfigure}
\centering
\includegraphics[width=0.9\textwidth]{mss-lin-figDotplotGlucoseOnly}
\caption{}
\label{figDotplotGlucoseOnly}
\end{sidewaysfigure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figure 5. Glucose and Galactose in UA159
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
<<figDotplotUA159Only,fig=TRUE,echo=FALSE,width=12,height=6>>=
par(mai=c(0, 0, 0, 0))
y <- tableUA159Only[,"Functional Class"]
z <- sapply(y, function(x) {x1=strsplit(as.character(x),"\\|\\|"); lapply(x1, trim)})
xm <- tableUA159Only[rep(seq(length(z)),sapply(z,length)),]
xm[,"Functional Class"] <- unlist(z)
# Use xm to plot a dotplot.
dotplot(xm[,"Functional Class"]~xm[,2], 
        main="",xlab="Log2 fold change",col="black",pch=88)
@
\begin{sidewaysfigure}
\centering
\includegraphics[width=0.9\textwidth]{mss-lin-figDotplotUA159Only}
\caption{}
\label{figDotplotUA159Only}
\end{sidewaysfigure}

\end{document}